### 2022-05-25 topic-ideas/topic-ideas.qmd
added heart dataset and glimpse!

---
title: "Topic ideas"
subtitle: "STA 210 - Project"
author: "Bayes' Harem - Christina Wang, Ethan Song, Kat Cottrell, David Goh"
format: pdf
editor: visual
---

```{r}
#| label: load-packages
#| message: false
#| warning: false

library(tidyverse)
library(tidymodels)
library(knitr)
library(ggfortify)
```

# Project idea 1

## Introduction and data

-   State the source of the data set.

-   Describe when and how it was originally collected (by the original data curator, not necessarily how you found the data)

-   Describe the observations and the general characteristics being measured in the data

## Research question

-   Describe a research question you're interested in answering using this data.

## Glimpse of data

-   Use the `glimpse` function to provide an overview of the dataset

```{r}
#| label: data-1

# add code to load and glimpse data here
```

\pagebreak

# Project idea 2

## Introduction and data

-   State the source of the data set.

-   Describe when and how it was originally collected (by the original data curator, not necessarily how you found the data)

-   Describe the observations and the general characteristics being measured in the data

## Research question

-   Describe a research question you're interested in answering using this data.

## Glimpse of data

-   Use the `glimpse` function to provide an overview of the dataset

```{r}
#| label: data-2

# add code to load and glimpse data here
```

\pagebreak

# Project idea 3

::: callout-important
Project idea 3 is optional. If you decide to submit only 2 ideas, please delete the section headings below and leave a note below stating so. If you decide to submit the 3rd idea, please delete this callout.
:::

## Introduction and data

-   State the source of the data set.

This dataset was published in this form by Kamil Pytlak, and it is taken from the Centers of Disease Control and Prevention (CDC) in the USA. It was retrieved from Pytlak's post on Kaggle.com. Citation: Kamil Pytlak. "Personal Key Indicators of Heart Disease". Kaggle. February 15 2022. Retrieved on May 25 2022 from <https://www.kaggle.com/datasets/kamilpytlak/personal-key-indicators-of-heart-disease/versions/2?resource=download>.

-   Describe when and how it was originally collected (by the original data curator, not necessarily how you found the data)

The data was collected by the Behavioral Risk Factor Surveillance System (BRFSS) which conducts **annual telephone surveys about the health status of U.S. residents in all 50 states, the District of Columbia and in the 3 US territories.** This data set includes data from **2020**, consisting of 401958 rows and 279 columns. Of note is that the classes are not balanced, and the author advises fixing the weights or undersampling. (Source: Pytlak, 2022)

-   Describe the observations and the general characteristics being measured in the data

The key indicators for the risk of heart disease include smoking, diabetic status, obesity (high BMI), not getting enough physical activity or drinking too much alcohol. In addition to these, the data records other information such as history of stroke, mental health, age category, race, general health, sleep time, kidney disease and skin cancer.

## Research question

-   Describe a research question you're interested in answering using this data.

What can we conclude about a person's likelihood of heart disease from their (1) BMI, (2) sex, (3) age category, (4) sleeping hours, (5) race, and (6) whether they are physically active?

Note that whether someone has heart disease is a binary variable. I have another research question prepared with the outcome variable set to a person's BMI, which is a continuous variable:

What can we conclude about a person's BMI from their (1) alcohol consumption, (2) sex, (3) age category, (4) sleeping hours, (5) race, and (6) whether they are physically active?

## Glimpse of data

-   Use the `glimpse` function to provide an overview of the dataset

```{r}
#| label: data-3

heart <- read_csv("data-3/heart_2020_cleaned.csv")
glimpse(heart)
```

### 2022-05-28 topic-ideas/topic-ideas.qmd
completed topic ideas

@@
 subtitle: "STA 210 - Project"
-author: "Bayes' Harem - Christina Wang, Ethan Song, Kat Cottrell, David Goh"
+author: "Bayes' Harem - Kat Cottrell, David Goh, Ethan Song, Christina Wang"
 format: pdf
@@
 
-```{r}
-#| label: load-packages
+```{r setup}
 #| message: false
@@
 
+# load packages
 library(tidyverse)
-library(tidymodels)
-library(knitr)
-library(ggfortify)
 ```
 
-# Project idea 1
+# Project idea 1 - Abortion attitudes
 
@@
 
+The dataset was put together by user svmiller on github. The user obtained the data from six waves of the World Values Survey (1982-2011). Link: https://github.com/svmiller/wvs-usa-abortion-attitudes
+
 -   Describe when and how it was originally collected (by the original data curator, not necessarily how you found the data)
 
+The data were collected as part of the World Values Survey, which is administered every few years and collects information about people’s values and beliefs worldwide.
+The survey aims to get a nationally representative sample of a minimum of 1200 for most countries, and the data are collected via face-to-face interviews at the respondents’ homes.
+The data included in this set specifically include responses from 6 waves of the survey (administered over the period 1982-2011).
+The responses included in this set are from people in the United States, and it examines their attitudes towards abortion.
+
 -   Describe the observations and the general characteristics being measured in the data
+
+The dataset includes 10,387 observations across all of the survey waves.
+The main variable that is assessed in this survey is abortion attitudes on a scale of 1-10 (1 being abortion is never justified, 10 being abortion is always justified).
+The survey also measures some demographic characteristics of the respondents (gender, education, age, etc.) and their attitudes on other issues (political ideology, their respect for authority, how important religion is to their lives, etc.)
 
@@
 
+How is an individual’s attitude towards abortion influenced by their demographic characteristics (such as age, gender, and education level) and personal attitudes towards other issues (such as political ideology, importance of religion to their lives, and their respect for authority)?
+
 ## Glimpse of data
@@
 
-```{r}
+```{r abortion-attitudes}
 #| label: data-1
 
-# add code to load and glimpse data here
+abortion_attitudes_data <- read_csv("data-1/wvs-usa-abortion-attitudes-data.csv")
+
+glimpse(abortion_attitudes_data)
 ```
@@
 
-# Project idea 2
+# Project idea 2 - Nutrition
 
@@
 
+This data comes from the National Health and Nutrition Examination Survey (NHANES), conducted in 2005-2006 by the Inter-university Consortium for Political and Social Research (ICPSR).
+It was retrieved from an ICPSR host page on the University of Michigan website.
+Possible variables used: DS218: Questionnaire: Drug Use, DS234: Questionnaire: Prescription Medications, DS229: Questionnaire: Osteoporosis.
+Link: https://www.icpsr.umich.edu/web/ICPSR/studies/25504/variables
+
 -   Describe when and how it was originally collected (by the original data curator, not necessarily how you found the data)
 
+The data were collected from approximately 5,000 interviewees who were first interviewed by trained administrators in the home and then later completed health examinations in mobile examination centers.
+The target population is average American non-institutionalized civilians.
+Low-income persons, adolescents 12-19 years of age, persons 60 years of age and older, African Americans, and Mexican Americans were oversampled in this data collection.
+
 -   Describe the observations and the general characteristics being measured in the data
+
+The NHANES 2005-2006 measured 7,449 variables pertaining to the health and nutritional status of Americans.  However, our research question only assesses injury and drug use variables.
 
@@
 -   Describe a research question you're interested in answering using this data.
+
+Does there exist any positive correlation between injuries and drug use?
 
@@
 
-# add code to load and glimpse data here
+drugs <- read.table(file ='data-2/25504-0218-Data.tsv', sep = '\t', header = TRUE)
+osteo <- read.table(file ='data-2/25504-0229-Data.tsv', sep = '\t', header = TRUE)
+rx    <- read.table(file ='data-2/25504-0234-Data.tsv', sep = '\t', header = TRUE)
+health <- inner_join(drugs, osteo, by=c('SEQN'='SEQN'))
+health <- inner_join(health, rx, by=c('SEQN'='SEQN'))
+
+glimpse(health)
 ```
@@
 
-# Project idea 3
-
-::: callout-important
-Project idea 3 is optional.
-If you decide to submit only 2 ideas, please delete the section headings below and leave a note below stating so.
-If you decide to submit the 3rd idea, please delete this callout.
-:::
+# Project idea 3 - Heart disease
 
@@
 This dataset was published in this form by Kamil Pytlak, and it is taken from the Centers of Disease Control and Prevention (CDC) in the USA.
-It was retrieved from Pytlak's post on Kaggle.com.
-Citation: Kamil Pytlak.
-"Personal Key Indicators of Heart Disease".
-Kaggle.
-February 15 2022.
-Retrieved on May 25 2022 from <https://www.kaggle.com/datasets/kamilpytlak/personal-key-indicators-of-heart-disease/versions/2?resource=download>.
+It was retrieved from Pytlak’s post on Kaggle.com.
+Link: https://www.kaggle.com/datasets/kamilpytlak/personal-key-indicators-of-heart-disease/versions/2?resource=download
 
@@
 
-The data was collected by the Behavioral Risk Factor Surveillance System (BRFSS) which conducts **annual telephone surveys about the health status of U.S.
-residents in all 50 states, the District of Columbia and in the 3 US territories.** This data set includes data from **2020**, consisting of 401958 rows and 279 columns.
+The data was collected by the Behavioral Risk Factor Surveillance System (BRFSS) which conducts annual telephone surveys about the health status of U.S.
+residents in all 50 states, the District of Columbia and in the 3 US territories.
+This data set includes data from 2020, consisting of 401958 rows and 279 columns.
 Of note is that the classes are not balanced, and the author advises fixing the weights or undersampling.
-(Source: Pytlak, 2022)
 
@@
 What can we conclude about a person's likelihood of heart disease from their (1) BMI, (2) sex, (3) age category, (4) sleeping hours, (5) race, and (6) whether they are physically active?
-
-Note that whether someone has heart disease is a binary variable. I have another research question prepared with the outcome variable set to a person's BMI, which is a continuous variable:
+Note that whether someone has heart disease is a binary variable. I have another research question prepared with the outcome variable set to a person’s BMI, which is a continuous variable:
 
@@
 
-heart <- read_csv("data-3/heart_2020_cleaned.csv")
-glimpse(heart)
+heart_disease_data <- read_csv("data-3/heart_2020_cleaned.csv")
+
+glimpse(heart_disease_data)
 ```

### 2022-06-06 proposal/proposal.qmd
copied the dataset to main data folder, explained eda

---
title: "Proposal"
subtitle: "STA 210 - Project"
author: "Bayes' Harem - Christina Wang, Kat Cottrell, David Goh, Ethan Song"
format: pdf
editor: source
---

```{r}
#| label: load-pkgs
#| message: false
#| warning: false
library(tidyverse)
library(tidymodels)
library(knitr)
library(ggfortify)
```

```{r load-data}
abortion_data_full <- read_csv(here::here("data/abortion-attitudes", "wvs-usa-abortion-attitudes-data.csv"))
```

## Introduction

[Christina Wang]

## Data description

[Kat Cottrell]

## Analysis approach

The response variable, “Justifiability of abortion”, is a numerical measure on a scale of 1-10 on the individual person’s attitude toward whether abortion is justifiable or not. The individuals responded “1” for “abortion is never justified” and “10” for “abortion is always justified.

Scrutinizing the data, we see that many survey questions were added only from the 1995 wave of the survey onward. To encompass all the predictor variables of the data set, we decided to remove the observations in generational waves before 1995 so that our data includes all the survey questions.

```{r filter-data}

abortion_data <- abortion_data_full %>% 
  filter(wave > "1994")
glimpse(abortion_data)

```

We now conduct exploratory data analysis (EDA) on each of the 15 predictor variables in the data set. The EDA for each variable comprises a histogram

## Data dictionary

The data dictionary can be found [here](ADD%20LINK!!!).

### 2022-06-06 proposal/proposal.qmd
finished section 3

@@
 library(ggfortify)
+library(GGally)
 ```
@@
 ```{r load-data}
+#| message: false
 abortion_data_full <- read_csv(here::here("data/abortion-attitudes", "wvs-usa-abortion-attitudes-data.csv"))
@@
 
-[Christina Wang]
+\[Christina Wang\]
 
@@
 
-[Kat Cottrell]
+\[Kat Cottrell\]
 
@@
 
-The response variable, “Justifiability of abortion”, is a numerical measure on a scale of 1-10 on the individual person’s attitude toward whether abortion is justifiable or not.
-The individuals responded “1” for “abortion is never justified” and “10” for “abortion is always justified.
+The response variable, **"Justifiability of abortion"**, is a numerical measure on a scale of 1-10 on the individual person's attitude toward whether abortion is justifiable or not.
+The individuals responded "1" for "abortion is never justified" and "10" for "abortion is always justified.
+
+We will conduct a **Multiple Linear Regression (MLR)** on this response variable against 6 other predictor variables for our project.
 
@@
 
+Moreover, the `wvsccode` for all observations is 840 because all surveys were conducted in the USA. We are hence removing it from our dataset.
+
 ```{r filter-data}
+abortion_data <- abortion_data_full %>% 
+  filter(year >= "1995") %>%
+  select(-starts_with("wave"), -starts_with("wvsccode"))
+```
 
-abortion_data <- abortion_data_full %>% 
-  filter(wave > "1994")
-glimpse(abortion_data)
+Before we select our predictor variables, we create a visualization and summary statistics for the response variable.
+
+``` {r visualization-summary}
+#| warning: false
+ggplot(abortion_data, aes(x = aj)) +
+  geom_histogram(binwidth = 1, aes(y=..density..)) + 
+  geom_density(color = "red") +
+  labs(title = "Histogram of abortion attitudes")
+
+summary(abortion_data$aj)
+```
+
+We now conduct exploratory data analysis (EDA) on each of the 15 predictor variables in the data set.
+The EDA for each variable comprises a histogram and a simple linear regression (SLR) against the outcome.
+
+```{r eda-year}
+#| warning: false
+#Year of survey
+ggplot(abortion_data, aes(x = year)) +
+  geom_histogram() + 
+  labs(title = "Histogram of year")
+ggplot(abortion_data, aes(x = year, y = aj)) +
+  geom_point() +
+  geom_smooth(method = "lm", se = FALSE)
+```
+
+```{r eda-age}
+#| warning: false
+#Age
+ggplot(abortion_data, aes(x = age)) +
+  geom_histogram(binwidth = 1) + 
+  labs(title = "Histogram of age")
+ggplot(abortion_data, aes(x = age, y = aj)) +
+  geom_point() +
+  geom_smooth(method = "lm", se = FALSE)
+```
+
+```{r eda-college}
+#| warning: false
+# College education
+ggplot(abortion_data, aes(x = collegeed)) +
+  geom_histogram() + 
+  labs(title = "Histogram of College education")
+ggplot(abortion_data, aes(x = collegeed, y = aj)) +
+  geom_point() +
+  geom_smooth(method = "lm", se = FALSE)
+```
+
+```{r eda-female}
+#| warning: false
+# Female
+ggplot(abortion_data, aes(x = female)) +
+  geom_histogram() + 
+  labs(title = "Histogram of female")
+ggplot(abortion_data, aes(x = female, y = aj)) +
+  geom_point() +
+  geom_smooth(method = "lm", se = FALSE)
+```
+
+```{r eda-unemployed}
+#| warning: false
+# Unemployed
+ggplot(abortion_data, aes(x = unemployed)) +
+  geom_histogram() + 
+  labs(title = "Histogram of unemployed")
+ggplot(abortion_data, aes(x = unemployed, y = aj)) +
+  geom_point() +
+  geom_smooth(method = "lm", se = FALSE)
+```
+
+```{r eda-ideology}
+#| warning: false
+# ideology
+ggplot(abortion_data, aes(x = ideology)) +
+  geom_histogram() + 
+  labs(title = "Histogram of ideology")
+ggplot(abortion_data, aes(x = ideology, y = aj)) +
+  geom_point() +
+  geom_smooth(method = "lm", se = FALSE)
+```
+
+```{r eda-financial}
+#| warning: false
+# Financial satisfaction
+ggplot(abortion_data, aes(x = satisfinancial)) +
+  geom_histogram() + 
+  labs(title = "Histogram of Financial satisfaction")
+ggplot(abortion_data, aes(x = satisfinancial, y = aj)) +
+  geom_point() +
+  geom_smooth(method = "lm", se = FALSE)
+```
+
+```{r eda-materialism}
+#| warning: false
+# WVS post-materialist index
+ggplot(abortion_data, aes(x = postma4)) +
+  geom_histogram() + 
+  labs(title = "Histogram of WVS post-materialist index")
+ggplot(abortion_data, aes(x = postma4, y = aj)) +
+  geom_point() +
+  geom_smooth(method = "lm", se = FALSE)
+```
+
+```{r eda-cai}
+#| warning: false
+# Child autonomy index
+ggplot(abortion_data, aes(x = cai)) +
+  geom_histogram() + 
+  labs(title = "Histogram of Child autonomy index")
+ggplot(abortion_data, aes(x = cai, y = aj)) +
+  geom_point() +
+  geom_smooth(method = "lm", se = FALSE)
+```
+
+```{r eda-trust}
+#| warning: false
+# Trust most people
+ggplot(abortion_data, aes(x = trustmostpeople)) +
+  geom_histogram() + 
+  labs(title = "Histogram of whether respondent trusted most people")
+ggplot(abortion_data, aes(x = trustmostpeople, y = aj)) +
+  geom_point() +
+  geom_smooth(method = "lm", se = FALSE)
+```
+
+```{r eda-god}
+#| warning: false
+# Importance of God
+ggplot(abortion_data, aes(x = godimportant)) +
+  geom_histogram() + 
+  labs(title = "Histogram of how respondent saw God's importance")
+ggplot(abortion_data, aes(x = godimportant, y = aj)) +
+  geom_point() +
+  geom_smooth(method = "lm", se = FALSE)
+```
+
+```{r eda-respect}
+#| warning: false
+# Respect for authority
+ggplot(abortion_data, aes(x = respectauthority)) +
+  geom_histogram() + 
+  labs(title = "Histogram of respect for authority")
+ggplot(abortion_data, aes(x = respectauthority, y = aj)) +
+  geom_point() +
+  geom_smooth(method = "lm", se = FALSE)
+```
+
+```{r eda-pride}
+#| warning: false
+# National Pride
+ggplot(abortion_data, aes(x = nationalpride)) +
+  geom_histogram() + 
+  labs(title = "Histogram of National pride")
+ggplot(abortion_data, aes(x = nationalpride, y = aj)) +
+  geom_point() +
+  geom_smooth(method = "lm", se = FALSE)
+```
+
+We now test if any of the predictors are strongly correlated with each other.
+
+```{r pair1}
+#| warning: false
+#I am doing ggpairs on 13 predictor variables. 
+#To ensure they fit into the screen, I will make 4 matrices with subsets
+
+ggpairs(abortion_data, 
+  columns = c("year", "age", "collegeed", "female", "unemployed",
+              "ideology", "satisfinancial"),
+  columnLabels = c("year", "age", "college", "female", 
+                   "unemployed", "ideology", "fin satis.")) +
+   theme(
+    axis.text.y = element_text(size = 10),
+    axis.text.x = element_text(angle = 45, size = 10),
+    strip.text.y = element_text(angle = 0, hjust = 0)
+    )
+```
+
+```{r pair2}
+#| warning: false
+ggpairs(abortion_data, 
+  columns = c("postma4", "cai", "trustmostpeople", "godimportant", 
+              "respectauthority","nationalpride"),
+  columnLabels = c("postmaterialism", "child autonomy", "trust", "god", 
+              "respectauthority","nationalpride")) +
+   theme(
+    axis.text.y = element_text(size = 10),
+    axis.text.x = element_text(angle = 45, size = 10),
+    strip.text.y = element_text(angle = 0, hjust = 0)
+    )
+```
+
+```{r pair3}
+#| warning: false
+ggpairs(abortion_data, 
+  columns = c("year", "age", "collegeed", "female", 
+              "respectauthority","nationalpride"),
+  columnLabels = c("year", "age", "collegeed", "female", 
+              "respectauthority","nationalpride")) +
+   theme(
+    axis.text.y = element_text(size = 10),
+    axis.text.x = element_text(angle = 45, size = 10),
+    strip.text.y = element_text(angle = 0, hjust = 0)
+    )
+```
+
+```{r pair4}
+#| warning: false
+ggpairs(abortion_data, 
+  columns = c("postma4", "cai", "trustmostpeople", "godimportant", 
+               "ideology", "satisfinancial"),
+  columnLabels = c("postmaterialism", "child autonomy", "trust", "god", 
+               "ideology", "satisfinancial")) +
+   theme(
+    axis.text.y = element_text(size = 10),
+    axis.text.x = element_text(angle = 45, size = 10),
+    strip.text.y = element_text(angle = 0, hjust = 0)
+    )
+```
+
+From these correlation matrices, we can conclude that the highest correlations (above 0.2) are those between `godimportant` and `cai` 0.370 `godimportant` and `respectauthority` 0.244 `godimportant` and `ideology` 0.243 `godimportant` and `nationalpride` 0.216 `nationalpride` and `age` 0.206 `nationalpride` and `respectauthority` 0.206 `respectauthority` and `cai` 0.205
+
+We want to choose predictor variables that are not strongly correlated with each other, are relatively balanced in our data set, and visually appear to have a significant relationship with the outcome.
+
+We now explore 3 different MLR models, each with 6 predictor variables chosen from the above.
+
+```{r model1}
+#| warning: false
+# predictors: year, age, female, ideology, cai, respectauthority
+model1_fit <- linear_reg() %>%
+  set_engine("lm") %>%
+  fit(aj ~ year + age + female + ideology + cai + respectauthority, data = abortion_data)
+
+glance(model1_fit) %>% select(r.squared, adj.r.squared, AIC, BIC)
 
@@
 
-We now conduct exploratory data analysis (EDA) on each of the 15 predictor variables in the data set. The EDA for each variable comprises a histogram
+```{r model2}
+#| warning: false
+# predictors: year, age, ideology, cai, respectauthority, nationalpride
+model2_fit <- linear_reg() %>%
+  set_engine("lm") %>%
+  fit(aj ~ year + age + ideology + cai + respectauthority + nationalpride, data = abortion_data)
+
+glance(model2_fit) %>% select(r.squared, adj.r.squared, AIC, BIC)
+
+```
+
+```{r model3}
+#| warning: false
+# predictors: year, age, female, ideology, cai, nationalpride
+model3_fit <- linear_reg() %>%
+  set_engine("lm") %>%
+  fit(aj ~ year + age + female + ideology + cai + nationalpride, data = abortion_data)
+
+glance(model3_fit) %>% select(r.squared, adj.r.squared, AIC, BIC)
+```
+
+Comparing the adjusted R-squared, AIC and BIC values, it is model 2 that has the highest adjusted R-squared and the lowest AIC and BIC of the 3 models.
+**Thus, we will use model 2 with the following predictors: year, age, ideology, cai, respectauthority, nationalpride.**
 

### 2022-06-08 proposal/proposal.qmd
completed introduction

@@
 
-\[Christina Wang\]
+Our project is on the topic of abortion attitudes in the US.
+Abortion has long been a divisive issue in the US, with both pro-choice and pro-life groups organizing in states across the country in support of political change that supported their respective sides (Ziegler, 2020).
+However, following the Supreme Court’s decision in Roe v.
+Wade in 1973, the issue has risen in political salience.
+Both the pro-choice and pro-life movement have gained national prominence and the two major political parties have polarized around the issue, with the Democratic Party in favor of and the Republican Party against policies legalizing and increasing access to abortion (Weinberger 2022).
+Abortion has also increasingly become a key issue that voters consider when they decide on candidates to vote for, with an increasing share of Americans becoming “single-issue voters” regarding abortion (Brenan 2020).
+
+In May 2022, a draft opinion was leaked revealing that the Supreme Court is prepared to overturn Roe v.
+Wade.
+If Roe is overturned, then it would dramatically change the trajectory of abortion politics in the US.
+Unless Congress passes a national policy, states would be able to decide whether or not to legalize abortion and would gain much greater leverage in regulating access to the procedure (Weinberger 2022).
+Given the potential overturning of Roe and the polarizing nature of the issue, it is important to understand how the American public feels about whether abortion should be legal or not, how accessible the procedure should be, and what factors influence these opinions.
+Understanding public opinion on the issue will ensure that policy experts are able to create and pass policies on this issue that accurately reflect the preferences of the American people.
+
+In this project, we are researching the factors that influence Americans’ abortion attitudes.
+Specifically, we would like to know which demographic factors (such as age, gender, and education level) and personal attitudes towards other issues (such as political ideology, importance of religion, and respect for authority) best predict an individual’s attitude towards abortion.
+Given that this issue has become highly polarized by political party, we predict that holding liberal political ideology and liberal attitudes on other issues, as well as being younger in age is correlated with believing that abortion is more justified.
+(the exact predictors we will use will be decided after model selection).
+
+Works Cited
+
+Brenan, M. (2020, July 7). One in Four Americans Consider Abortion a Key Voting Issue. Gallup. https://news.gallup.com/poll/313316/one-four-americans-consider-abortion-key-voting-issue.aspx
+
+Weinberger, J. (2022, May 6). How we got here: Roe v. Wade from 1973 to today. Vox. https://www.vox.com/23055389/roe-v-wade-timeline-abortion-overturn-political-polarization
+
+Ziegler, M.
+(2020, October 22).
+Abortion politics polarized before Roe.
+When it’s gone, the fighting won’t stop.
+The Washington Post.
+https://www.washingtonpost.com/outlook/2020/10/22/roe-polarize-abortion-politics/
 
@@
 
-``` {r visualization-summary}
+```{r visualization-summary}
 #| warning: false

### 2022-06-08 proposal/proposal.qmd
Added section 2\

@@
 
-\[Kat Cottrell\]
+The dataset observes “attitudes on the justifiability of abortion in the United States across six waves of World Values Survey data” (README.md) and some basic qualities of the respondants.
+
+Observations include:
+ - WVS country code
+ - Generational wave (1982, 1990, 1995, 1999, 2006, or 2011)
+ - Justifiability of abortion (1-10)
+ - Age (17 to 96)
+ - College graduate (1 for yes)
+ - Female (1 for women)
+ - Unemployed (1 = currently unemployed)
+ - Ideology (1-10 for left-right)
+ - Financial satisfaction (1-10 for least-most)
+ - WVS post-materialist index (-1 = materialist. 2 = mixed. 3 = post-materialist)
+ - Child autonomy index (-2 to 2 for obedience and religious faith-determination and independence)
+ - Trust (1 = believes most people can be trusted)
+ - Importance of God (1-10)
+ - Opinion of respect for authority (-1-1 for bad to good)
+ - National pride (1 = very proud to be an American)
+
+The data were collected as part of the World Values Survey, which is administered every few years and collects information about people’s values and beliefs worldwide.
+The survey aims to get a nationally representative sample of a minimum of 1200 for most countries, and the data are collected via face-to-face interviews at the respondents’ homes.
+The data included in this set specifically include responses from 6 waves of the survey (administered over the period 1982-2011).
+The responses included in this set are from people in the United States, and it examines their attitudes towards abortion.
 

### 2022-06-08 proposal/proposal.qmd
fixed bullets

--- 2022-06-08 proposal/proposal.qmd
@@
 editor: source
+editor_options: 
+  markdown: 
+    wrap: 72
 ---
@@
 
-Our project is on the topic of abortion attitudes in the US.
-Abortion has long been a divisive issue in the US, with both pro-choice and pro-life groups organizing in states across the country in support of political change that supported their respective sides (Ziegler, 2020).
-However, following the Supreme Court’s decision in Roe v.
-Wade in 1973, the issue has risen in political salience.
-Both the pro-choice and pro-life movement have gained national prominence and the two major political parties have polarized around the issue, with the Democratic Party in favor of and the Republican Party against policies legalizing and increasing access to abortion (Weinberger 2022).
-Abortion has also increasingly become a key issue that voters consider when they decide on candidates to vote for, with an increasing share of Americans becoming “single-issue voters” regarding abortion (Brenan 2020).
+Our project is on the topic of abortion attitudes in the US. Abortion
+has long been a divisive issue in the US, with both pro-choice and
+pro-life groups organizing in states across the country in support of
+political change that supported their respective sides (Ziegler, 2020).
+However, following the Supreme Court's decision in Roe v. Wade in 1973,
+the issue has risen in political salience. Both the pro-choice and
+pro-life movement have gained national prominence and the two major
+political parties have polarized around the issue, with the Democratic
+Party in favor of and the Republican Party against policies legalizing
+and increasing access to abortion (Weinberger 2022). Abortion has also
+increasingly become a key issue that voters consider when they decide on
+candidates to vote for, with an increasing share of Americans becoming
+"single-issue voters" regarding abortion (Brenan 2020).
 
-In May 2022, a draft opinion was leaked revealing that the Supreme Court is prepared to overturn Roe v.
-Wade.
-If Roe is overturned, then it would dramatically change the trajectory of abortion politics in the US.
-Unless Congress passes a national policy, states would be able to decide whether or not to legalize abortion and would gain much greater leverage in regulating access to the procedure (Weinberger 2022).
-Given the potential overturning of Roe and the polarizing nature of the issue, it is important to understand how the American public feels about whether abortion should be legal or not, how accessible the procedure should be, and what factors influence these opinions.
-Understanding public opinion on the issue will ensure that policy experts are able to create and pass policies on this issue that accurately reflect the preferences of the American people.
+In May 2022, a draft opinion was leaked revealing that the Supreme Court
+is prepared to overturn Roe v. Wade. If Roe is overturned, then it would
+dramatically change the trajectory of abortion politics in the US.
+Unless Congress passes a national policy, states would be able to decide
+whether or not to legalize abortion and would gain much greater leverage
+in regulating access to the procedure (Weinberger 2022). Given the
+potential overturning of Roe and the polarizing nature of the issue, it
+is important to understand how the American public feels about whether
+abortion should be legal or not, how accessible the procedure should be,
+and what factors influence these opinions. Understanding public opinion
+on the issue will ensure that policy experts are able to create and pass
+policies on this issue that accurately reflect the preferences of the
+American people.
 
-In this project, we are researching the factors that influence Americans’ abortion attitudes.
-Specifically, we would like to know which demographic factors (such as age, gender, and education level) and personal attitudes towards other issues (such as political ideology, importance of religion, and respect for authority) best predict an individual’s attitude towards abortion.
-Given that this issue has become highly polarized by political party, we predict that holding liberal political ideology and liberal attitudes on other issues, as well as being younger in age is correlated with believing that abortion is more justified.
-(the exact predictors we will use will be decided after model selection).
+In this project, we are researching the factors that influence
+Americans' abortion attitudes. Specifically, we would like to know which
+demographic factors (such as age, gender, and education level) and
+personal attitudes towards other issues (such as political ideology,
+importance of religion, and respect for authority) best predict an
+individual's attitude towards abortion. Given that this issue has become
+highly polarized by political party, we predict that holding liberal
+political ideology and liberal attitudes on other issues, as well as
+being younger in age is correlated with believing that abortion is more
+justified. (the exact predictors we will use will be decided after model
+selection).
 
@@
 
-Brenan, M. (2020, July 7). One in Four Americans Consider Abortion a Key Voting Issue. Gallup. https://news.gallup.com/poll/313316/one-four-americans-consider-abortion-key-voting-issue.aspx
+Brenan, M. (2020, July 7). One in Four Americans Consider Abortion a Key
+Voting Issue. Gallup.
+https://news.gallup.com/poll/313316/one-four-americans-consider-abortion-key-voting-issue.aspx
 
-Weinberger, J. (2022, May 6). How we got here: Roe v. Wade from 1973 to today. Vox. https://www.vox.com/23055389/roe-v-wade-timeline-abortion-overturn-political-polarization
+Weinberger, J. (2022, May 6). How we got here: Roe v. Wade from 1973 to
+today. Vox.
+https://www.vox.com/23055389/roe-v-wade-timeline-abortion-overturn-political-polarization
 
-Ziegler, M.
-(2020, October 22).
-Abortion politics polarized before Roe.
-When it’s gone, the fighting won’t stop.
-The Washington Post.
+Ziegler, M. (2020, October 22). Abortion politics polarized before Roe.
+When it's gone, the fighting won't stop. The Washington Post.
 https://www.washingtonpost.com/outlook/2020/10/22/roe-polarize-abortion-politics/
@@
 
-\[Kat Cottrell\]
+The dataset observes "attitudes on the justifiability of abortion in the
+United States across six waves of World Values Survey data" (README.md)
+and some basic qualities of the respondants.
+
+Observations include:
+
+-   WVS country code
+
+-   Generational wave (1982, 1990, 1995, 1999, 2006, or 2011)
+
+-   Justifiability of abortion (1-10)
+
+-   Age (17 to 96)
+
+-   College graduate (1 for yes)
+
+-   Female (1 for women) - Unemployed (1 = currently unemployed)
+
+-   Ideology (1-10 for left-right)
+
+-   Financial satisfaction (1-10 for least-most)
+
+```{=html}
+<!-- -->
+```
+-   WVS post-materialist index (-1 = materialist. 2 = mixed. 3 =
+    post-materialist)
+
+-   Child autonomy index (-2 to 2 for obedience and religious
+    faith-determination and independence)
+
+-   Trust (1 = believes most people can be trusted)
+
+-   Importance of God (1-10)
+
+-   Opinion of respect for authority (-1-1 for bad to good)
+
+-   National pride (1 = very proud to be an American)
+
+The data were collected as part of the World Values Survey, which is
+administered every few years and collects information about people's
+values and beliefs worldwide. The survey aims to get a nationally
+representative sample of a minimum of 1200 for most countries, and the
+data are collected via face-to-face interviews at the respondents'
+homes. The data included in this set specifically include responses from
+6 waves of the survey (administered over the period 1982-2011). The
+responses included in this set are from people in the United States, and
+it examines their attitudes towards abortion.
 
@@
 
-The response variable, **"Justifiability of abortion"**, is a numerical measure on a scale of 1-10 on the individual person's attitude toward whether abortion is justifiable or not.
-The individuals responded "1" for "abortion is never justified" and "10" for "abortion is always justified.
+The response variable, **"Justifiability of abortion"**, is a numerical
+measure on a scale of 1-10 on the individual person's attitude toward
+whether abortion is justifiable or not. The individuals responded "1"
+for "abortion is never justified" and "10" for "abortion is always
+justified.
 
-We will conduct a **Multiple Linear Regression (MLR)** on this response variable against 6 other predictor variables for our project.
+We will conduct a **Multiple Linear Regression (MLR)** on this response
+variable against 6 other predictor variables for our project.
 
-Scrutinizing the data, we see that many survey questions were added only from the 1995 wave of the survey onward.
-To encompass all the predictor variables of the data set, we decided to remove the observations in generational waves before 1995 so that our data includes all the survey questions.
+Scrutinizing the data, we see that many survey questions were added only
+from the 1995 wave of the survey onward. To encompass all the predictor
+variables of the data set, we decided to remove the observations in
+generational waves before 1995 so that our data includes all the survey
+questions.
 
-Moreover, the `wvsccode` for all observations is 840 because all surveys were conducted in the USA. We are hence removing it from our dataset.
+Moreover, the `wvsccode` for all observations is 840 because all surveys
+were conducted in the USA. We are hence removing it from our dataset.
 
@@
 
-Before we select our predictor variables, we create a visualization and summary statistics for the response variable.
+Before we select our predictor variables, we create a visualization and
+summary statistics for the response variable.
 
@@
 
-We now conduct exploratory data analysis (EDA) on each of the 15 predictor variables in the data set.
-The EDA for each variable comprises a histogram and a simple linear regression (SLR) against the outcome.
+We now conduct exploratory data analysis (EDA) on each of the 15
+predictor variables in the data set. The EDA for each variable comprises
+a histogram and a simple linear regression (SLR) against the outcome.
 
@@
 
-We now test if any of the predictors are strongly correlated with each other.
+We now test if any of the predictors are strongly correlated with each
+other.
 
@@
 
-From these correlation matrices, we can conclude that the highest correlations (above 0.2) are those between `godimportant` and `cai` 0.370 `godimportant` and `respectauthority` 0.244 `godimportant` and `ideology` 0.243 `godimportant` and `nationalpride` 0.216 `nationalpride` and `age` 0.206 `nationalpride` and `respectauthority` 0.206 `respectauthority` and `cai` 0.205
+From these correlation matrices, we can conclude that the highest
+correlations (above 0.2) are those between `godimportant` and `cai`
+0.370 `godimportant` and `respectauthority` 0.244 `godimportant` and
+`ideology` 0.243 `godimportant` and `nationalpride` 0.216
+`nationalpride` and `age` 0.206 `nationalpride` and `respectauthority`
+0.206 `respectauthority` and `cai` 0.205
 
-We want to choose predictor variables that are not strongly correlated with each other, are relatively balanced in our data set, and visually appear to have a significant relationship with the outcome.
+We want to choose predictor variables that are not strongly correlated
+with each other, are relatively balanced in our data set, and visually
+appear to have a significant relationship with the outcome.
 
-We now explore 3 different MLR models, each with 6 predictor variables chosen from the above.
+We now explore 3 different MLR models, each with 6 predictor variables
+chosen from the above.
 
@@
 
-Comparing the adjusted R-squared, AIC and BIC values, it is model 2 that has the highest adjusted R-squared and the lowest AIC and BIC of the 3 models.
-**Thus, we will use model 2 with the following predictors: year, age, ideology, cai, respectauthority, nationalpride.**
+Comparing the adjusted R-squared, AIC and BIC values, it is model 2 that
+has the highest adjusted R-squared and the lowest AIC and BIC of the 3
+models. **Thus, we will use model 2 with the following predictors: year,
+age, ideology, cai, respectauthority, nationalpride.**
 

### 2022-06-08 proposal/proposal.qmd
removed the augment analysis on the 3 models

@@
 
-We now explore 3 different MLR models, each with 6 predictor variables
-chosen from the above.
-
-```{r model1}
-#| warning: false
-# predictors: year, age, female, ideology, cai, respectauthority
-model1_fit <- linear_reg() %>%
-  set_engine("lm") %>%
-  fit(aj ~ year + age + female + ideology + cai + respectauthority, data = abortion_data)
-
-glance(model1_fit) %>% select(r.squared, adj.r.squared, AIC, BIC)
-
-```
-
-```{r model2}
-#| warning: false
-# predictors: year, age, ideology, cai, respectauthority, nationalpride
-model2_fit <- linear_reg() %>%
-  set_engine("lm") %>%
-  fit(aj ~ year + age + ideology + cai + respectauthority + nationalpride, data = abortion_data)
-
-glance(model2_fit) %>% select(r.squared, adj.r.squared, AIC, BIC)
-
-```
-
-```{r model3}
-#| warning: false
-# predictors: year, age, female, ideology, cai, nationalpride
-model3_fit <- linear_reg() %>%
-  set_engine("lm") %>%
-  fit(aj ~ year + age + female + ideology + cai + nationalpride, data = abortion_data)
-
-glance(model3_fit) %>% select(r.squared, adj.r.squared, AIC, BIC)
-```
-
-Comparing the adjusted R-squared, AIC and BIC values, it is model 2 that
-has the highest adjusted R-squared and the lowest AIC and BIC of the 3
-models. **Thus, we will use model 2 with the following predictors: year,
-age, ideology, cai, respectauthority, nationalpride.**
-
 ## Data dictionary

### 2022-06-10 proposal/CopyOfproposal.qmd
started editing the proposal

--- 2022-06-08 proposal/proposal.qmd
@@
 
+We can observe from this histogram that the distribution of the outcome variable is not a bell shape, and it is trimodal.
+This is likely because the question's phrasing is similar to a yes/no question, but respondents were asked to give their level of agreement on a scale of 1-10.
+This may result in our model not being a good fit for the data if we attempt a multiple linear regression model.
+We have two backup plans for this, if our MLR eventually has a poor performance.
+First, we can truncate this data into a categorical outcome variable such as (Agree, Disagree, Undecided), and conduct a binomial or multinomial logistic regression.
+Second, we can filter our population based on various characteristics (if we have good reason to do so) so that our outcome data follows a bell shape.
+
 We now conduct exploratory data analysis (EDA) on each of the 15
@@
   labs(title = "Histogram of year")
-ggplot(abortion_data, aes(x = year, y = aj)) +
-  geom_point() +
-  geom_smooth(method = "lm", se = FALSE)
+ggplot(abortion_data, aes(group = year, y = aj)) +
+  geom_boxplot() 
 ```
@@
   labs(title = "Histogram of age")
-ggplot(abortion_data, aes(x = age, y = aj)) +
-  geom_point() +
-  geom_smooth(method = "lm", se = FALSE)
+#ggplot(abortion_data, aes(group = age, y = aj)) +
+  #geom_density() try to group together
 ```
@@
 
+ Is there a plan B ? how to deal ? MLR compare
+
 ## Data dictionary

### 2022-06-10 proposal/proposal.qmd
data dictionary link added

@@
 
-Our project is on the topic of abortion attitudes in the US. Abortion
-has long been a divisive issue in the US, with both pro-choice and
-pro-life groups organizing in states across the country in support of
-political change that supported their respective sides (Ziegler, 2020).
-However, following the Supreme Court's decision in Roe v. Wade in 1973,
-the issue has risen in political salience. Both the pro-choice and
-pro-life movement have gained national prominence and the two major
-political parties have polarized around the issue, with the Democratic
-Party in favor of and the Republican Party against policies legalizing
-and increasing access to abortion (Weinberger 2022). Abortion has also
-increasingly become a key issue that voters consider when they decide on
-candidates to vote for, with an increasing share of Americans becoming
-"single-issue voters" regarding abortion (Brenan 2020).
+Our project is on the topic of abortion attitudes in the US.
+Abortion has long been a divisive issue in the US, with both pro-choice and pro-life groups organizing in states across the country in support of political change that supported their respective sides (Ziegler, 2020).
+However, following the Supreme Court’s decision in Roe v.
+Wade in 1973, the issue has risen in political salience.
+Both the pro-choice and pro-life movements have gained national prominence and the two major political parties have polarized around the issue, with the Democratic Party in favor of and the Republican Party against policies legalizing and increasing access to abortion (Weinberger 2022).
+Abortion has also increasingly become a key issue that voters consider when they decide on candidates to vote for, with an increasing share of Americans becoming “single-issue voters” regarding abortion (Brenan 2020).
 
-In May 2022, a draft opinion was leaked revealing that the Supreme Court
-is prepared to overturn Roe v. Wade. If Roe is overturned, then it would
-dramatically change the trajectory of abortion politics in the US.
-Unless Congress passes a national policy, states would be able to decide
-whether or not to legalize abortion and would gain much greater leverage
-in regulating access to the procedure (Weinberger 2022). Given the
-potential overturning of Roe and the polarizing nature of the issue, it
-is important to understand how the American public feels about whether
-abortion should be legal or not, how accessible the procedure should be,
-and what factors influence these opinions. Understanding public opinion
-on the issue will ensure that policy experts are able to create and pass
-policies on this issue that accurately reflect the preferences of the
-American people.
+In May 2022, a draft opinion was leaked revealing that the Supreme Court is prepared to overturn Roe v.
+Wade.
+If Roe is overturned, then it would dramatically change the trajectory of abortion politics in the US.
+Unless Congress passes a national policy, states would be able to decide whether or not to legalize abortion and would gain much greater leverage in regulating access to the procedure (Weinberger 2022).
+Given the potential overturning of Roe and the polarizing nature of the issue, it is important to understand how the American public feels about whether abortion should be legal or not, how accessible the procedure should be, and what factors influence these opinions.
+Understanding public opinion on the issue will ensure that political leaders will be able to mobilize the correct constituencies and that policy experts are able to pass policies on this issue that accurately reflect the preferences of the American people.
 
-In this project, we are researching the factors that influence
-Americans' abortion attitudes. Specifically, we would like to know which
-demographic factors (such as age, gender, and education level) and
-personal attitudes towards other issues (such as political ideology,
-importance of religion, and respect for authority) best predict an
-individual's attitude towards abortion. Given that this issue has become
-highly polarized by political party, we predict that holding liberal
-political ideology and liberal attitudes on other issues, as well as
-being younger in age is correlated with believing that abortion is more
-justified. (the exact predictors we will use will be decided after model
-selection).
+In this project, we are researching the factors that influence Americans’ abortion attitudes.
+Specifically, we would like to know which demographic factors (such as age, gender, and education level) and personal attitudes towards other issues (such as political ideology, importance of religion, and respect for authority) best predict an individual’s attitude towards abortion.
+We will conduct our analysis using data collected about Americans’ abortion attitudes, demographic information, as well as attitudes about other issues.
+The data are a nationally representative sample of the American people; hence we can infer that all the observations are independent and the variables may show a linear relationship.
+Given that this issue has become highly polarized by political party, we predict that holding liberal political ideology and liberal attitudes on other issues, as well as being younger in age is correlated with believing that abortion is more justified.
+(the exact predictors we will use will be decided after model selection).
 
@@
 Brenan, M. (2020, July 7). One in Four Americans Consider Abortion a Key
-Voting Issue. Gallup.
-https://news.gallup.com/poll/313316/one-four-americans-consider-abortion-key-voting-issue.aspx
+Voting Issue. Gallup. https://news.gallup.com/poll/313316/one-four-americans-consider-abortion-key-voting-issue.aspx
 
 Weinberger, J. (2022, May 6). How we got here: Roe v. Wade from 1973 to
-today. Vox.
-https://www.vox.com/23055389/roe-v-wade-timeline-abortion-overturn-political-polarization
+today. Vox. https://www.vox.com/23055389/roe-v-wade-timeline-abortion-overturn-political-polarization
 
 Ziegler, M. (2020, October 22). Abortion politics polarized before Roe.
-When it's gone, the fighting won't stop. The Washington Post.
-https://www.washingtonpost.com/outlook/2020/10/22/roe-polarize-abortion-politics/
+When it's gone, the fighting won't stop. The Washington Post. https://www.washingtonpost.com/outlook/2020/10/22/roe-polarize-abortion-politics/
 
@@
 
-```{=html}
 <!-- -->
-```
+
 -   WVS post-materialist index (-1 = materialist. 2 = mixed. 3 =
@@
 
-The data dictionary can be found [here](ADD%20LINK!!!).
+The data dictionary can be found
+[here](https://github.com/STA210-Summer22/project-bayes-harem/blob/main/data/README.md).

### 2022-06-10 proposal/proposal2.qmd
renamed proposal and changed analysis section

--- 2022-06-10 proposal/CopyOfproposal.qmd
@@
 #Age
+abortion_data <- abortion_data %>% 
+  mutate(agegroup = case_when(age >= 90 ~ '9',
+                                             age >= 80  & age <= 89 ~ '8',
+                                             age >= 70  & age <= 79 ~ '7',
+                                             age >= 60  & age <= 69 ~ '6',
+                                             age >= 50  & age <= 59 ~ '5',
+                                             age >= 40  & age <= 49 ~ '4',
+                                             age >= 30  & age <= 39 ~ '3',
+                                             age >= 20  & age <= 29 ~ '2',
+                              age <= 19 ~ '1'))
+abortion_data
 ggplot(abortion_data, aes(x = age)) +
@@
   labs(title = "Histogram of age")
-#ggplot(abortion_data, aes(group = age, y = aj)) +
-  #geom_density() try to group together
+ggplot(abortion_data, aes(x = agegroup, group = agegroup, y = aj)) +
+  geom_boxplot()
 ```
@@
   labs(title = "Histogram of College education")
-ggplot(abortion_data, aes(x = collegeed, y = aj)) +
-  geom_point() +
-  geom_smooth(method = "lm", se = FALSE)
+ggplot(abortion_data, aes(x = collegeed, group = collegeed, y = aj)) +
+  geom_boxplot()
 ```
@@
   labs(title = "Histogram of female")
-ggplot(abortion_data, aes(x = female, y = aj)) +
-  geom_point() +
-  geom_smooth(method = "lm", se = FALSE)
+ggplot(abortion_data, aes(x = female,group = female, y = aj)) +
+  geom_boxplot()
 ```
@@
   labs(title = "Histogram of unemployed")
-ggplot(abortion_data, aes(x = unemployed, y = aj)) +
-  geom_point() +
-  geom_smooth(method = "lm", se = FALSE)
+ggplot(abortion_data, aes(x = unemployed, group = unemployed, y = aj)) +
+  geom_boxplot()
 ```
@@
   labs(title = "Histogram of ideology")
-ggplot(abortion_data, aes(x = ideology, y = aj)) +
-  geom_point() +
-  geom_smooth(method = "lm", se = FALSE)
+ggplot(abortion_data, aes(x = ideology,group = ideology, y = aj)) +
+  geom_boxplot()
 ```
@@
   labs(title = "Histogram of Financial satisfaction")
-ggplot(abortion_data, aes(x = satisfinancial, y = aj)) +
-  geom_point() +
-  geom_smooth(method = "lm", se = FALSE)
+ggplot(abortion_data, aes(x = satisfinancial,group = satisfinancial, y = aj)) +
+  geom_boxplot()
 ```
@@
   labs(title = "Histogram of WVS post-materialist index")
-ggplot(abortion_data, aes(x = postma4, y = aj)) +
-  geom_point() +
-  geom_smooth(method = "lm", se = FALSE)
+ggplot(abortion_data, aes(x = postma4, group = postma4,y = aj)) +
+  geom_boxplot()
 ```
@@
   labs(title = "Histogram of Child autonomy index")
-ggplot(abortion_data, aes(x = cai, y = aj)) +
-  geom_point() +
-  geom_smooth(method = "lm", se = FALSE)
+ggplot(abortion_data, aes(x = cai, group = cai, y = aj)) +
+  geom_boxplot()
 ```
@@
   labs(title = "Histogram of whether respondent trusted most people")
-ggplot(abortion_data, aes(x = trustmostpeople, y = aj)) +
-  geom_point() +
-  geom_smooth(method = "lm", se = FALSE)
+ggplot(abortion_data, aes(x = trustmostpeople, group = trustmostpeople, y = aj)) +
+  geom_boxplot()
 ```
@@
   labs(title = "Histogram of how respondent saw God's importance")
-ggplot(abortion_data, aes(x = godimportant, y = aj)) +
-  geom_point() +
-  geom_smooth(method = "lm", se = FALSE)
+ggplot(abortion_data, aes(x = godimportant, group = godimportant, y = aj)) +
+  geom_boxplot()
 ```
@@
   labs(title = "Histogram of respect for authority")
-ggplot(abortion_data, aes(x = respectauthority, y = aj)) +
-  geom_point() +
-  geom_smooth(method = "lm", se = FALSE)
+ggplot(abortion_data, aes(x = respectauthority, group = respectauthority, y = aj)) +
+  geom_boxplot()
 ```
@@
   labs(title = "Histogram of National pride")
-ggplot(abortion_data, aes(x = nationalpride, y = aj)) +
-  geom_point() +
-  geom_smooth(method = "lm", se = FALSE)
+ggplot(abortion_data, aes(x = nationalpride, group = nationalpride, y = aj)) +
+  geom_boxplot()
 ```
@@
 
-We want to choose predictor variables that are not strongly correlated
-with each other, are relatively balanced in our data set, and visually
-appear to have a significant relationship with the outcome.
-
- Is there a plan B ? how to deal ? MLR compare
+We want to choose predictor variables that are not strongly correlated with each other, are relatively balanced in our data set, and visually appear to have a significant relationship with the outcome.
+Now that we have calculated the correlation matrix and know that some variables may have multicollinearity, we will select variables that do not exhibit significant multicollinearity for our model.
+Moreover, we can run several multiple linear regression models with different combinations of predictors to see which model performs best and is least affected by multicollinearity.
 

### 2022-06-13 draft/draft.qmd
created new qmd called draft

--- 2022-06-10 proposal/proposal2.qmd
@@
 ---
-title: "Proposal"
+title: "Draft"
 subtitle: "STA 210 - Project"

### 2022-06-13 draft/draft.qmd
some items

--- 2022-06-10 proposal/proposal2.qmd
@@
 
-Our project is on the topic of abortion attitudes in the US. Abortion
-has long been a divisive issue in the US, with both pro-choice and
-pro-life groups organizing in states across the country in support of
-political change that supported their respective sides (Ziegler, 2020).
-However, following the Supreme Court's decision in Roe v. Wade in 1973,
-the issue has risen in political salience. Both the pro-choice and
-pro-life movement have gained national prominence and the two major
-political parties have polarized around the issue, with the Democratic
-Party in favor of and the Republican Party against policies legalizing
-and increasing access to abortion (Weinberger 2022). Abortion has also
-increasingly become a key issue that voters consider when they decide on
-candidates to vote for, with an increasing share of Americans becoming
-"single-issue voters" regarding abortion (Brenan 2020).
-
-In May 2022, a draft opinion was leaked revealing that the Supreme Court
-is prepared to overturn Roe v. Wade. If Roe is overturned, then it would
-dramatically change the trajectory of abortion politics in the US.
-Unless Congress passes a national policy, states would be able to decide
-whether or not to legalize abortion and would gain much greater leverage
-in regulating access to the procedure (Weinberger 2022). Given the
-potential overturning of Roe and the polarizing nature of the issue, it
-is important to understand how the American public feels about whether
-abortion should be legal or not, how accessible the procedure should be,
-and what factors influence these opinions. Understanding public opinion
-on the issue will ensure that policy experts are able to create and pass
-policies on this issue that accurately reflect the preferences of the
-American people.
-
-In this project, we are researching the factors that influence
-Americans' abortion attitudes. Specifically, we would like to know which
-demographic factors (such as age, gender, and education level) and
-personal attitudes towards other issues (such as political ideology,
-importance of religion, and respect for authority) best predict an
-individual's attitude towards abortion. Given that this issue has become
-highly polarized by political party, we predict that holding liberal
-political ideology and liberal attitudes on other issues, as well as
-being younger in age is correlated with believing that abortion is more
-justified. (the exact predictors we will use will be decided after model
-selection).
-
-Works Cited
-
-Brenan, M. (2020, July 7). One in Four Americans Consider Abortion a Key
-Voting Issue. Gallup.
-https://news.gallup.com/poll/313316/one-four-americans-consider-abortion-key-voting-issue.aspx
-
-Weinberger, J. (2022, May 6). How we got here: Roe v. Wade from 1973 to
-today. Vox.
-https://www.vox.com/23055389/roe-v-wade-timeline-abortion-overturn-political-polarization
-
-Ziegler, M. (2020, October 22). Abortion politics polarized before Roe.
-When it's gone, the fighting won't stop. The Washington Post.
-https://www.washingtonpost.com/outlook/2020/10/22/roe-polarize-abortion-politics/
+Paste-updated-content-here
 
@@
 
-The response variable, **"Justifiability of abortion"**, is a numerical
-measure on a scale of 1-10 on the individual person's attitude toward
-whether abortion is justifiable or not. The individuals responded "1"
-for "abortion is never justified" and "10" for "abortion is always
-justified.
-
-We will conduct a **Multiple Linear Regression (MLR)** on this response
-variable against 6 other predictor variables for our project.
-
-Scrutinizing the data, we see that many survey questions were added only
-from the 1995 wave of the survey onward. To encompass all the predictor
-variables of the data set, we decided to remove the observations in
-generational waves before 1995 so that our data includes all the survey
-questions.
-
-Moreover, the `wvsccode` for all observations is 840 because all surveys
-were conducted in the USA. We are hence removing it from our dataset.
+Paste-updated-content-here
 

### 2022-06-15 draft/draft.qmd
improved line spacing with correlation list
added introduction and references section

--- 2022-06-13 draft/draft.qmd
@@
 
-## Introduction
+## Introduction and Research Question
 
@@
 Specifically, we would like to know which demographic factors (such as age, gender, and education level) and personal attitudes towards other issues (such as political ideology, importance of religion, and respect for authority) best predict an individual’s attitude towards abortion.
-We will conduct our analysis using data collected about Americans’ abortion attitudes, demographic information, as well as attitudes about other issues.
+We will conduct our analysis using data collected through the World Values Survey about Americans’ abortion attitudes, demographic information, as well as attitudes about other issues.
 The data are a nationally representative sample of the American people; hence we can infer that all the observations are independent and the variables may show a linear relationship.
@@
 (the exact predictors we will use will be decided after model selection).
-
-Works Cited
-
-Brenan, M. (2020, July 7). One in Four Americans Consider Abortion a Key
-Voting Issue. Gallup. https://news.gallup.com/poll/313316/one-four-americans-consider-abortion-key-voting-issue.aspx
-
-Weinberger, J. (2022, May 6). How we got here: Roe v. Wade from 1973 to
-today. Vox. https://www.vox.com/23055389/roe-v-wade-timeline-abortion-overturn-political-polarization
-
-Ziegler, M. (2020, October 22). Abortion politics polarized before Roe.
-When it's gone, the fighting won't stop. The Washington Post. https://www.washingtonpost.com/outlook/2020/10/22/roe-polarize-abortion-politics/
 
@@
 From these correlation matrices, we can conclude that the highest
-correlations (above 0.2) are those between `godimportant` and `cai`
-0.370 `godimportant` and `respectauthority` 0.244 `godimportant` and
-`ideology` 0.243 `godimportant` and `nationalpride` 0.216
-`nationalpride` and `age` 0.206 `nationalpride` and `respectauthority`
-0.206 `respectauthority` and `cai` 0.205
+correlations (above 0.2) are those between 
+`godimportant` and `cai` 0.370 
+`godimportant` and `respectauthority` 0.244 
+`godimportant` and `ideology` 0.243 
+`godimportant` and `nationalpride` 0.216
+`nationalpride` and `age` 0.206 
+`nationalpride` and `respectauthority` 0.206 
+`respectauthority` and `cai` 0.205
 
@@
 The data dictionary can be found [here](ADD%20LINK!!!).
+
+## References
+
+Brenan, M. (2020, July 7). One in Four Americans Consider Abortion a Key
+Voting Issue. Gallup. https://news.gallup.com/poll/313316/one-four-americans-consider-abortion-key-voting-issue.aspx
+
+Weinberger, J. (2022, May 6). How we got here: Roe v. Wade from 1973 to
+today. Vox. https://www.vox.com/23055389/roe-v-wade-timeline-abortion-overturn-political-polarization
+
+Ziegler, M. (2020, October 22). Abortion politics polarized before Roe.
+When it's gone, the fighting won't stop. The Washington Post. https://www.washingtonpost.com/outlook/2020/10/22/roe-polarize-abortion-politics/

### 2022-06-15 draft/draft.qmd
finished eda and method writeup

--- 2022-06-13 draft/draft.qmd
@@
 
-Paste-updated-content-here
-
-```{r filter-data}
-abortion_data <- abortion_data_full %>% 
-  filter(year >= "1995") %>%
-  select(-starts_with("wave"), -starts_with("wvsccode"))
-```
-
-Before we select our predictor variables, we create a visualization and
-summary statistics for the response variable.
+We create a visualization and summary statistics for the response variable.
 
@@
 
-We can observe from this histogram that the distribution of the outcome variable is not a bell shape, and it is trimodal.
-This is likely because the question's phrasing is similar to a yes/no question, but respondents were asked to give their level of agreement on a scale of 1-10.
-This may result in our model not being a good fit for the data if we attempt a multiple linear regression model.
-We have two backup plans for this, if our MLR eventually has a poor performance.
-First, we can truncate this data into a categorical outcome variable such as (Agree, Disagree, Undecided), and conduct a binomial or multinomial logistic regression.
-Second, we can filter our population based on various characteristics (if we have good reason to do so) so that our outcome data follows a bell shape.
+We can observe from this histogram that the distribution of the outcome
+variable is not a bell shape, and it is trimodal. This is likely because
+the question's phrasing is similar to a yes/no question, but respondents
+were asked to give their level of agreement on a scale of 1-10. This may
+result in our model not being a good fit for the data if we attempt a
+multiple linear regression model. We have two backup plans for this, if
+our MLR eventually has a poor performance. First, we can truncate this
+data into a categorical outcome variable such as (Agree, Disagree,
+Undecided), and conduct a binomial or multinomial logistic regression.
+Second, we can filter our population based on various characteristics
+(if we have good reason to do so) so that our outcome data follows a
+bell shape.
 
-We now conduct exploratory data analysis (EDA) on each of the 15
-predictor variables in the data set. The EDA for each variable comprises
-a histogram and a simple linear regression (SLR) against the outcome.
+We now extend our exploratory data analysis (EDA) to some predictor
+variables of interest; namely, the year, ideology, Child Autonomy Index,
+Importance of God, Respect for Authority and National Pride predictors.
+The EDA for each variable comprises a histogram and a boxplot of the
+response variable grouped by predictor value. Additionally, we have a
+jitter plot to explore the potential for an interaction effect between
+year and importance of God.
 
@@
 ggplot(abortion_data, aes(group = year, y = aj)) +
-  geom_boxplot() 
-```
-
-```{r eda-age}
-#| warning: false
-#Age
-abortion_data <- abortion_data %>% 
-  mutate(agegroup = case_when(age >= 90 ~ '9',
-                                             age >= 80  & age <= 89 ~ '8',
-                                             age >= 70  & age <= 79 ~ '7',
-                                             age >= 60  & age <= 69 ~ '6',
-                                             age >= 50  & age <= 59 ~ '5',
-                                             age >= 40  & age <= 49 ~ '4',
-                                             age >= 30  & age <= 39 ~ '3',
-                                             age >= 20  & age <= 29 ~ '2',
-                              age <= 19 ~ '1'))
-abortion_data
-ggplot(abortion_data, aes(x = age)) +
-  geom_histogram(binwidth = 1) + 
-  labs(title = "Histogram of age")
-ggplot(abortion_data, aes(x = agegroup, group = agegroup, y = aj)) +
-  geom_boxplot()
-```
-
-```{r eda-college}
-#| warning: false
-# College education
-ggplot(abortion_data, aes(x = collegeed)) +
-  geom_histogram() + 
-  labs(title = "Histogram of College education")
-ggplot(abortion_data, aes(x = collegeed, group = collegeed, y = aj)) +
-  geom_boxplot()
-```
-
-```{r eda-female}
-#| warning: false
-# Female
-ggplot(abortion_data, aes(x = female)) +
-  geom_histogram() + 
-  labs(title = "Histogram of female")
-ggplot(abortion_data, aes(x = female,group = female, y = aj)) +
-  geom_boxplot()
-```
-
-```{r eda-unemployed}
-#| warning: false
-# Unemployed
-ggplot(abortion_data, aes(x = unemployed)) +
-  geom_histogram() + 
-  labs(title = "Histogram of unemployed")
-ggplot(abortion_data, aes(x = unemployed, group = unemployed, y = aj)) +
   geom_boxplot() 
@@
 
-```{r eda-financial}
-#| warning: false
-# Financial satisfaction
-ggplot(abortion_data, aes(x = satisfinancial)) +
-  geom_histogram() + 
-  labs(title = "Histogram of Financial satisfaction")
-ggplot(abortion_data, aes(x = satisfinancial,group = satisfinancial, y = aj)) +
-  geom_boxplot()
-```
-
-```{r eda-materialism}
-#| warning: false
-# WVS post-materialist index
-ggplot(abortion_data, aes(x = postma4)) +
-  geom_histogram() + 
-  labs(title = "Histogram of WVS post-materialist index")
-ggplot(abortion_data, aes(x = postma4, group = postma4,y = aj)) +
-  geom_boxplot()
-```
-
 ```{r eda-cai}
@@
 ggplot(abortion_data, aes(x = cai, group = cai, y = aj)) +
-  geom_boxplot()
-```
-
-```{r eda-trust}
-#| warning: false
-# Trust most people
-ggplot(abortion_data, aes(x = trustmostpeople)) +
-  geom_histogram() + 
-  labs(title = "Histogram of whether respondent trusted most people")
-ggplot(abortion_data, aes(x = trustmostpeople, group = trustmostpeople, y = aj)) +
   geom_boxplot()
@@
 
-```{r pair1}
-#| warning: false
-#I am doing ggpairs on 13 predictor variables. 
-#To ensure they fit into the screen, I will make 4 matrices with subsets
-
-ggpairs(abortion_data, 
-  columns = c("year", "age", "collegeed", "female", "unemployed",
-              "ideology", "satisfinancial"),
-  columnLabels = c("year", "age", "college", "female", 
-                   "unemployed", "ideology", "fin satis.")) +
-   theme(
-    axis.text.y = element_text(size = 10),
-    axis.text.x = element_text(angle = 45, size = 10),
-    strip.text.y = element_text(angle = 0, hjust = 0)
-    )
-```
-
-```{r pair2}
+```{r correlations}
 #| warning: false
 ggpairs(abortion_data, 
-  columns = c("postma4", "cai", "trustmostpeople", "godimportant", 
-              "respectauthority","nationalpride"),
-  columnLabels = c("postmaterialism", "child autonomy", "trust", "god", 
-              "respectauthority","nationalpride")) +
-   theme(
-    axis.text.y = element_text(size = 10),
-    axis.text.x = element_text(angle = 45, size = 10),
-    strip.text.y = element_text(angle = 0, hjust = 0)
-    )
-```
-
-```{r pair3}
-#| warning: false
-ggpairs(abortion_data, 
-  columns = c("year", "age", "collegeed", "female", 
-              "respectauthority","nationalpride"),
-  columnLabels = c("year", "age", "collegeed", "female", 
-              "respectauthority","nationalpride")) +
-   theme(
-    axis.text.y = element_text(size = 10),
-    axis.text.x = element_text(angle = 45, size = 10),
-    strip.text.y = element_text(angle = 0, hjust = 0)
-    )
-```
-
-```{r pair4}
-#| warning: false
-ggpairs(abortion_data, 
-  columns = c("postma4", "cai", "trustmostpeople", "godimportant", 
-               "ideology", "satisfinancial"),
-  columnLabels = c("postmaterialism", "child autonomy", "trust", "god", 
-               "ideology", "satisfinancial")) +
+  columns = c("year", "ideology", "cai", "godimportant", "respectauthority", 
+              "nationalpride"),
+  columnLabels = c("year", "ideology", "child autonomy", "imptnce of God", "resp for autho", 
+              "natl pride")) +
    theme(
@@
 `ideology` 0.243 `godimportant` and `nationalpride` 0.216
-`nationalpride` and `age` 0.206 `nationalpride` and `respectauthority`
-0.206 `respectauthority` and `cai` 0.205
+`nationalpride` and `respectauthority` 0.206 `respectauthority` and
+`cai` 0.205
 
-We want to choose predictor variables that are not strongly correlated with each other, are relatively balanced in our data set, and visually appear to have a significant relationship with the outcome.
-Now that we have calculated the correlation matrix and know that some variables may have multicollinearity, we will select variables that do not exhibit significant multicollinearity for our model.
-Moreover, we can run several multiple linear regression models with different combinations of predictors to see which model performs best and is least affected by multicollinearity.
+\pagebreak
+
+## METHOD
+
+We will conduct a Multiple Linear Regression (MLR) on the attitudes on
+the justifiability of abortion against several other predictor variables
+found in this data set. This is because our outcome variable, the
+attitude towards abortion, is measured on a numeric scale from 1 to 10,
+and there are multiple predictor variables that we feel are potentially
+correlated with it.
+
+We will now discuss the data cleaning and predictor selection process.
+For observations, we removed the observations that were part of
+"generational waves" before 1995, because many survey questions were
+only added from 1995 onward, and observations before 1995 have a high
+number of null values.
+
+We removed the variable "wvsccode" as it is the same for all the
+observations. The WVSC Code depends on the country the surveys were
+conducted in, and all surveys in this data set were conducted in the
+United States of America.
+
+```{r filter-data}
+abortion_data <- abortion_data_full %>% 
+  filter(year >= "1995") %>%
+  select(-starts_with("wave"), -starts_with("wvsccode"))
+```
+
+We chose to consider predictor variables that, from our exploratory data
+analysis (EDA), looked like they were potentially correlated with the
+outcome. Because the predictor variables were discrete, our EDA was
+presented as a series of box plots. We observed the differences in
+median and quartile values of abortion attitude across different
+predictor values and ruled out variables with no visible hint of
+relationship.
+
+This led to our first set of predictor variables, which we designate as
+Recipe1: 
+- Year 
+- Ideology 
+- Child Autonomy Index 
+- Importance of God 
+- Respect for Authority 
+- National Pride
+
+We also wanted to choose predictor variables that are not strongly
+correlated with each other, to avoid multicollinearity. From our EDA, we
+found that several predictors had correlations of 0.2 or higher with
+each other. Given these correlations, we decided to make a second recipe
+(`Recipe2`) with the following predictor variables, after excluding
+those with a high correlation with the `godimportant` variable:
+
+-   Year
+-   Importance of God
+
+Finally, we considered that the distribution for the `godimportant`
+variable is heavily left-skewed in our data set, which we observed in
+the EDA. This may reduce the ability of a model that relies on the
+"importance of God" predictor to explain the variations we observe in
+the outcome. The same can be said of the `respectauthority` variable,
+with more than 8 times the observations indicating "1" compared to the
+observations indicating "-1".
+
+Consequently, our third recipe (Recipe3) excludes `godimportant` and
+uses the other variables that are not correlated with each other more
+than 0.200:
+
+-   Year
+-   Ideology
+-   Child Autonomy Index
+-   National Pride
+
+## Data Split
+
+## Fit Model
+
+Note: make sure to create recipes, specify engine, make workflows, and
+fit models here
+
+## Inference
+
+\pagebreak
+
+##RESULTS
+
+## Model Evaluation
+
+## Model Comparison
+
+## Model Diagnosis
+
+## Discussion of Influential Points
 

### 2022-06-15 draft/draft.qmd

@@
 
-## Introduction
+## Introduction and Research Question
 
-Paste-updated-content-here
+Our project is on the topic of abortion attitudes in the US.
+Abortion has long been a divisive issue in the US, with both pro-choice and pro-life groups organizing in states across the country in support of political change that supported their respective sides (Ziegler, 2020).
+However, following the Supreme Court’s decision in Roe v.
+Wade in 1973, the issue has risen in political salience.
+Both the pro-choice and pro-life movements have gained national prominence and the two major political parties have polarized around the issue, with the Democratic Party in favor of and the Republican Party against policies legalizing and increasing access to abortion (Weinberger 2022).
+Abortion has also increasingly become a key issue that voters consider when they decide on candidates to vote for, with an increasing share of Americans becoming “single-issue voters” regarding abortion (Brenan 2020).
+
+In May 2022, a draft opinion was leaked revealing that the Supreme Court is prepared to overturn Roe v.
+Wade.
+If Roe is overturned, then it would dramatically change the trajectory of abortion politics in the US.
+Unless Congress passes a national policy, states would be able to decide whether or not to legalize abortion and would gain much greater leverage in regulating access to the procedure (Weinberger 2022).
+Given the potential overturning of Roe and the polarizing nature of the issue, it is important to understand how the American public feels about whether abortion should be legal or not, how accessible the procedure should be, and what factors influence these opinions.
+Understanding public opinion on the issue will ensure that political leaders will be able to mobilize the correct constituencies and that policy experts are able to pass policies on this issue that accurately reflect the preferences of the American people.
+
+In this project, we are researching the factors that influence Americans’ abortion attitudes.
+Specifically, we would like to know which demographic factors (such as age, gender, and education level) and personal attitudes towards other issues (such as political ideology, importance of religion, and respect for authority) best predict an individual’s attitude towards abortion.
+We will conduct our analysis using data collected through the World Values Survey about Americans’ abortion attitudes, demographic information, as well as attitudes about other issues.
+The data are a nationally representative sample of the American people; hence we can infer that all the observations are independent and the variables may show a linear relationship.
+Given that this issue has become highly polarized by political party, we predict that holding liberal political ideology and liberal attitudes on other issues, as well as being younger in age is correlated with believing that abortion is more justified.
+(the exact predictors we will use will be decided after model selection).
 

### 2022-06-15 draft/draft.qmd
latest version

--- 2022-06-15 draft/draft.qmd
@@
 
-## Introduction
+## Introduction and Research Question
 
-Paste-updated-content-here
+# Motivation
+
+Understanding public attitudes on divisive political issues is an important way for political leaders to mobilize voters and for lawmakers to draft laws that represent their constituents.
+While it is simple to poll constituents’ positions on an issue, it can be challenging to assess the complex factors that influence and predict those stances.
+
+Abortion is one such divisive issue in the United States, with strong organization of pro-choice and pro-life groups in states across the country in support of legislation for their respective sides (Ziegler, 2020).
+However, following the Supreme Court’s 1973 decision in Roe v.
+Wade, a ruling that has historically protected women's rights to abortion without excessive government restriction in America, the issue has risen in political salience.
+Both the pro-choice and pro-life movements have gained national prominence, and the two major political parties have polarized around the issue, with the Democratic Party in favor of and the Republican Party against policies legalization and increased access to abortion (Weinberger 2022).
+Abortion has also increasingly become a key issue that voters consider, with an increasing share of Americans identifying as “single-issue voters” regarding abortion (Brenan 2020).
+
+In May 2022, a leaked draft opinion revealed that the US Supreme Court is prepared to overturn Roe v.
+Wade.
+Overturning Roe would dramatically change the trajectory of abortion politics in the US.
+Unless the US Congress passed a national policy, states would be able to decide whether or not to legalize abortion and gain much greater leverage in regulating access to the procedure (Weinberger 2022).
+
+Given the potential overturning of Roe and the polarizing nature of the issue, it is important to understand how the American public feels about whether abortion should be legal or not, how accessible the procedure should be, and what factors influence these opinions.
+Understanding public opinion on the issue will ensure that political leaders are able to mobilize the correct constituencies, and that policy experts are able to pass policies on this issue that accurately reflect the preferences of the American people.
 
@@
 
+We will conduct our analysis on data collected about Americans’ abortion attitudes, demographic information, and other ideologies, using an EDA-informed predictive model.
+The data are a nationally representative sample of the American people; hence we can infer that all the observations are independent and the variables may show a linear relationship.
+Given that this issue has become highly polarized by political party, we predict that liberal political attitudes and youth will correlate with belief that abortion is more justified.
+
 The dataset observes "attitudes on the justifiability of abortion in the
 United States across six waves of World Values Survey data" (README.md)
-and some basic qualities of the respondants.
+and some basic qualities of the respondents.
 
@@
 -   Financial satisfaction (1-10 for least-most)
-
-<!-- -->
 
@@
 
-## Analysis approach
+## Research Question and Response Variable
+
+In this project, we investigate the correlation between attitudes on the justifiability of abortion with demographic factors (such as age, gender, and education level) and personal attitudes towards other issues (such as political ideology, importance of religion, and respect for authority) in a representative sample of American citizens from 1982-2011.
+
+The response variable, `Justifiability of abortion`, is a numerical measure on a scale of 1 to 10 on the individual person’s attitude toward whether abortion is justifiable or not.
+Individuals responded 1 for “abortion is never justified” and 10 for “abortion is always justified.
+
+## Exploratory Data Analysis
 
@@
 
+In the next section, we create and apply our recipes to fit MLR models.
+
 ## Data Split

### 2022-06-19 Final/FinalReport.qmd
updated structure according to YR's outline. left code untouched

---
title: "Draft"
subtitle: "STA 210 - Project"
author: "Bayes' Harem - Christina Wang, Kat Cottrell, David Goh, Ethan Song"
format: pdf
editor: source
editor_options: 
  markdown: 
    wrap: 72
---

```{r}
#| label: load-pkgs
#| message: false
#| warning: false
library(tidyverse)
library(tidymodels)
library(knitr)
library(ggfortify)
library(GGally)
```

```{r load-data}
#| message: false
abortion_data_full<-read_csv(here::here("data/abortion-attitudes", 
                                        "wvs-usa-abortion-attitudes-data.csv"))
```

## Introduction and Research Question

Understanding public attitudes on divisive political issues is an important way for political leaders to mobilize voters and for lawmakers to draft laws that represent their constituents. While it is simple to poll constituents' positions on an issue, it can be challenging to assess the complex factors that influence and predict those stances.

Abortion is one such divisive issue in the United States, with strong organization of pro-choice and pro-life groups in states across the country in support of legislation for their respective sides (Ziegler, 2020). However, following the Supreme Court's 1973 decision in Roe v. Wade, a ruling that has historically protected women's rights to abortion without excessive government restriction in America, the issue has risen in political salience. Both the pro-choice and pro-life movements have gained national prominence, and the two major political parties have polarized around the issue, with the Democratic Party in favor of and the Republican Party against policies legalization and increased access to abortion (Weinberger 2022). Abortion has also increasingly become a key issue that voters consider, with an increasing share of Americans identifying as "single-issue voters" regarding abortion (Brenan 2020).

In May 2022, a leaked draft opinion revealed that the US Supreme Court is prepared to overturn Roe v. Wade. Overturning Roe would dramatically change the trajectory of abortion politics in the US. Unless the US Congress passed a national policy, states would be able to decide whether or not to legalize abortion and gain much greater leverage in regulating access to the procedure (Weinberger 2022).

Given the potential overturning of Roe and the polarizing nature of the issue, it is important to understand how the American public feels about whether abortion should be legal or not, how accessible the procedure should be, and what factors influence these opinions. Understanding public opinion on the issue will ensure that political leaders are able to mobilize the correct constituencies, and that policy experts are able to pass policies on this issue that accurately reflect the preferences of the American people.

This dataset observes attitudes on the justifiability of abortion among respondents in in the United States across six “waves” of the World Values Survey (1982-2011), which is administered every few years and collects information about people’s values and beliefs worldwide, alongside basic demographic characteristics such as age and college education. The data are collected via face-to-face interviews at the respondents’ homes. In this data set, the response variable, justifiability of abortion, is a numerical measure on a scale of 1 to 10 on the individual person’s attitude toward whether abortion is justifiable or not.  Individuals responded 1 for “abortion is never justified” and 10 for “abortion is always justified.

Our research question is as follows: “What is the relationship between attitudes on the justifiability of abortion and personal attitudes towards other issues, such as political ideology, importance of religion, respect for authority and national pride?” We will attempt to answer this question using an EDA-informed predictive model. Given that this issue has become highly polarized by political party, we predict that liberal political attitudes will correlate with belief that abortion is more justified.

## Exploratory Data Analysis

```{r visualization-summary}
#| warning: false
ggplot(abortion_data_full, aes(x = aj)) +
  geom_histogram(binwidth = 1, aes(y=..density..)) + 
  geom_density(color = "red") +
  labs(title = "Histogram of Abortion Attitudes",
       x = "Abortion Justifiability (1-Least, 10-Most)")

summary(abortion_data_full$aj)
```

From the above visualization of the distribution of attitudes toward the justifiability of abortion, we observe that it is not a bell shape, but rather trimodal. This is likely because the question's phrasing is similar to a yes/no question, but respondents were asked to give their level of agreement on a scale of 1-10. This may worsen the ability of a multiple linear regression (MLR) model to fit the data. Consequently, further research may choose to truncate the attitude on the justifiability of abortion into a categorical variable such as (Agree, Disagree, Undecided), conducting a binomial or multinomial logistic regression thereafter.

We now extend our exploratory data analysis (EDA) to the predictor variables of interest: year of survey, ideology, Child Autonomy Index, Importance of God, Respect for Authority, and National Pride. The EDA for each variable comprises a histogram and a boxplot of the response variable, grouped by value of the predictor. We also provide a correlation matrix to detect any multicollinearity between our predictor variables, which would increase the uncertainty of our model’s parameters.

```{r eda-year}
#| warning: false
#Year of survey
ggplot(abortion_data_full, aes(x = year)) +
  geom_histogram() + 
  labs(title = "Histogram of year")
ggplot(abortion_data_full, aes(group = year, y = aj)) +
  geom_boxplot() 
```

```{r eda-ideology}
#| warning: false
# ideology
ggplot(abortion_data_full, aes(x = ideology)) +
  geom_histogram() + 
  labs(title = "Histogram of ideology")
ggplot(abortion_data_full, aes(x = ideology,group = ideology, y = aj)) +
  geom_boxplot()
```

```{r eda-cai}
#| warning: false
# Child autonomy index
ggplot(abortion_data_full, aes(x = cai)) +
  geom_histogram() + 
  labs(title = "Histogram of Child autonomy index")
ggplot(abortion_data_full, aes(x = cai, group = cai, y = aj)) +
  geom_boxplot()
```

```{r eda-god}
#| warning: false
# Importance of God
ggplot(abortion_data_full, aes(x = godimportant)) +
  geom_histogram() + 
  labs(title = "Histogram of how respondent saw God's importance")
ggplot(abortion_data_full, aes(x = godimportant, group = godimportant, 
                               y = aj)) +
  geom_boxplot()
```

```{r eda-respect}
#| warning: false
# Respect for authority
ggplot(abortion_data_full, aes(x = respectauthority)) +
  geom_histogram() + 
  labs(title = "Histogram of respect for authority")
ggplot(abortion_data_full, aes(x = respectauthority, group = respectauthority, 
                          y = aj)) +
  geom_boxplot()
```

```{r eda-pride}
#| warning: false
# National Pride
ggplot(abortion_data_full, aes(x = nationalpride)) +
  geom_histogram() + 
  labs(title = "Histogram of National pride")
ggplot(abortion_data_full, aes(x = nationalpride, group = nationalpride, y = aj)) +
  geom_boxplot()
```

```{r correlations}
#| warning: false
ggpairs(abortion_data_full, 
  columns = c("year", "ideology", "cai", "godimportant", "respectauthority", 
              "nationalpride"),
  columnLabels = c("year", "ideology", "child autonomy", "imptnce of God", 
                   "resp for autho", "natl pride")) +
   theme(
    axis.text.y = element_text(size = 10),
    axis.text.x = element_text(angle = 45, size = 10),
    strip.text.y = element_text(angle = 0, hjust = 0)
    )
```

\pagebreak

## Methodology

Given that our response variable, attitude towards the justifiability of abortion, is measured on a numeric scale from 1 to 10, and that there are multiple predictor variables that we suspect to be potentially related to it, we will conduct multiple linear regression (MLR) to model the effect of these predictors on the variation in abortion attitude.

We chose to consider predictor variables that, from our exploratory data analysis (EDA), looked like they were potentially correlated with the outcome. Because the predictor variables were discrete, our EDA was presented as a series of box plots. We observed the differences in median and quartile values of abortion attitude across different predictor values, ruling out variables with no visible hint of relationship. This led to the following predictor variables: Year of survey, Political Ideology, Child Autonomy Index, Importance of God, Respect for Authority, and National Pride.

A brief note on data cleaning: many survey questions were only added from 1995 onward, and observations before 1995 have a high number of missing values; therefore, we removed the observations that were part of “generational waves” before 1995.

```{r filter-data}
abortion_data <- abortion_data_full %>% 
  filter(year >= "1995") %>%
  select(-starts_with("wave"), -starts_with("wvsccode"))
```

## Fitting the Model

We conducted a 75%-25% data split into training and testing sets, using the seed “206”. Fitting the MLR, we obtain the following table of parameter point estimates.

```{r splits}
set.seed(206)
abortion_split <- initial_split(abortion_data)
abortion_training <- training(abortion_split)
abortion_testing <- testing(abortion_split)
```

```{r spec}
abortion_spec <- linear_reg() %>%
  set_engine("lm")
```

```{r recipe1}
abortion_rec1 <- recipe(aj ~ year + ideology + cai + godimportant +
                        respectauthority, nationalpride,
                        data = abortion_data) %>%
  step_center(year) %>%
  step_dummy(all_nominal_predictors()) %>%
  step_zv(all_predictors())

abortion_rec1
```

```{r recipe2}
abortion_rec2 <- recipe(aj ~ year + godimportant,
                        data = abortion_data) %>%
  step_center(year) %>%
  step_dummy(all_nominal_predictors()) %>%
  step_zv(all_predictors())

abortion_rec2
```

```{r recipe3}
abortion_rec3 <- recipe(aj ~ year + ideology + cai + nationalpride,
                        data = abortion_data) %>%
  step_center(year) %>%
  step_dummy(all_nominal_predictors()) %>%
  step_zv(all_predictors())

abortion_rec3
```

```{r workflows}
abortion_wflow1 <- workflow() %>%
  add_model(abortion_spec) %>%
  add_recipe(abortion_rec1)

abortion_wflow2 <- workflow() %>%
  add_model(abortion_spec) %>%
  add_recipe(abortion_rec2)

abortion_wflow3 <- workflow() %>%
  add_model(abortion_spec) %>%
  add_recipe(abortion_rec3)
```

```{r fit-models}
abortion_fit1 <- abortion_wflow1 %>%
  fit(data = abortion_training)
tidy(abortion_fit1) %>%
  kable(digits = 3)

abortion_fit2 <- abortion_wflow2 %>%
  fit(data = abortion_training)
tidy(abortion_fit2) %>%
  kable(digits = 3)

abortion_fit3 <- abortion_wflow3 %>%
  fit(data = abortion_training)
tidy(abortion_fit3) %>%
  kable(digits = 3)
```

[table]

Hence, our MLR model equation is as follows, where x1 is __, x2 is __, x3 is __, x4 is __, and x5 is __:

[equation in latex]

## Model Inference and Prediction

Inference: CI and HT

We will take a quick look at the 95% confidence intervals of the parameters from our fitted model, displayed in the following plots:

[plots of parameters’ .estimate against the parameter names]

```{r ci}
tidy(abortion_fit1, conf.int = TRUE) %>%
  kable(digits = 3)
tidy(abortion_fit2, conf.int = TRUE) %>%
  kable(digits = 3)
tidy(abortion_fit3, conf.int = TRUE) %>%
  kable(digits = 3)
```

We then proceed to our hypothesis test. As stated in our introduction, given that this issue has become highly polarized by political party, we suspect that liberal political attitudes will correlate with belief that abortion is more justified, leading to the following hypotheses.

[write out null hypothesis and alternative hypothesis]

From the p-value in our output displayed in Figure 5(?), we conclude …

Prediction:

We now consider the strength of our model in predicting the attitude toward the justifiability of abortion of a person of certain characteristics.

For a person who [describe liberal characteristics here], our model predicts that their attitude toward the justifiability of abortion is [write what the model would predict].

This does/does not support our hypothesis based on contextual knowledge of this issue.

However, calculating the R-squared and root-mean-squared-error (RMSE) values of our model shows that our model is not very strong at predicting attitude toward the justifiability of abortion, with a R^2 of __ and RMSE of __:

[R^2, RMSE output]

```{r predict}
abortion_training_pred1 <- predict(abortion_fit1, abortion_training) %>%
  bind_cols(abortion_training %>% select(aj, year, ideology, cai, 
                                         godimportant, respectauthority, 
                                         nationalpride))
abortion_training_pred1

abortion_training_pred2 <- predict(abortion_fit2, abortion_training) %>%
  bind_cols(abortion_training %>% select(aj, year, godimportant))
abortion_training_pred2

abortion_training_pred3 <- predict(abortion_fit3, abortion_training) %>%
  bind_cols(abortion_training %>% select(aj, year, ideology, cai, 
                                         nationalpride))
abortion_training_pred3
```

We will discuss potential reasons for this lackluster performance in our model diagnosis and conclusion sections; however, we will first conduct model evaluation and comparison to ascertain if our chosen model is still a better fit to the data compared to some alternatives.

A quick glance at our data might seem that the models are not that strong at predicting the abortion attitude (aj) of a given observation. We will look into this further now with cross-validation, then a test in AIC and BIC statistics.

## Model Evaluation

To see how well our model fits the data on the attitudes toward the justifiability of abortion, we conducted a 10-fold cross-validation, calculating various metrics – the Akaike information criterion (AIC), Bayesian Information Criterion (BIC) and adjusted R-squared value – to evaluate the performance of our model on 10 subsets of our data.

We also compared this to two alternative models that used other combinations of predictors from the data set.

[3 line plots, displaying AIC, BIC and R^2 value respectively, against Fold number. Each plot should have the line graph for each of the 3 models we used)

```{r cv}
set.seed(206)
folds <- vfold_cv(abortion_training, v = 10)
abortion_fit_rs1 <- abortion_wflow1 %>%
  fit_resamples(folds)

abortion_fit_rs2 <- abortion_wflow2 %>%
  fit_resamples(folds)

abortion_fit_rs3 <- abortion_wflow3 %>%
  fit_resamples(folds)
```

```{r cv-table}
cv_metrics1 <- collect_metrics(abortion_fit_rs1, summarize = FALSE) 
cv_metrics2 <- collect_metrics(abortion_fit_rs2, summarize = FALSE) 
cv_metrics3 <- collect_metrics(abortion_fit_rs3, summarize = FALSE)

cv_metrics1 %>%
  mutate(.estimate = round(.estimate, 3)) %>%
  pivot_wider(id_cols = id, names_from = .metric, values_from = .estimate) %>%
  kable(col.names = c("Fold", "RMSE", "R-squared"))

cv_metrics2 %>%
  mutate(.estimate = round(.estimate, 3)) %>%
  pivot_wider(id_cols = id, names_from = .metric, values_from = .estimate) %>%
  kable(col.names = c("Fold", "RMSE", "R-squared"))

cv_metrics3 %>%
  mutate(.estimate = round(.estimate, 3)) %>%
  pivot_wider(id_cols = id, names_from = .metric, values_from = .estimate) %>%
  kable(col.names = c("Fold", "RMSE", "R-squared"))
```

```{r aic-bic}
glance(abortion_fit1) %>% 
  select(AIC, BIC)

glance(abortion_fit2) %>% 
  select(AIC, BIC)

glance(abortion_fit3) %>% 
  select(AIC, BIC)
```

From the above AIC and BIC values, we have evidence that our chosen model is the strongest model and best fit for the data compared to two alternatives with fewer predictors. Our model has the lowest AIC and BIC values, which are usually penalized for having more predictors, but in this case the model is strong enough to overcome that penalty.

## Model Diagnosis

Conditions check

Linearity [residual plot vs fitted values]

Randomness

Independence [residual plot vs id]

## Conclusion + Discussion

[paste here after finalizing the google doc]

## Data dictionary

The data dictionary can be found [here](ADD%20LINK!!!).

## References

Brenan, M. (2020, July 7). One in Four Americans Consider Abortion a Key
Voting Issue. Gallup.
https://news.gallup.com/poll/313316/one-four-americans-consider-abortion-key-voting-issue.aspx

Weinberger, J. (2022, May 6). How we got here: Roe v. Wade from 1973 to
today. Vox.
https://www.vox.com/23055389/roe-v-wade-timeline-abortion-overturn-political-polarization

Ziegler, M. (2020, October 22). Abortion politics polarized before Roe.
When it's gone, the fighting won't stop. The Washington Post.
https://www.washingtonpost.com/outlook/2020/10/22/roe-polarize-abortion-politics/

### 2022-06-19 Final/FinalReport.qmd
edited introduction and research question wording

@@
 Understanding public attitudes on divisive political issues is an important way for political leaders to mobilize voters and for lawmakers to draft laws that represent their constituents.
-While it is simple to poll constituents' positions on an issue, it can be challenging to assess the complex factors that influence and predict those stances.
+While it may be easier to poll constituents' positions on an issue, it can be challenging to assess the complex factors that influence and predict those stances.
 
-Abortion is one such divisive issue in the United States, with strong organization of pro-choice and pro-life groups in states across the country in support of legislation for their respective sides (Ziegler, 2020).
+Abortion is one such divisive issue in the United States.
+Both pro-choice and pro-life groups have a history of mobilizing in states across the country in support of legislation for their respective sides (Ziegler, 2020).
 However, following the Supreme Court's 1973 decision in Roe v.
-Wade, a ruling that has historically protected women's rights to abortion without excessive government restriction in America, the issue has risen in political salience.
-Both the pro-choice and pro-life movements have gained national prominence, and the two major political parties have polarized around the issue, with the Democratic Party in favor of and the Republican Party against policies legalization and increased access to abortion (Weinberger 2022).
-Abortion has also increasingly become a key issue that voters consider, with an increasing share of Americans identifying as "single-issue voters" regarding abortion (Brenan 2020).
+Wade, a ruling which protected an individual’s right to have an abortion before fetal viability, the issue has risen in political salience.
+Both the pro-choice and pro-life movements have gained national prominence, and the two major political parties have polarized around the issue, with the Democratic Party in favor of and the Republican Party against policies legalizing and increasing access to abortion (Weinberger 2022).
+Abortion has also increasingly become a key issue that voters consider when making their choice at the ballot box, with an increasing share of Americans identifying as "single-issue voters" regarding abortion (Brenan 2020).
 
@@
 Overturning Roe would dramatically change the trajectory of abortion politics in the US.
-Unless the US Congress passed a national policy, states would be able to decide whether or not to legalize abortion and gain much greater leverage in regulating access to the procedure (Weinberger 2022).
+Unless Congress passes a national policy, states would be able to decide whether or not to legalize abortion and gain much greater leverage in regulating access to the procedure (Weinberger 2022).
 
-Given the potential overturning of Roe and the polarizing nature of the issue, it is important to understand how the American public feels about whether abortion should be legal or not, how accessible the procedure should be, and what factors influence these opinions.
+Given the potential overturning of Roe and the polarizing nature of the issue, it is important to understand how the American public feels about whether abortion should be legal or not, how accessible the procedure should be, and which factors influence these opinions.
 Understanding public opinion on the issue will ensure that political leaders are able to mobilize the correct constituencies, and that policy experts are able to pass policies on this issue that accurately reflect the preferences of the American people.
 
-This dataset observes attitudes on the justifiability of abortion among respondents in in the United States across six “waves” of the World Values Survey (1982-2011), which is administered every few years and collects information about people’s values and beliefs worldwide, alongside basic demographic characteristics such as age and college education.
+The dataset used in this analysis observes attitudes on the justifiability of abortion among respondents in in the United States across six “waves” of the World Values Survey (1982-2011), which is administered every few years and collects information about people’s values and beliefs worldwide, alongside basic demographic characteristics.
 The data are collected via face-to-face interviews at the respondents’ homes.
@@
 
-Our research question is as follows: “What is the relationship between attitudes on the justifiability of abortion and personal attitudes towards other issues, such as political ideology, importance of religion, respect for authority and national pride?” We will attempt to answer this question using an EDA-informed predictive model.
-Given that this issue has become highly polarized by political party, we predict that liberal political attitudes will correlate with belief that abortion is more justified.
+Our research question is as follows: “Do an individual’s political ideology and their personal attitudes/preferences towards other issues, such as the importance of religion in their life and their respect for authority, among others, predict their attitude on the justifiability of abortion?” We will attempt to answer this question using an EDA-informed predictive model.
+Given that this issue has become highly polarized by political party, we predict that liberal ideology and liberal-leaning attitudes on other issues will correlate with belief that abortion is more justified.
 

### 2022-06-20 Final/FinalReport.qmd
adjustments to variable types in dataset, small edits to recipes

@@
   select(-starts_with("wave"), -starts_with("wvsccode"))
+
+abortion_data$year <- as.factor(abortion_data$year)
+abortion_data$respectauthority <- as.factor(abortion_data$respectauthority)
+abortion_data$respectauthority <- factor(abortion_data$respectauthority, levels=c('0', '-1', '1'))
+abortion_data$nationalpride <- as.factor(abortion_data$nationalpride)
 ```
@@
 abortion_rec1 <- recipe(aj ~ year + ideology + cai + godimportant +
-                        respectauthority, nationalpride,
+                        respectauthority + nationalpride,
                         data = abortion_data) %>%
-  step_center(year) %>%
+  step_center(ideology) %>%
   step_dummy(all_nominal_predictors()) %>%
@@
                         data = abortion_data) %>%
-  step_center(year) %>%
   step_dummy(all_nominal_predictors()) %>%
@@
                         data = abortion_data) %>%
-  step_center(year) %>%
+  step_center(ideology) %>%
   step_dummy(all_nominal_predictors()) %>%
@@
 
-```{r ci}
+```{r inference}
 tidy(abortion_fit1, conf.int = TRUE) %>%
-  kable(digits = 3)
-tidy(abortion_fit2, conf.int = TRUE) %>%
-  kable(digits = 3)
-tidy(abortion_fit3, conf.int = TRUE) %>%
   kable(digits = 3)
@@
 We then proceed to our hypothesis test.
-As stated in our introduction, given that this issue has become highly polarized by political party, we suspect that liberal political attitudes will correlate with belief that abortion is more justified, leading to the following hypotheses.
+As stated in our introduction, given that this issue has become highly polarized by political party, we suspect that liberal ideology and political attitudes will correlate with belief that abortion is more justified, leading to the following hypotheses:
 
@@
 abortion_training_pred1
-
-abortion_training_pred2 <- predict(abortion_fit2, abortion_training) %>%
-  bind_cols(abortion_training %>% select(aj, year, godimportant))
-abortion_training_pred2
-
-abortion_training_pred3 <- predict(abortion_fit3, abortion_training) %>%
-  bind_cols(abortion_training %>% select(aj, year, ideology, cai, 
-                                         nationalpride))
-abortion_training_pred3
 ```

### 2022-06-20 Final/FinalReport.qmd
Added Methodology and Fitting the Model

@@
 
-Given that our response variable, attitude towards the justifiability of abortion, is measured on a numeric scale from 1 to 10, and that there are multiple predictor variables that we suspect to be potentially related to it, we will conduct multiple linear regression (MLR) to model the effect of these predictors on the variation in abortion attitude.
+Given that our response variable, attitude towards the justifiability of abortion, is measured on a numeric scale from 1 to 10, and that there are multiple predictor variables being tested against the null hypothesis, we will conduct multiple linear regression (MLR) to model the effect of these predictors on the variation in abortion attitudes.
 
-We chose to consider predictor variables that, from our exploratory data analysis (EDA), looked like they were potentially correlated with the outcome.
+We chose to consider the predictor variables our exploratory data analysis (EDA) suggested were potentially correlated with the response variable.
 Because the predictor variables were discrete, our EDA was presented as a series of box plots.
-We observed the differences in median and quartile values of abortion attitude across different predictor values, ruling out variables with no visible hint of relationship.
+We observed the differences in median and quartile values of abortion attitude across different predictor values, ruling out variables with no significant influence on the outcome.
 This led to the following predictor variables: Year of survey, Political Ideology, Child Autonomy Index, Importance of God, Respect for Authority, and National Pride.
 
-A brief note on data cleaning: many survey questions were only added from 1995 onward, and observations before 1995 have a high number of missing values; therefore, we removed the observations that were part of “generational waves” before 1995.
+Year of survey, Respect for Authority, and National Pride were made factors since <CHRISTINA EXPLAIN HERE>.
+Political Ideology, Child Autonomy Index, and Importance of God had numerical sample spaces and were therefore treated as numerical variables.
+
+*A brief note on data cleaning: Many survey questions were only added from 1995 onward, so observations before 1995 have a high number of missing values.
+For these reasons, observations prior to 1995 were removed.*
 
@@
 
-We conducted a 75%-25% data split into training and testing sets, using the seed “206”. Fitting the MLR, we obtain the following table of parameter point estimates.
+We conducted a 75% - 25% data split into training and testing sets, using the random seed $206$. Fitting the MLR, we obtain the following table of parameter point estimates.
+
+```{r fit-model}
+#| echo = false
+abortion_fit <- abortion_wflow %>%
+  fit(data = abortion_training)
+tidy(abortion_fit) %>%
+  kable(digits = 3)
+```
+
+Hence, our MLR model equation is
+
+$$y ~=~ 8.5 ~+~ 0.029x_1 ~+~ 0.469x_2 ~-~ (0.284x_3 ~+~ 0.288x_4 ~+~ 0.19x_5)$$
+
+where $x_1 ~=~ \text{year}$, $x_2 ~=~ \text{child autonomy index}$, $x_3 ~=~ \text{ideology}$, $x_4 ~=~ \text{god is important}$, and $x_5 ~=~ \text{respect for authority}$.
 

### 2022-06-20 Final/FinalReport.qmd
completed hypothesis test

@@
 From the above visualization of the distribution of attitudes toward the justifiability of abortion, we observe that it is not a bell shape, but rather trimodal.
-This is likely because the question's phrasing is similar to a yes/no question, but respondents were asked to give their level of agreement on a scale of 1-10.
+This may be because the question's phrasing is similar to a yes/no question, but respondents were asked to give their level of agreement on a scale of 1-10.
 This may worsen the ability of a multiple linear regression (MLR) model to fit the data.
@@
 
-Year of survey, Respect for Authority, and National Pride were made factors since <CHRISTINA EXPLAIN HERE>.
-Political Ideology, Child Autonomy Index, and Importance of God had numerical sample spaces and were therefore treated as numerical variables.
+*A note on data cleaning: Many survey questions were only added from 1995 onward, so observations before 1995 have a high number of missing values.
+For these reasons, observations prior to 1995 were removed.
 
-*A brief note on data cleaning: Many survey questions were only added from 1995 onward, so observations before 1995 have a high number of missing values.
-For these reasons, observations prior to 1995 were removed.*
+Year of survey, Respect for Authority, and National Pride were changed into factors since they are categorical variables.
+Year has 4 levels (one corresponding to each year), Respect for Authority has 3 levels (0 being the baseline - neutral, -1 being respect for authority bad, and 1 being respect for authority good), and National Pride is a binary (0 being the baseline - having no feelings of national pride, and 1 being having feelings of national pride) Political Ideology, Child Autonomy Index, and Importance of God had numerical sample spaces and were therefore treated as numerical variables.*
 
-```{r filter-data}
+```{r clean-data}
 abortion_data <- abortion_data_full %>% 
@@
 
-We conducted a 75% - 25% data split into training and testing sets, using the random seed $206$. Fitting the MLR, we obtain the following table of parameter point estimates.
-
-```{r fit-model}
-#| echo = false
-abortion_fit <- abortion_wflow %>%
-  fit(data = abortion_training)
-tidy(abortion_fit) %>%
-  kable(digits = 3)
-```
-
-Hence, our MLR model equation is
-
-$$y ~=~ 8.5 ~+~ 0.029x_1 ~+~ 0.469x_2 ~-~ (0.284x_3 ~+~ 0.288x_4 ~+~ 0.19x_5)$$
-
-where $x_1 ~=~ \text{year}$, $x_2 ~=~ \text{child autonomy index}$, $x_3 ~=~ \text{ideology}$, $x_4 ~=~ \text{god is important}$, and $x_5 ~=~ \text{respect for authority}$.
+We conducted a 75% - 25% data split into training and testing sets, using the random seed $206$.
 
@@
 ```{r recipe1}
-abortion_rec1 <- recipe(aj ~ year + ideology + cai + godimportant +
+abortion_rec <- recipe(aj ~ year + ideology + cai + godimportant +
                         respectauthority + nationalpride,
@@
 
-abortion_rec1
+abortion_rec
 ```
@@
 ```{r workflows}
-abortion_wflow1 <- workflow() %>%
+abortion_wflow <- workflow() %>%
   add_model(abortion_spec) %>%
-  add_recipe(abortion_rec1)
+  add_recipe(abortion_rec)
 
@@
 ```{r fit-models}
-abortion_fit1 <- abortion_wflow1 %>%
+abortion_fit <- abortion_wflow %>%
   fit(data = abortion_training)
-tidy(abortion_fit1) %>%
+tidy(abortion_fit) %>%
   kable(digits = 3)
@@
 
-[table]
+Fitting the MLR, we obtain the following table of parameter point estimates.
 
-Hence, our MLR model equation is as follows, where x1 is __, x2 is __, x3 is __, x4 is __, and x5 is __:
+```{r fit-model}
+#| echo: false
+abortion_fit <- abortion_wflow %>%
+  fit(data = abortion_training)
+tidy(abortion_fit) %>%
+  kable(digits = 3)
+```
 
-[equation in latex]
+Hence, our MLR model equation is
+
+$$y ~=~ 8.5 ~+~ 0.029x_1 ~+~ 0.469x_2 ~-~ (0.284x_3 ~+~ 0.288x_4 ~+~ 0.19x_5)$$
+
+where $x_1 ~=~ \text{year}$, $x_2 ~=~ \text{child autonomy index}$, $x_3 ~=~ \text{ideology}$, $x_4 ~=~ \text{god is important}$, and $x_5 ~=~ \text{respect for authority}$.
 
@@
 ```{r inference}
-tidy(abortion_fit1, conf.int = TRUE) %>%
+tidy(abortion_fit, conf.int = TRUE) %>%
   kable(digits = 3)
@@
 We then proceed to our hypothesis test.
-As stated in our introduction, given that this issue has become highly polarized by political party, we suspect that liberal ideology and political attitudes will correlate with belief that abortion is more justified, leading to the following hypotheses:
+As stated in our introduction, given that this issue has become highly polarized by political party, we suspect that liberal political ideology and political attitudes will correlate with belief that abortion is more justified.
 
-[write out null hypothesis and alternative hypothesis]
+We define the following as liberal political ideology and attitudes:
+- ideology: values closer to 1 (most left-leaning)
+- cai: values closer to 2 (belief in child self-determination and independence)
+- godimportant: values closer to 1 (belief that God is not important)
+- respectauthority: -1 (belief that greater respect for authority is a bad thing)
+- nationalpride: 0 (not proud to be American)
 
-From the p-value in our output displayed in Figure 5(?), we conclude …
+This leads to the following hypotheses:
+
+H0 (null hypothesis): An individual’s political ideology (ideology), their beliefs about child autonomy (cai), the importance of God in their life (godimportant), their view on whether respect for authority is good (respectauthority), and their pride in being American (nationalpride) are not significant predictors of aj, and the variables’ respective slope point estimates will be 0.
+
+HA (alternative hypothesis: An individual’s political ideology (ideology), their beliefs about child autonomy (cai), the importance of God in their life (godimportant), their view on whether respect for authority is good (respectauthority), and their pride in being American (nationalpride) are significant predictors of aj, and the variables’ respective slope point estimates will not be 0.
+
+Looking at the p-values in our output displayed above, we can reject the null hypothesis for the predictors ideology, cai, and godimportant, as the p-values are very small, way below the 0.05 significance threshold.
+Hence, we can conclude that the data provide convincing evidence that an individual’s political ideology, beliefs about child autonomy, and the importance of God in their life are significant predictors of their abortion attitudes.
+However, we fail to reject the null hypothesis for the predictors respectauthority and nationalpride, as the p-values for each level of these predictors are above the 0.05 significance threshold.
+Hence, we cannot conclude that the data provide convincing evidence that an individual’s view on respect for authority and their pride in being American are significant predictors of their abortion attitudes.
 
@@
 ```{r predict}
-abortion_training_pred1 <- predict(abortion_fit1, abortion_training) %>%
+abortion_training_pred <- predict(abortion_fit, abortion_training) %>%
   bind_cols(abortion_training %>% select(aj, year, ideology, cai, 
@@
                                          nationalpride))
-abortion_training_pred1
+abortion_training_pred
 ```
@@
 folds <- vfold_cv(abortion_training, v = 10)
-abortion_fit_rs1 <- abortion_wflow1 %>%
+abortion_fit_rs <- abortion_wflow %>%
   fit_resamples(folds)
@@
 ```{r cv-table}
-cv_metrics1 <- collect_metrics(abortion_fit_rs1, summarize = FALSE) 
+cv_metrics <- collect_metrics(abortion_fit_rs, summarize = FALSE) 
 cv_metrics2 <- collect_metrics(abortion_fit_rs2, summarize = FALSE) 
@@
 
-cv_metrics1 %>%
+cv_metrics %>%
   mutate(.estimate = round(.estimate, 3)) %>%
@@
 ```{r aic-bic}
-glance(abortion_fit1) %>% 
+glance(abortion_fit) %>% 
   select(AIC, BIC)

### 2022-06-20 Final/FinalReport.qmd
Update FinalReport.qmd
david's changes. issues: natl pride in CI and predict; CI dealing with the factor variables; remove recipes and workflows for models 2,3

@@
 #| warning: false
+#| echo: false
 library(tidyverse)
@@
 library(GGally)
+library(cowplot)
+library(readr)
 ```
@@
 #| message: false
+#| echo: false
 abortion_data_full<-read_csv(here::here("data/abortion-attitudes", 
                                         "wvs-usa-abortion-attitudes-data.csv"))
+```
+
+```{r clean-data}
+#| echo: false
+abortion_data <- abortion_data_full %>% 
+  filter(year >= "1995") %>%
+  select(-starts_with("wave"), -starts_with("wvsccode"))
 ```
@@
 #| warning: false
-ggplot(abortion_data_full, aes(x = aj)) +
-  geom_histogram(binwidth = 1, aes(y=..density..)) + 
+#| echo: false
+ggplot(abortion_data, aes(x = aj)) +
+  geom_histogram(binwidth = 1, aes(y=..density..), fill = "light blue") + 
   geom_density(color = "red") +
@@
        x = "Abortion Justifiability (1-Least, 10-Most)")
-
-summary(abortion_data_full$aj)
 ```
@@
 
-```{r eda-year}
+```{r eda-histograms}
 #| warning: false
-#Year of survey
-ggplot(abortion_data_full, aes(x = year)) +
-  geom_histogram() + 
-  labs(title = "Histogram of year")
-ggplot(abortion_data_full, aes(group = year, y = aj)) +
-  geom_boxplot() 
+#| message: false
+#| echo: false
+h1 <- ggplot(abortion_data, aes(x = year)) +
+  geom_histogram(fill = "orange") + 
+  labs(x = "Year of Survey")
+
+h2 <- ggplot(abortion_data, aes(x = ideology)) +
+  geom_histogram(fill = "orange") + 
+  labs(x = "Political Ideology")
+
+h3 <- ggplot(abortion_data, aes(x = cai)) +
+  geom_histogram(fill = "orange") + 
+  labs(x = "Child Autonomy Index")
+
+h4 <- ggplot(abortion_data, aes(x = godimportant)) +
+  geom_histogram(fill = "orange") + 
+  labs(x = "Importance of God")
+
+h5 <- ggplot(abortion_data, aes(x = respectauthority)) +
+  geom_histogram(fill = "orange") + 
+  labs(x = "Respect for Authority")
+
+plot_grid(h1, h2, h3, h4, h5, ncol = 3, nrow = 2)
+
 ```
 
-```{r eda-ideology}
+```{r eda-boxplots}
 #| warning: false
-# ideology
-ggplot(abortion_data_full, aes(x = ideology)) +
-  geom_histogram() + 
-  labs(title = "Histogram of ideology")
-ggplot(abortion_data_full, aes(x = ideology,group = ideology, y = aj)) +
-  geom_boxplot()
-```
+#| echo: false
+b1 <- ggplot(abortion_data, aes(x = year, group = year, y = aj)) +
+  geom_boxplot() +
+  labs(y = "Abortion Attitude", x = "Year of Survey ")
 
-```{r eda-cai}
-#| warning: false
-# Child autonomy index
-ggplot(abortion_data_full, aes(x = cai)) +
-  geom_histogram() + 
-  labs(title = "Histogram of Child autonomy index")
-ggplot(abortion_data_full, aes(x = cai, group = cai, y = aj)) +
-  geom_boxplot()
-```
+b2 <- ggplot(abortion_data, aes(x = ideology,group = ideology, y = aj)) +
+  geom_boxplot()+
+  labs(y = "Abortion Attitude", x = "Political Ideology")
 
-```{r eda-god}
-#| warning: false
-# Importance of God
-ggplot(abortion_data_full, aes(x = godimportant)) +
-  geom_histogram() + 
-  labs(title = "Histogram of how respondent saw God's importance")
-ggplot(abortion_data_full, aes(x = godimportant, group = godimportant, 
+b3 <- ggplot(abortion_data, aes(x = cai, group = cai, y = aj)) +
+  geom_boxplot()+
+  labs(y = "Abortion Attitude", x = "Child Autonomy Index")
+
+b4 <- ggplot(abortion_data, 
+             aes(x = godimportant, group = godimportant, y = aj)) +
+  geom_boxplot()+
+  labs(y = "Abortion Attitude", x = "Importance of God")
+
+b5 <- ggplot(abortion_data, aes(x = respectauthority, group = respectauthority, 
                           y = aj)) +
-  geom_boxplot()
-```
+  geom_boxplot()+
+  labs(y = "Abortion Attitude", x = "Respect For Authority")
 
-```{r eda-respect}
-#| warning: false
-# Respect for authority
-ggplot(abortion_data_full, aes(x = respectauthority)) +
-  geom_histogram() + 
-  labs(title = "Histogram of respect for authority")
-ggplot(abortion_data_full, aes(x = respectauthority, group = respectauthority, 
-                          y = aj)) +
-  geom_boxplot()
-```
-
-```{r eda-pride}
-#| warning: false
-# National Pride
-ggplot(abortion_data_full, aes(x = nationalpride)) +
-  geom_histogram() + 
-  labs(title = "Histogram of National pride")
-ggplot(abortion_data_full, aes(x = nationalpride, group = nationalpride, y = aj)) +
-  geom_boxplot()
+plot_grid(b1, b2, b3, b4, b5, ncol = 3, nrow = 2)
 ```
@@
 #| warning: false
-ggpairs(abortion_data_full, 
-  columns = c("year", "ideology", "cai", "godimportant", "respectauthority", 
-              "nationalpride"),
-  columnLabels = c("year", "ideology", "child autonomy", "imptnce of God", 
-                   "resp for autho", "natl pride")) +
+#| echo: false
+ggpairs(abortion_data, 
+  columns = c("year", "ideology", "cai", "godimportant", "respectauthority"),
+  columnLabels = c("YS", "PI", "CAI", "IG", 
+                   "RA")) +
    theme(
-    axis.text.y = element_text(size = 10),
-    axis.text.x = element_text(angle = 45, size = 10),
+    axis.text.y = element_text(size = 8),
+    axis.text.x = element_text(angle = 45, size = 9),
     strip.text.y = element_text(angle = 0, hjust = 0)
@@
 
-Year of survey, Respect for Authority, and National Pride were changed into factors since they are categorical variables.
-Year has 4 levels (one corresponding to each year), Respect for Authority has 3 levels (0 being the baseline - neutral, -1 being respect for authority bad, and 1 being respect for authority good), and National Pride is a binary (0 being the baseline - having no feelings of national pride, and 1 being having feelings of national pride) Political Ideology, Child Autonomy Index, and Importance of God had numerical sample spaces and were therefore treated as numerical variables.*
-
-```{r clean-data}
-abortion_data <- abortion_data_full %>% 
-  filter(year >= "1995") %>%
-  select(-starts_with("wave"), -starts_with("wvsccode"))
-
+```{r clean-data2}
+#| echo: false
 abortion_data$year <- as.factor(abortion_data$year)
@@
 
+Year of survey, Respect for Authority, and National Pride were changed into factors since they are categorical variables.
+Year has 4 levels (one corresponding to each year), Respect for Authority has 3 levels (0 being the baseline - neutral, -1 being respect for authority bad, and 1 being respect for authority good), and National Pride is a binary (0 being the baseline - having no feelings of national pride, and 1 being having feelings of national pride) Political Ideology, Child Autonomy Index, and Importance of God had numerical sample spaces and were therefore treated as numerical variables.*
+
 ## Fitting the Model
@@
 ```{r splits}
+#| echo: false
 set.seed(206)
@@
 ```{r spec}
+#| echo: false
 abortion_spec <- linear_reg() %>%
@@
 ```{r recipe1}
+#| echo: false
 abortion_rec <- recipe(aj ~ year + ideology + cai + godimportant +
@@
 ```{r recipe2}
+#| echo: false
 abortion_rec2 <- recipe(aj ~ year + godimportant,
@@
 ```{r recipe3}
+#| echo: false
 abortion_rec3 <- recipe(aj ~ year + ideology + cai + nationalpride,
@@
 ```{r workflows}
+#| echo: false
 abortion_wflow <- workflow() %>%
@@
 ```{r fit-models}
+#| echo: false
 abortion_fit <- abortion_wflow %>%
@@
 ```{r fit-model}
-#| echo: false
+#| echo = false
 abortion_fit <- abortion_wflow %>%
@@
 
-$$y ~=~ 8.5 ~+~ 0.029x_1 ~+~ 0.469x_2 ~-~ (0.284x_3 ~+~ 0.288x_4 ~+~ 0.19x_5)$$
+$$\begin{align}y ~=~ &6.596 ~+~ 0.468x_1 ~+~ 0.33x_2 ~+~ 0.339x_3 ~+~ 0.521x_4 ~+~ 0.14x_5 \\&~-~ (0.28x_6 ~+~ 0.287x_7 ~+~ 0.179x_8 ~+~ 0.11x_9)\end{align}$$
 
-where $x_1 ~=~ \text{year}$, $x_2 ~=~ \text{child autonomy index}$, $x_3 ~=~ \text{ideology}$, $x_4 ~=~ \text{god is important}$, and $x_5 ~=~ \text{respect for authority}$.
+where $x_1 ~=~ \text{child autonomy index}$, $x_2 ~=~ \text{year 1999}$, $x_3 ~=~ \text{year 2006}$, $x_4 ~=~ \text{year 2011}$, $x_5 ~=~ \text{respect authority -1}$, $x_6 ~=~ \text{ideology}$, $x_7 ~=~ \text{God important}$, $x_8 ~=~ \text{respect authority 1}$, and $x_9 ~=~ \text{national pride}$.
 
@@
 
-Inference: CI and HT
+We will take a quick look at the 95% **confidence intervals** of the parameters from our fitted model, displayed in the following plots:
 
-We will take a quick look at the 95% confidence intervals of the parameters from our fitted model, displayed in the following plots:
+```{r ci}
+#| echo: false
+df <- tidy(abortion_fit, conf.int = TRUE)
+intercept <- df[1,]
+yr <- df[2,]
+cai <- df[4,]
+df <- df[-c(1,2,4),]
 
-[plots of parameters’ .estimate against the parameter names]
+p1 <- ggplot(intercept, aes(x = term, y = estimate, ymin = conf.low, 
+                            ymax = conf.high)) +
+  geom_pointrange() +
+  labs(x = NULL)
+p2 <- ggplot(yr, aes(x = term, y = estimate, ymin = conf.low, ymax = conf.high)) +
+  geom_pointrange() +
+  labs(x = NULL) +
+  scale_x_discrete(labels=c("Year of Survey"))
+p3 <- ggplot(cai, aes(x = term, y = estimate, ymin = conf.low, ymax = conf.high)) +
+  geom_pointrange()+
+  labs(x = NULL)+
+  scale_x_discrete(labels=c("Child Autonomy Index"))
 
-```{r inference}
-tidy(abortion_fit, conf.int = TRUE) %>%
-  kable(digits = 3)
+p4 <- ggplot(df, aes(x = term, y = estimate, ymin = conf.low, ymax = conf.high)) +
+  geom_pointrange()+
+  labs(x = NULL) +
+  scale_x_discrete(breaks=c("godimportant","ideology","respectauthority"),
+        labels=c("Importance of God", "Political Ideology", "Respect for Authority"))
+
+title <- ggdraw() + 
+  draw_label(
+    "Confidence Intervals of Model Parameter Estimates",
+    x = 0,
+    hjust = 0
+  ) +
+  theme(
+    plot.margin = margin(0, 0, 0, 7)
+  )
+
+top_row <- plot_grid(p1,p2,p3, ncol = 3, nrow = 1)
+plot_grid(title, 
+          top_row, p4, ncol = 1, rel_heights = c(0.1, 1,1))
+
 ```
 
-We then proceed to our hypothesis test.
+We then proceed to our **hypothesis test**.
 As stated in our introduction, given that this issue has become highly polarized by political party, we suspect that liberal political ideology and political attitudes will correlate with belief that abortion is more justified.
@@
 
-HA (alternative hypothesis: An individual’s political ideology (ideology), their beliefs about child autonomy (cai), the importance of God in their life (godimportant), their view on whether respect for authority is good (respectauthority), and their pride in being American (nationalpride) are significant predictors of aj, and the variables’ respective slope point estimates will not be 0.
+HA (alternative hypothesis): An individual’s political ideology (ideology), their beliefs about child autonomy (cai), the importance of God in their life (godimportant), their view on whether respect for authority is good (respectauthority), and their pride in being American (nationalpride) are significant predictors of aj, and the variables’ respective slope point estimates will not be 0.
 
@@
 
-Prediction:
+We now consider the strength of our model in **predicting** the attitude toward the justifiability of abortion of a person of certain characteristics.
 
-We now consider the strength of our model in predicting the attitude toward the justifiability of abortion of a person of certain characteristics.
+Earlier, we guessed that liberal ideology and liberal-leaning attitudes on other social issues will correlate with belief that abortion is more justified.
+For a person surveyed in 2011 who identifies with a left leaning political ideology ("1"), has a high child autonomy index of 2, considers God and religion unimportant ("1"), and has little respect for authority, our our model predicts that their attitude toward the justifiability of abortion is 9.27, that is to say, very amenable to the justifiability of abortion.
+This supports our hypothesis based on contextual knowledge of this issue.
 
-For a person who [describe liberal characteristics here], our model predicts that their attitude toward the justifiability of abortion is [write what the model would predict].
+```{r predict-new}
+#| echo: false
+new_person <- tibble(year = 2011, ideology = 1, cai = 2, godimportant = 1, respectauthority = -1)
+predict(abortion_fit, new_data = new_person)
 
-This does/does not support our hypothesis based on contextual knowledge of this issue.
-
-However, calculating the R-squared and root-mean-squared-error (RMSE) values of our model shows that our model is not very strong at predicting attitude toward the justifiability of abortion, with a R^2 of __ and RMSE of __:
-
-[R^2, RMSE output]
-
-```{r predict}
-abortion_training_pred <- predict(abortion_fit, abortion_training) %>%
-  bind_cols(abortion_training %>% select(aj, year, ideology, cai, 
-                                         godimportant, respectauthority, 
-                                         nationalpride))
-abortion_training_pred
 ```
+However, calculating the R-squared and root-mean-squared-error (RMSE) values of our model shows that our model is not very strong at predicting attitude toward the justifiability of abortion, with a R^2 of 0.247 and RMSE of 2.53.
 
@@
 We will look into this further now with cross-validation, then a test in AIC and BIC statistics.
+
+```{r metrics}
+#| echo: false
+abortion_training_pred <- predict(abortion_fit, abortion_training) %>%
+  bind_cols(abortion_training %>% select(aj, year, ideology, cai, 
+                                         godimportant, respectauthority))
+rsq(abortion_training_pred, 
+                 truth = aj, 
+                 estimate = .pred)
+rmse(abortion_training_pred,
+                 truth = aj,
+                 estimate = .pred)
+```
 
@@
 ```{r cv}
+#| echo: false
 set.seed(206)
@@
 ```{r cv-table}
+#| echo: false
 cv_metrics <- collect_metrics(abortion_fit_rs, summarize = FALSE) 
@@
 ```{r aic-bic}
+#| echo: false
 glance(abortion_fit) %>% 

### 2022-06-20 Final/FinalReport.qmd
edited EDA and Prediction to fit full model with NP

@@
   select(-starts_with("wave"), -starts_with("wvsccode"))
+```
+
+```{r make-factor}
+#| echo: false
+abortion_data$year <- as.factor(abortion_data$year)
+abortion_data$respectauthority <- as.factor(abortion_data$respectauthority)
+abortion_data$respectauthority <- factor(abortion_data$respectauthority, levels=c('0', '-1', '1'))
+abortion_data$nationalpride <- as.factor(abortion_data$nationalpride)
 ```
@@
 h1 <- ggplot(abortion_data, aes(x = year)) +
-  geom_histogram(fill = "orange") + 
+  geom_bar(fill = "orange") + 
   labs(x = "Year of Survey")
@@
 h5 <- ggplot(abortion_data, aes(x = respectauthority)) +
-  geom_histogram(fill = "orange") + 
+  geom_bar(fill = "orange") +
   labs(x = "Respect for Authority")
 
-plot_grid(h1, h2, h3, h4, h5, ncol = 3, nrow = 2)
+h6 <- ggplot(abortion_data, aes(x = nationalpride)) +
+  geom_bar(fill = "orange") +
+  labs(x = "National Pride")
+
+plot_grid(h1, h2, h3, h4, h5, h6, ncol = 3, nrow = 2)
 
@@
 
-plot_grid(b1, b2, b3, b4, b5, ncol = 3, nrow = 2)
+b6 <- ggplot(abortion_data, aes(x = nationalpride, group = nationalpride, 
+                          y = aj)) +
+  geom_boxplot()+
+  labs(y = "Abortion Attitude", x = "National Pride")
+
+plot_grid(b1, b2, b3, b4, b5, b6, ncol = 3, nrow = 2)
 ```
@@
 
-```{r clean-data2}
-#| echo: false
-abortion_data$year <- as.factor(abortion_data$year)
-abortion_data$respectauthority <- as.factor(abortion_data$respectauthority)
-abortion_data$respectauthority <- factor(abortion_data$respectauthority, levels=c('0', '-1', '1'))
-abortion_data$nationalpride <- as.factor(abortion_data$nationalpride)
-```
-
 Year of survey, Respect for Authority, and National Pride were changed into factors since they are categorical variables.
@@
   step_zv(all_predictors())
-
-abortion_rec
 ```
@@
   step_zv(all_predictors())
-
-abortion_rec2
 ```
@@
   step_zv(all_predictors())
-
-abortion_rec3
 ```
@@
   fit(data = abortion_training)
-tidy(abortion_fit) %>%
-  kable(digits = 3)
 
@@
   fit(data = abortion_training)
-tidy(abortion_fit2) %>%
-  kable(digits = 3)
 
@@
   fit(data = abortion_training)
-tidy(abortion_fit3) %>%
-  kable(digits = 3)
 ```
@@
 
-```{r fit-model}
-#| echo = false
-abortion_fit <- abortion_wflow %>%
-  fit(data = abortion_training)
+```{r display-model}
+#| echo: false
 tidy(abortion_fit) %>%
@@
 
+```{r}
+df <- tidy(abortion_fit, conf.int = TRUE)
+p0 <- ggplot(df, aes(x = term, y = estimate, ymin = conf.low, ymax = conf.high)) +
+  geom_pointrange()
+df
+p0
+```
+
 ```{r ci}
@@
 intercept <- df[1,]
-yr <- df[2,]
-cai <- df[4,]
-df <- df[-c(1,2,4),]
+PI_IG <- df[c(2,4),]
+CAI <- df[3,]
+YS <- df[c(5,6,7),]
+RA_NP <- df[c(8,9,10),]
 
@@
   labs(x = NULL)
-p2 <- ggplot(yr, aes(x = term, y = estimate, ymin = conf.low, ymax = conf.high)) +
+
+p2 <- ggplot(YS, aes(x = term, y = estimate, ymin = conf.low, ymax = conf.high)) +
   geom_pointrange() +
   labs(x = NULL) +
-  scale_x_discrete(labels=c("Year of Survey"))
-p3 <- ggplot(cai, aes(x = term, y = estimate, ymin = conf.low, ymax = conf.high)) +
+  scale_x_discrete(labels=c("1999", "2006", "2011"))
+
+p3 <- ggplot(CAI, aes(x = term, y = estimate, ymin = conf.low, ymax = conf.high)) +
   geom_pointrange()+
@@
 
-p4 <- ggplot(df, aes(x = term, y = estimate, ymin = conf.low, ymax = conf.high)) +
+p4 <- ggplot(PI_IG, aes(x = term, y = estimate, ymin = conf.low, ymax = conf.high)) +
   geom_pointrange()+
   labs(x = NULL) +
-  scale_x_discrete(breaks=c("godimportant","ideology","respectauthority"),
-        labels=c("Importance of God", "Political Ideology", "Respect for Authority"))
+  scale_x_discrete(breaks=c("ideology","godimportant"),
+        labels=c("Ideology", "Importance God"))
 
-title <- ggdraw() + 
+p5 <- ggplot(RA_NP, aes(x = term, y = estimate, ymin = conf.low, ymax = conf.high)) +
+  geom_pointrange()+
+  labs(x = NULL) +
+  scale_x_discrete(breaks=c("respectauthority_X.1", "respectauthority_X1", "nationalpride_X1"),
+        labels=c("RA=No", "RA=Yes", "NP=Yes"))
+
+title_factor <- ggdraw() + 
   draw_label(
-    "Confidence Intervals of Model Parameter Estimates",
+    "Confidence Intervals of Estimated Model Parameters (Factors)",
     x = 0,
@@
 
-top_row <- plot_grid(p1,p2,p3, ncol = 3, nrow = 1)
-plot_grid(title, 
-          top_row, p4, ncol = 1, rel_heights = c(0.1, 1,1))
+title_numeric <- ggdraw() + 
+  draw_label(
+    "Confidence Intervals of Estimated Model Parameters (Numeric)",
+    x = 0,
+    hjust = 0
+  ) +
+  theme(
+    plot.margin = margin(0, 0, 0, 7)
+  )
+
+top_row <- plot_grid(p2,p5, ncol = 2, nrow = 1)
+bottom_row <- plot_grid(p1,p3,p4, ncol = 3, nrow = 1)
+plot_grid(title_factor, top_row, title_numeric, bottom_row, 
+          ncol = 1, rel_heights = c(0.1, 1,0.1,1))
 
@@
 Hence, we can conclude that the data provide convincing evidence that an individual’s political ideology, beliefs about child autonomy, and the importance of God in their life are significant predictors of their abortion attitudes.
+
 However, we fail to reject the null hypothesis for the predictors respectauthority and nationalpride, as the p-values for each level of these predictors are above the 0.05 significance threshold.
@@
 Earlier, we guessed that liberal ideology and liberal-leaning attitudes on other social issues will correlate with belief that abortion is more justified.
-For a person surveyed in 2011 who identifies with a left leaning political ideology ("1"), has a high child autonomy index of 2, considers God and religion unimportant ("1"), and has little respect for authority, our our model predicts that their attitude toward the justifiability of abortion is 9.27, that is to say, very amenable to the justifiability of abortion.
+For a person surveyed in 2011 who identifies with a left leaning political ideology ("1"), has a high child autonomy index of 2, considers God and religion unimportant ("1"), and has little respect for authority, our our model predicts that their attitude toward the justifiability of abortion is 9.24, that is to say, very amenable to the justifiability of abortion.
 This supports our hypothesis based on contextual knowledge of this issue.
@@
 #| echo: false
-new_person <- tibble(year = 2011, ideology = 1, cai = 2, godimportant = 1, respectauthority = -1)
+new_person <- tibble(year = "2011", ideology = 1, cai = 2, godimportant = 1, 
+                     respectauthority = "-1", nationalpride = "0")
 predict(abortion_fit, new_data = new_person)

### 2022-06-20 Final/FinalReport.qmd
fixed latex

@@
 
-Understanding public attitudes on divisive political issues is an important way for political leaders to mobilize voters and for lawmakers to draft laws that represent their constituents.
-While it may be easier to poll constituents' positions on an issue, it can be challenging to assess the complex factors that influence and predict those stances.
+Understanding public attitudes on divisive political issues is an
+important way for political leaders to mobilize voters and for lawmakers
+to draft laws that represent their constituents. While it may be easier
+to poll constituents' positions on an issue, it can be challenging to
+assess the complex factors that influence and predict those stances.
 
-Abortion is one such divisive issue in the United States.
-Both pro-choice and pro-life groups have a history of mobilizing in states across the country in support of legislation for their respective sides (Ziegler, 2020).
-However, following the Supreme Court's 1973 decision in Roe v.
-Wade, a ruling which protected an individual’s right to have an abortion before fetal viability, the issue has risen in political salience.
-Both the pro-choice and pro-life movements have gained national prominence, and the two major political parties have polarized around the issue, with the Democratic Party in favor of and the Republican Party against policies legalizing and increasing access to abortion (Weinberger 2022).
-Abortion has also increasingly become a key issue that voters consider when making their choice at the ballot box, with an increasing share of Americans identifying as "single-issue voters" regarding abortion (Brenan 2020).
+Abortion is one such divisive issue in the United States. Both
+pro-choice and pro-life groups have a history of mobilizing in states
+across the country in support of legislation for their respective sides
+(Ziegler, 2020). However, following the Supreme Court's 1973 decision in
+Roe v. Wade, a ruling which protected an individual's right to have an
+abortion before fetal viability, the issue has risen in political
+salience. Both the pro-choice and pro-life movements have gained
+national prominence, and the two major political parties have polarized
+around the issue, with the Democratic Party in favor of and the
+Republican Party against policies legalizing and increasing access to
+abortion (Weinberger 2022). Abortion has also increasingly become a key
+issue that voters consider when making their choice at the ballot box,
+with an increasing share of Americans identifying as "single-issue
+voters" regarding abortion (Brenan 2020).
 
-In May 2022, a leaked draft opinion revealed that the US Supreme Court is prepared to overturn Roe v.
-Wade.
-Overturning Roe would dramatically change the trajectory of abortion politics in the US.
-Unless Congress passes a national policy, states would be able to decide whether or not to legalize abortion and gain much greater leverage in regulating access to the procedure (Weinberger 2022).
+In May 2022, a leaked draft opinion revealed that the US Supreme Court
+is prepared to overturn Roe v. Wade. Overturning Roe would dramatically
+change the trajectory of abortion politics in the US. Unless Congress
+passes a national policy, states would be able to decide whether or not
+to legalize abortion and gain much greater leverage in regulating access
+to the procedure (Weinberger 2022).
 
-Given the potential overturning of Roe and the polarizing nature of the issue, it is important to understand how the American public feels about whether abortion should be legal or not, how accessible the procedure should be, and which factors influence these opinions.
-Understanding public opinion on the issue will ensure that political leaders are able to mobilize the correct constituencies, and that policy experts are able to pass policies on this issue that accurately reflect the preferences of the American people.
+Given the potential overturning of Roe and the polarizing nature of the
+issue, it is important to understand how the American public feels about
+whether abortion should be legal or not, how accessible the procedure
+should be, and which factors influence these opinions. Understanding
+public opinion on the issue will ensure that political leaders are able
+to mobilize the correct constituencies, and that policy experts are able
+to pass policies on this issue that accurately reflect the preferences
+of the American people.
 
-The dataset used in this analysis observes attitudes on the justifiability of abortion among respondents in in the United States across six “waves” of the World Values Survey (1982-2011), which is administered every few years and collects information about people’s values and beliefs worldwide, alongside basic demographic characteristics.
-The data are collected via face-to-face interviews at the respondents’ homes.
-In this data set, the response variable, justifiability of abortion, is a numerical measure on a scale of 1 to 10 on the individual person’s attitude toward whether abortion is justifiable or not.
-Individuals responded 1 for “abortion is never justified” and 10 for “abortion is always justified.
+The dataset used in this analysis observes attitudes on the
+justifiability of abortion among respondents in in the United States
+across six "waves" of the World Values Survey (1982-2011), which is
+administered every few years and collects information about people's
+values and beliefs worldwide, alongside basic demographic
+characteristics. The data are collected via face-to-face interviews at
+the respondents' homes. In this data set, the response variable,
+justifiability of abortion, is a numerical measure on a scale of 1 to 10
+on the individual person's attitude toward whether abortion is
+justifiable or not. Individuals responded 1 for "abortion is never
+justified" and 10 for "abortion is always justified.
 
-Our research question is as follows: “Do an individual’s political ideology and their personal attitudes/preferences towards other issues, such as the importance of religion in their life and their respect for authority, among others, predict their attitude on the justifiability of abortion?” We will attempt to answer this question using an EDA-informed predictive model.
-Given that this issue has become highly polarized by political party, we predict that liberal ideology and liberal-leaning attitudes on other issues will correlate with belief that abortion is more justified.
+Our research question is as follows: "Do an individual's political
+ideology and their personal attitudes/preferences towards other issues,
+such as the importance of religion in their life and their respect for
+authority, among others, predict their attitude on the justifiability of
+abortion?" We will attempt to answer this question using an EDA-informed
+predictive model. Given that this issue has become highly polarized by
+political party, we predict that liberal ideology and liberal-leaning
+attitudes on other issues will correlate with belief that abortion is
+more justified.
 
@@
 
-From the above visualization of the distribution of attitudes toward the justifiability of abortion, we observe that it is not a bell shape, but rather trimodal.
-This may be because the question's phrasing is similar to a yes/no question, but respondents were asked to give their level of agreement on a scale of 1-10.
-This may worsen the ability of a multiple linear regression (MLR) model to fit the data.
-Consequently, further research may choose to truncate the attitude on the justifiability of abortion into a categorical variable such as (Agree, Disagree, Undecided), conducting a binomial or multinomial logistic regression thereafter.
+From the above visualization of the distribution of attitudes toward the
+justifiability of abortion, we observe that it is not a bell shape, but
+rather trimodal. This may be because the question's phrasing is similar
+to a yes/no question, but respondents were asked to give their level of
+agreement on a scale of 1-10. This may worsen the ability of a multiple
+linear regression (MLR) model to fit the data. Consequently, further
+research may choose to truncate the attitude on the justifiability of
+abortion into a categorical variable such as (Agree, Disagree,
+Undecided), conducting a binomial or multinomial logistic regression
+thereafter.
 
-We now extend our exploratory data analysis (EDA) to the predictor variables of interest: year of survey, ideology, Child Autonomy Index, Importance of God, Respect for Authority, and National Pride.
-The EDA for each variable comprises a histogram and a boxplot of the response variable, grouped by value of the predictor.
-We also provide a correlation matrix to detect any multicollinearity between our predictor variables, which would increase the uncertainty of our model’s parameters.
+We now extend our exploratory data analysis (EDA) to the predictor
+variables of interest: year of survey, ideology, Child Autonomy Index,
+Importance of God, Respect for Authority, and National Pride. The EDA
+for each variable comprises a histogram and a boxplot of the response
+variable, grouped by value of the predictor. We also provide a
+correlation matrix to detect any multicollinearity between our predictor
+variables, which would increase the uncertainty of our model's
+parameters.
 
@@
 
-Given that our response variable, attitude towards the justifiability of abortion, is measured on a numeric scale from 1 to 10, and that there are multiple predictor variables being tested against the null hypothesis, we will conduct multiple linear regression (MLR) to model the effect of these predictors on the variation in abortion attitudes.
+Given that our response variable, attitude towards the justifiability of
+abortion, is measured on a numeric scale from 1 to 10, and that there
+are multiple predictor variables being tested against the null
+hypothesis, we will conduct multiple linear regression (MLR) to model
+the effect of these predictors on the variation in abortion attitudes.
 
-We chose to consider the predictor variables our exploratory data analysis (EDA) suggested were potentially correlated with the response variable.
-Because the predictor variables were discrete, our EDA was presented as a series of box plots.
-We observed the differences in median and quartile values of abortion attitude across different predictor values, ruling out variables with no significant influence on the outcome.
-This led to the following predictor variables: Year of survey, Political Ideology, Child Autonomy Index, Importance of God, Respect for Authority, and National Pride.
+We chose to consider the predictor variables our exploratory data
+analysis (EDA) suggested were potentially correlated with the response
+variable. Because the predictor variables were discrete, our EDA was
+presented as a series of box plots. We observed the differences in
+median and quartile values of abortion attitude across different
+predictor values, ruling out variables with no significant influence on
+the outcome. This led to the following predictor variables: Year of
+survey, Political Ideology, Child Autonomy Index, Importance of God,
+Respect for Authority, and National Pride.
 
-*A note on data cleaning: Many survey questions were only added from 1995 onward, so observations before 1995 have a high number of missing values.
-For these reasons, observations prior to 1995 were removed.
+\*A note on data cleaning: Many survey questions were only added from
+1995 onward, so observations before 1995 have a high number of missing
+values. For these reasons, observations prior to 1995 were removed.
 
-Year of survey, Respect for Authority, and National Pride were changed into factors since they are categorical variables.
-Year has 4 levels (one corresponding to each year), Respect for Authority has 3 levels (0 being the baseline - neutral, -1 being respect for authority bad, and 1 being respect for authority good), and National Pride is a binary (0 being the baseline - having no feelings of national pride, and 1 being having feelings of national pride) Political Ideology, Child Autonomy Index, and Importance of God had numerical sample spaces and were therefore treated as numerical variables.*
+Year of survey, Respect for Authority, and National Pride were changed
+into factors since they are categorical variables. Year has 4 levels
+(one corresponding to each year), Respect for Authority has 3 levels (0
+being the baseline - neutral, -1 being respect for authority bad, and 1
+being respect for authority good), and National Pride is a binary (0
+being the baseline - having no feelings of national pride, and 1 being
+having feelings of national pride) Political Ideology, Child Autonomy
+Index, and Importance of God had numerical sample spaces and were
+therefore treated as numerical variables.\*
 
@@
 
-We conducted a 75% - 25% data split into training and testing sets, using the random seed $206$.
+We conducted a 75% - 25% data split into training and testing sets,
+using the random seed $206$.
 
@@
 
-Fitting the MLR, we obtain the following table of parameter point estimates.
+Fitting the MLR, we obtain the following table of parameter point
+estimates.
 
@@
 
-$$\begin{align}y ~=~ &6.596 ~+~ 0.468x_1 ~+~ 0.33x_2 ~+~ 0.339x_3 ~+~ 0.521x_4 ~+~ 0.14x_5 \\&~-~ (0.28x_6 ~+~ 0.287x_7 ~+~ 0.179x_8 ~+~ 0.11x_9)\end{align}$$
+$$\begin{aligned}y ~=~ &6.596 ~+~ 0.468x_1 ~+~ 0.33x_2 ~+~ 0.339x_3 ~+~ 0.521x_4 ~+~ 0.14x_5 \\
+&~-~ (0.28x_6 ~+~ 0.287x_7 ~+~ 0.179x_8 ~+~ 0.11x_9)\end{aligned}$$
 
-where $x_1 ~=~ \text{child autonomy index}$, $x_2 ~=~ \text{year 1999}$, $x_3 ~=~ \text{year 2006}$, $x_4 ~=~ \text{year 2011}$, $x_5 ~=~ \text{respect authority -1}$, $x_6 ~=~ \text{ideology}$, $x_7 ~=~ \text{God important}$, $x_8 ~=~ \text{respect authority 1}$, and $x_9 ~=~ \text{national pride}$.
+where $x_1 ~=~ \text{child autonomy index}$, $x_2 ~=~ \text{year 1999}$,
+$x_3 ~=~ \text{year 2006}$, $x_4 ~=~ \text{year 2011}$,
+$x_5 ~=~ \text{respect authority -1}$, $x_6 ~=~ \text{ideology}$,
+$x_7 ~=~ \text{God important}$, $x_8 ~=~ \text{respect authority 1}$,
+and $x_9 ~=~ \text{national pride}$.
 
@@
 
-We will take a quick look at the 95% **confidence intervals** of the parameters from our fitted model, displayed in the following plots:
+We will take a quick look at the 95% **confidence intervals** of the
+parameters from our fitted model, displayed in the following plots:
 
@@
 
-We then proceed to our **hypothesis test**.
-As stated in our introduction, given that this issue has become highly polarized by political party, we suspect that liberal political ideology and political attitudes will correlate with belief that abortion is more justified.
+We then proceed to our **hypothesis test**. As stated in our
+introduction, given that this issue has become highly polarized by
+political party, we suspect that liberal political ideology and
+political attitudes will correlate with belief that abortion is more
+justified.
 
-We define the following as liberal political ideology and attitudes:
-- ideology: values closer to 1 (most left-leaning)
-- cai: values closer to 2 (belief in child self-determination and independence)
-- godimportant: values closer to 1 (belief that God is not important)
-- respectauthority: -1 (belief that greater respect for authority is a bad thing)
-- nationalpride: 0 (not proud to be American)
+We define the following as liberal political ideology and attitudes: -
+ideology: values closer to 1 (most left-leaning) - cai: values closer to
+2 (belief in child self-determination and independence) - godimportant:
+values closer to 1 (belief that God is not important) -
+respectauthority: -1 (belief that greater respect for authority is a bad
+thing) - nationalpride: 0 (not proud to be American)
 
@@
 
-H0 (null hypothesis): An individual’s political ideology (ideology), their beliefs about child autonomy (cai), the importance of God in their life (godimportant), their view on whether respect for authority is good (respectauthority), and their pride in being American (nationalpride) are not significant predictors of aj, and the variables’ respective slope point estimates will be 0.
+H0 (null hypothesis): An individual's political ideology (ideology),
+their beliefs about child autonomy (cai), the importance of God in their
+life (godimportant), their view on whether respect for authority is good
+(respectauthority), and their pride in being American (nationalpride)
+are not significant predictors of aj, and the variables' respective
+slope point estimates will be 0.
 
-HA (alternative hypothesis): An individual’s political ideology (ideology), their beliefs about child autonomy (cai), the importance of God in their life (godimportant), their view on whether respect for authority is good (respectauthority), and their pride in being American (nationalpride) are significant predictors of aj, and the variables’ respective slope point estimates will not be 0.
+HA (alternative hypothesis): An individual's political ideology
+(ideology), their beliefs about child autonomy (cai), the importance of
+God in their life (godimportant), their view on whether respect for
+authority is good (respectauthority), and their pride in being American
+(nationalpride) are significant predictors of aj, and the variables'
+respective slope point estimates will not be 0.
 
-Looking at the p-values in our output displayed above, we can reject the null hypothesis for the predictors ideology, cai, and godimportant, as the p-values are very small, way below the 0.05 significance threshold.
-Hence, we can conclude that the data provide convincing evidence that an individual’s political ideology, beliefs about child autonomy, and the importance of God in their life are significant predictors of their abortion attitudes.
+Looking at the p-values in our output displayed above, we can reject the
+null hypothesis for the predictors ideology, cai, and godimportant, as
+the p-values are very small, way below the 0.05 significance threshold.
+Hence, we can conclude that the data provide convincing evidence that an
+individual's political ideology, beliefs about child autonomy, and the
+importance of God in their life are significant predictors of their
+abortion attitudes.
 
-However, we fail to reject the null hypothesis for the predictors respectauthority and nationalpride, as the p-values for each level of these predictors are above the 0.05 significance threshold.
-Hence, we cannot conclude that the data provide convincing evidence that an individual’s view on respect for authority and their pride in being American are significant predictors of their abortion attitudes.
+However, we fail to reject the null hypothesis for the predictors
+respectauthority and nationalpride, as the p-values for each level of
+these predictors are above the 0.05 significance threshold. Hence, we
+cannot conclude that the data provide convincing evidence that an
+individual's view on respect for authority and their pride in being
+American are significant predictors of their abortion attitudes.
 
-We now consider the strength of our model in **predicting** the attitude toward the justifiability of abortion of a person of certain characteristics.
+We now consider the strength of our model in **predicting** the attitude
+toward the justifiability of abortion of a person of certain
+characteristics.
 
-Earlier, we guessed that liberal ideology and liberal-leaning attitudes on other social issues will correlate with belief that abortion is more justified.
-For a person surveyed in 2011 who identifies with a left leaning political ideology ("1"), has a high child autonomy index of 2, considers God and religion unimportant ("1"), and has little respect for authority, our our model predicts that their attitude toward the justifiability of abortion is 9.24, that is to say, very amenable to the justifiability of abortion.
-This supports our hypothesis based on contextual knowledge of this issue.
+Earlier, we guessed that liberal ideology and liberal-leaning attitudes
+on other social issues will correlate with belief that abortion is more
+justified. For a person surveyed in 2011 who identifies with a left
+leaning political ideology ("1"), has a high child autonomy index of 2,
+considers God and religion unimportant ("1"), and has little respect for
+authority, our our model predicts that their attitude toward the
+justifiability of abortion is 9.24, that is to say, very amenable to the
+justifiability of abortion. This supports our hypothesis based on
+contextual knowledge of this issue.
 
@@
 ```
-However, calculating the R-squared and root-mean-squared-error (RMSE) values of our model shows that our model is not very strong at predicting attitude toward the justifiability of abortion, with a R^2 of 0.247 and RMSE of 2.53.
 
-We will discuss potential reasons for this lackluster performance in our model diagnosis and conclusion sections; however, we will first conduct model evaluation and comparison to ascertain if our chosen model is still a better fit to the data compared to some alternatives.
+However, calculating the R-squared and root-mean-squared-error (RMSE)
+values of our model shows that our model is not very strong at
+predicting attitude toward the justifiability of abortion, with a R\^2
+of 0.247 and RMSE of 2.53.
 
-A quick glance at our data might seem that the models are not that strong at predicting the abortion attitude (aj) of a given observation.
-We will look into this further now with cross-validation, then a test in AIC and BIC statistics.
+We will discuss potential reasons for this lackluster performance in our
+model diagnosis and conclusion sections; however, we will first conduct
+model evaluation and comparison to ascertain if our chosen model is
+still a better fit to the data compared to some alternatives.
+
+A quick glance at our data might seem that the models are not that
+strong at predicting the abortion attitude (aj) of a given observation.
+We will look into this further now with cross-validation, then a test in
+AIC and BIC statistics.
 
@@
 
-To see how well our model fits the data on the attitudes toward the justifiability of abortion, we conducted a 10-fold cross-validation, calculating various metrics – the Akaike information criterion (AIC), Bayesian Information Criterion (BIC) and adjusted R-squared value – to evaluate the performance of our model on 10 subsets of our data.
+To see how well our model fits the data on the attitudes toward the
+justifiability of abortion, we conducted a 10-fold cross-validation,
+calculating various metrics -- the Akaike information criterion (AIC),
+Bayesian Information Criterion (BIC) and adjusted R-squared value -- to
+evaluate the performance of our model on 10 subsets of our data.
 
-We also compared this to two alternative models that used other combinations of predictors from the data set.
+We also compared this to two alternative models that used other
+combinations of predictors from the data set.
 
-[3 line plots, displaying AIC, BIC and R^2 value respectively, against Fold number. Each plot should have the line graph for each of the 3 models we used)
+\[3 line plots, displaying AIC, BIC and R\^2 value respectively, against
+Fold number. Each plot should have the line graph for each of the 3
+models we used)
 
@@
 
-From the above AIC and BIC values, we have evidence that our chosen model is the strongest model and best fit for the data compared to two alternatives with fewer predictors.
-Our model has the lowest AIC and BIC values, which are usually penalized for having more predictors, but in this case the model is strong enough to overcome that penalty.
+From the above AIC and BIC values, we have evidence that our chosen
+model is the strongest model and best fit for the data compared to two
+alternatives with fewer predictors. Our model has the lowest AIC and BIC
+values, which are usually penalized for having more predictors, but in
+this case the model is strong enough to overcome that penalty.
 
@@
 
-Linearity [residual plot vs fitted values]
+Linearity \[residual plot vs fitted values\]
 
@@
 
-Independence [residual plot vs id]
+Independence \[residual plot vs id\]
 
@@
 
-[paste here after finalizing the google doc]
+\[paste here after finalizing the google doc\]
 

### 2022-06-20 Final/FinalReport.qmd
cosmetic changes complete

@@
 ---
-title: "Draft"
+title: "Investigating Predictors for American Attitudes on Justifiability of Abortion"
 subtitle: "STA 210 - Project"
@@
 
-plot_grid(h1, h2, h3, h4, h5, h6, ncol = 3, nrow = 2)
+title1 <- ggdraw() + 
+  draw_label(
+    "Distributions of Predictor variables",
+    x = 0,
+    hjust = 0
+  ) +
+  theme(
+    plot.margin = margin(0, 0, 0, 7)
+  )
 
+top_row1 <- plot_grid(h1,h2,h3, ncol = 3, nrow = 1)
+bottom_row1 <- plot_grid(h4,h5,h6, ncol = 3, nrow = 1)
+plot_grid(title1, top_row1, bottom_row1, 
+          ncol = 1, rel_heights = c(0.1, 1,1))
+# ADD CAPTION
 ```
@@
 plot_grid(b1, b2, b3, b4, b5, b6, ncol = 3, nrow = 2)
+
+# ADD TITLE AND CAPTION
 ```
@@
 ggpairs(abortion_data, 
-  columns = c("year", "ideology", "cai", "godimportant", "respectauthority"),
+  columns = c("year", "ideology", "cai", "godimportant", "respectauthority", "nationalpride"),
   columnLabels = c("YS", "PI", "CAI", "IG", 
-                   "RA")) +
+                   "RA", "NP")) +
    theme(
@@
     )
+# ADD TITLE AND CAPTION
 ```
@@
 
-\*A note on data cleaning: Many survey questions were only added from
-1995 onward, so observations before 1995 have a high number of missing
-values. For these reasons, observations prior to 1995 were removed.
+Many survey questions were only added from 1995 onward, so observations before 1995 have a high number of missing values. For these reasons, observations prior to 1995 were removed.
 
@@
 Index, and Importance of God had numerical sample spaces and were
-therefore treated as numerical variables.\*
+therefore treated as numerical variables.
 
@@
 #| echo: false
+#| warning: false
 set.seed(206)
@@
 #| echo: false
+#| warning: false
 abortion_spec <- linear_reg() %>%
@@
 #| echo: false
+#| warning: false
 abortion_rec <- recipe(aj ~ year + ideology + cai + godimportant +
@@
 
-```{r}
-df <- tidy(abortion_fit, conf.int = TRUE)
-p0 <- ggplot(df, aes(x = term, y = estimate, ymin = conf.low, ymax = conf.high)) +
-  geom_pointrange()
-df
-p0
-```
-
 ```{r ci}
@@
           ncol = 1, rel_heights = c(0.1, 1,0.1,1))
-
+#ADD CAPTION
 ```
@@
 
-H0 (null hypothesis): An individual's political ideology (ideology),
+$H_0$ (null hypothesis): An individual's political ideology (ideology),
 their beliefs about child autonomy (cai), the importance of God in their
@@
 
-HA (alternative hypothesis): An individual's political ideology
+$H_A$ (alternative hypothesis): An individual's political ideology
 (ideology), their beliefs about child autonomy (cai), the importance of
@@
 Looking at the p-values in our output displayed above, we can reject the
-null hypothesis for the predictors ideology, cai, and godimportant, as
+null hypothesis for the predictors political ideology, child autonomy index, and importance of God, as
 the p-values are very small, way below the 0.05 significance threshold.
@@
 However, we fail to reject the null hypothesis for the predictors
-respectauthority and nationalpride, as the p-values for each level of
+respect for authority and national pride, as the p-values for each level of
 these predictors are above the 0.05 significance threshold. Hence, we
@@
 #| echo: false
+#| eval: false
 new_person <- tibble(year = "2011", ideology = 1, cai = 2, godimportant = 1, 
@@
 #| echo: false
+#| eval: false
+#| warning: false
 abortion_training_pred <- predict(abortion_fit, abortion_training) %>%

### 2022-06-20 Final/FinalReport.qmd
minor edits to introduction

@@
 
-The dataset used in this analysis observes attitudes on the
-justifiability of abortion among respondents in in the United States
-across six "waves" of the World Values Survey (1982-2011), which is
-administered every few years and collects information about people's
-values and beliefs worldwide, alongside basic demographic
-characteristics. The data are collected via face-to-face interviews at
-the respondents' homes. In this data set, the response variable,
-justifiability of abortion, is a numerical measure on a scale of 1 to 10
-on the individual person's attitude toward whether abortion is
-justifiable or not. Individuals responded 1 for "abortion is never
-justified" and 10 for "abortion is always justified.
+The dataset used in this analysis observes attitudes on the justifiability of abortion among respondents in in the United States across six “waves” of the World Values Survey (1982-2011), which is administered every few years and collects information about people’s values and beliefs worldwide, alongside basic demographic characteristics.
+The data are a nationally representative sample of the American population and are collected via face-to-face interviews at the respondents’ homes.
+In this data set, the response variable, justifiability of abortion, is a numerical measure on a scale of 1 to 10 on the individual person’s attitude toward whether abortion is justifiable or not.
+Individuals responded 1 for “abortion is never justified” and 10 for “abortion is always justified.
 
-Our research question is as follows: "Do an individual's political
-ideology and their personal attitudes/preferences towards other issues,
-such as the importance of religion in their life and their respect for
-authority, among others, predict their attitude on the justifiability of
-abortion?" We will attempt to answer this question using an EDA-informed
-predictive model. Given that this issue has become highly polarized by
-political party, we predict that liberal ideology and liberal-leaning
-attitudes on other issues will correlate with belief that abortion is
-more justified.
+Our research question is as follows: “Do an individual’s political ideology and their personal attitudes/preferences towards other issues, such as the importance of religion in their life and their respect for authority, among others, predict their attitude on the justifiability of abortion?” We will attempt to answer this question using an EDA-informed predictive model.
+Because the data are nationally representative, we can infer that all the observations are independent and the variables of interest may show a linear relationship.
+Given that this issue has become highly polarized by political party, we predict that liberal ideology and liberal-leaning attitudes on other issues will correlate with belief that abortion is more justified.
 
@@
 Looking at the p-values in our output displayed above, we can reject the
-null hypothesis for the predictors political ideology, child autonomy index, and importance of God, as
-the p-values are very small, way below the 0.05 significance threshold.
+null hypothesis for the predictors political ideology, child autonomy index, and importance of God, as the p-values are very small, way below the 0.05 significance threshold.
 Hence, we can conclude that the data provide convincing evidence that an

### 2022-06-20 Final/FinalReport.qmd
model eval done

--- 2022-06-20 Final/FinalReport.qmd
@@
 values and beliefs worldwide, alongside basic demographic
-characteristics. The data are collected via face-to-face interviews at
-the respondents' homes. In this data set, the response variable,
+characteristics. The data are a nationally representative sample of the
+American population and are collected via face-to-face interviews at the
+respondents' homes. In this data set, the response variable,
 justifiability of abortion, is a numerical measure on a scale of 1 to 10
@@
 abortion?" We will attempt to answer this question using an EDA-informed
-predictive model. Given that this issue has become highly polarized by
-political party, we predict that liberal ideology and liberal-leaning
-attitudes on other issues will correlate with belief that abortion is
-more justified.
+predictive model. Because the data are nationally representative, we can
+infer that all the observations are independent and the variables of
+interest may show a linear relationship. Given that this issue has
+become highly polarized by political party, we predict that liberal
+ideology and liberal-leaning attitudes on other issues will correlate
+with belief that abortion is more justified.
 
@@
 
-Many survey questions were only added from 1995 onward, so observations before 1995 have a high number of missing values. For these reasons, observations prior to 1995 were removed.
+Many survey questions were only added from 1995 onward, so observations
+before 1995 have a high number of missing values. For these reasons,
+observations prior to 1995 were removed.
 
@@
 Looking at the p-values in our output displayed above, we can reject the
-null hypothesis for the predictors political ideology, child autonomy index, and importance of God, as
-the p-values are very small, way below the 0.05 significance threshold.
-Hence, we can conclude that the data provide convincing evidence that an
-individual's political ideology, beliefs about child autonomy, and the
-importance of God in their life are significant predictors of their
-abortion attitudes.
+null hypothesis for the predictors political ideology, child autonomy
+index, and importance of God, as the p-values are very small, way below
+the 0.05 significance threshold. Hence, we can conclude that the data
+provide convincing evidence that an individual's political ideology,
+beliefs about child autonomy, and the importance of God in their life
+are significant predictors of their abortion attitudes.
 
 However, we fail to reject the null hypothesis for the predictors
-respect for authority and national pride, as the p-values for each level of
-these predictors are above the 0.05 significance threshold. Hence, we
+respect for authority and national pride, as the p-values for each level
+of these predictors are above the 0.05 significance threshold. Hence, we
 cannot conclude that the data provide convincing evidence that an
@@
 
-\[3 line plots, displaying AIC, BIC and R\^2 value respectively, against
-Fold number. Each plot should have the line graph for each of the 3
-models we used)
-
 ```{r cv}
+#| message: false
 #| echo: false
@@
 folds <- vfold_cv(abortion_training, v = 10)
+
 abortion_fit_rs <- abortion_wflow %>%
   fit_resamples(folds)
-
 abortion_fit_rs2 <- abortion_wflow2 %>%
   fit_resamples(folds)
-
 abortion_fit_rs3 <- abortion_wflow3 %>%
   fit_resamples(folds)
+
+metrics1 <- collect_metrics(abortion_fit_rs, summarize = FALSE) %>%
+  mutate(model = "Model 1")
+metrics2 <- collect_metrics(abortion_fit_rs2, summarize = FALSE) %>%
+  mutate(model = "Model 2")
+metrics3 <- collect_metrics(abortion_fit_rs3, summarize = FALSE) %>%
+  mutate(model = "Model 3")
+
+metrics <- bind_rows(metrics1, metrics2, metrics3) %>%
+  arrange(.metric)
+
+ggplot(metrics %>% filter(.metric == "rsq"),
+	aes(x = id, y = .estimate, group = model, color = model)) +
+    geom_point() +
+    geom_line() +
+    labs(title = "R-squared")
 ```
 
-```{r cv-table}
+From the above visualization of R-squared values, we have evidence that
+our chosen model is the strongest model and best fit for the data
+compared to two alternatives with fewer predictors. Below we will
+observe the AIC, BIC, and Adjusted R-squared values for each model as
+well.
+
+```{r aic-bic-adjrsq}
 #| echo: false
-cv_metrics <- collect_metrics(abortion_fit_rs, summarize = FALSE) 
-cv_metrics2 <- collect_metrics(abortion_fit_rs2, summarize = FALSE) 
-cv_metrics3 <- collect_metrics(abortion_fit_rs3, summarize = FALSE)
+abortion_table = matrix(c(glance(abortion_fit)$AIC,
+                          glance(abortion_fit)$BIC,
+                          glance(abortion_fit)$adj.r.squared,
+                          glance(abortion_fit2)$AIC,
+                          glance(abortion_fit2)$BIC, 
+                          glance(abortion_fit2)$adj.r.squared, 
+                          glance(abortion_fit3)$AIC,
+                          glance(abortion_fit3)$BIC,
+                          glance(abortion_fit3)$adj.r.squared), 
+                        ncol = 3, byrow = TRUE)
 
-cv_metrics %>%
-  mutate(.estimate = round(.estimate, 3)) %>%
-  pivot_wider(id_cols = id, names_from = .metric, values_from = .estimate) %>%
-  kable(col.names = c("Fold", "RMSE", "R-squared"))
+colnames(abortion_table) = c('AIC','BIC','Adj. R-squared')
+rownames(abortion_table) = c('Model 1','Model 2','Model 3')
 
-cv_metrics2 %>%
-  mutate(.estimate = round(.estimate, 3)) %>%
-  pivot_wider(id_cols = id, names_from = .metric, values_from = .estimate) %>%
-  kable(col.names = c("Fold", "RMSE", "R-squared"))
-
-cv_metrics3 %>%
-  mutate(.estimate = round(.estimate, 3)) %>%
-  pivot_wider(id_cols = id, names_from = .metric, values_from = .estimate) %>%
-  kable(col.names = c("Fold", "RMSE", "R-squared"))
+final_table = as.table(abortion_table)
+final_table %>%
+  kable(digits = 3)
 ```
 
-```{r aic-bic}
-#| echo: false
-glance(abortion_fit) %>% 
-  select(AIC, BIC)
-
-glance(abortion_fit2) %>% 
-  select(AIC, BIC)
-
-glance(abortion_fit3) %>% 
-  select(AIC, BIC)
-```
-
-From the above AIC and BIC values, we have evidence that our chosen
-model is the strongest model and best fit for the data compared to two
-alternatives with fewer predictors. Our model has the lowest AIC and BIC
-values, which are usually penalized for having more predictors, but in
-this case the model is strong enough to overcome that penalty.
+The table above further supports that Model 1 is our best fitting model.
 

### 2022-06-20 Final/FinalReport.qmd
conditions code started

@@
 
-Conditions check
+Below we will check each of the following conditions for our model:
+Linearity, Constant Variance, Normality, and Independence.
 
-Linearity \[residual plot vs fitted values\]
+```{r fit-aug}
+final_fit <- linear_reg() %>%
+  set_engine("lm") %>%
+  fit(aj ~ year + ideology + cai + godimportant +
+                        respectauthority + nationalpride,
+      data = abortion_data)
 
-Randomness
+final_fit_aug <- augment(final_fit$fit)
+```
 
-Independence \[residual plot vs id\]
+Linearity
+
+```{r linearity}
+#| echo: false
+ggplot(data = final_fit_aug, aes(x = .fitted, y = .resid)) +
+  geom_point(alpha = 0.5) +
+  geom_hline(yintercept = 0, color = "red", linetype = "dashed") +
+  labs(x = "Predicted values", y = "Residuals")
+```
+
+Constant Variance
+
+Normality
+
+```{r normality}
+#| echo: false
+ggplot(final_fit_aug, aes(.resid)) +
+  geom_histogram(aes(y = after_stat(density)), binwidth = 1) +
+  stat_function(fun = dnorm, 
+    args = list(mean = mean(final_fit_aug$.resid), sd = sd(final_fit_aug$.resid)), 
+    lwd = 2,  color = "red"  )
+```
+
+Independence
+
+```{r independence}
+#| echo: false
+ggplot(final_fit_aug, aes(y = .resid, x = 1:nrow(final_fit_aug))) +
+  geom_point(alpha = 0.5) +
+  labs(x = "Order of data collection", y = "Residuals")
+```
 

### 2022-06-20 Final/FinalReport.qmd
added captions and fixed errors

--- 2022-06-20 Final/FinalReport.qmd
@@
 The dataset used in this analysis observes attitudes on the
-justifiability of abortion among respondents in in the United States
+justifiability of abortion among respondents in the United States
 across six "waves" of the World Values Survey (1982-2011), which is
@@
 
-```{r visualization-summary}
+```{r vis-response}
 #| warning: false
@@
   labs(title = "Histogram of Abortion Attitudes",
-       x = "Abortion Justifiability (1-Least, 10-Most)")
+       x = "Abortion Justifiability (1-Least, 10-Most)",
+       caption = "Fig.1: The visual shows that the distribution of attitude toward abortion is trimodal, rather than bell-shaped. This may worsen our model's ability to fit the data.")+
+  theme(plot.caption = element_text(hjust = 0.5))
 ```
@@
   draw_label(
-    "Distributions of Predictor variables",
+    "Distributions of variables of interest",
     x = 0,
@@
 
+caption1 <- ggdraw() + 
+  draw_label(
+    "Fig.2: The visuals show that while some variables in the data set have bell-shaped distributions, others are left-skewed.",
+    x = 0,
+    hjust = 0, 
+    size = 8
+  ) +
+  theme(
+    plot.margin = margin(0, 0, 0, 7)
+  )
+
 top_row1 <- plot_grid(h1,h2,h3, ncol = 3, nrow = 1)
 bottom_row1 <- plot_grid(h4,h5,h6, ncol = 3, nrow = 1)
-plot_grid(title1, top_row1, bottom_row1, 
-          ncol = 1, rel_heights = c(0.1, 1,1))
-# ADD CAPTION
+plot_grid(title1, top_row1, bottom_row1, caption1, 
+          ncol = 1, rel_heights = c(0.1, 1,1, 0.1)) 
 ```
@@
 
-plot_grid(b1, b2, b3, b4, b5, b6, ncol = 3, nrow = 2)
+title2 <-  ggdraw() + 
+  draw_label("Boxplots of Abortion Attitudes against interesting variables",
+    x = 0,
+    hjust = 0
+  ) +
+  theme(
+    plot.margin = margin(0, 0, 0, 7)
+  )
 
-# ADD TITLE AND CAPTION
+caption2 <- ggdraw() + 
+  draw_label("Fig.3: The visuals show that there may be a relationship between the response and each of these variables, making them likely predictors.",
+    x = 0,
+    hjust = 0, size = 7
+  ) +
+  theme(
+    plot.margin = margin(0, 0, 0, 7)
+  )
+
+middle <- plot_grid(b1, b2, b3, b4, b5, b6, ncol = 3, nrow = 2)
+
+plot_grid(title2, middle, caption2, 
+          ncol = 1, rel_heights = c(0.1, 1, 0.1))
+
 ```
@@
    theme(
-    axis.text.y = element_text(size = 8),
+    axis.text.y = element_text(size = 6),
     axis.text.x = element_text(angle = 45, size = 9),
-    strip.text.y = element_text(angle = 0, hjust = 0)
-    )
-# ADD TITLE AND CAPTION
+    strip.text.y = element_text(angle = 0, hjust = 0),
+    plot.caption = element_text(hjust = 0.5)
+    ) +
+  labs(title = "Correlation Matrix of Predictors of Interest", caption = "Fig.4: This matrix shows that the correlations between pairs of predictors are low, with the highest being 0.370.")
+
 ```
-
 \pagebreak
@@
 We conducted a 75% - 25% data split into training and testing sets,
-using the random seed $206$.
+using the random seed $206$. Fitting the MLR, we obtain the following table of parameter point estimates.
 
@@
 #| echo: false
+#| warning: false
 abortion_rec2 <- recipe(aj ~ year + godimportant,
@@
 #| echo: false
+#| warning: false
 abortion_rec3 <- recipe(aj ~ year + ideology + cai + nationalpride,
@@
 #| echo: false
+#| warning: false
 abortion_wflow <- workflow() %>%
@@
 #| echo: false
+#| warning: false
 abortion_fit <- abortion_wflow %>%
@@
 
-Fitting the MLR, we obtain the following table of parameter point
-estimates.
-
 ```{r display-model}
@@
 tidy(abortion_fit) %>%
-  kable(digits = 3)
+  kable(digits = 3, caption = "This table displays the estimate, standard error, statistic and p.value for each parameter of the model.")
 ```
@@
 #| echo: false
+#| warning: false
 df <- tidy(abortion_fit, conf.int = TRUE)
@@
 bottom_row <- plot_grid(p1,p3,p4, ncol = 3, nrow = 1)
-plot_grid(title_factor, top_row, title_numeric, bottom_row, 
-          ncol = 1, rel_heights = c(0.1, 1,0.1,1))
-#ADD CAPTION
+caption3 <- ggdraw() + 
+  draw_label("Fig.5: 95% confidence intervals of estimates for model parameters, using points and error bars.",
+    x = 0,
+    hjust = 0, size = 10
+  ) +
+  theme(
+    plot.margin = margin(0, 0, 0, 7)
+  )
+
+middle <- plot_grid(b1, b2, b3, b4, b5, b6, ncol = 3, nrow = 2)
+
+plot_grid(title_factor, top_row, title_numeric, bottom_row, caption3, 
+          ncol = 1, rel_heights = c(0.1, 1,0.1,1,0.1))
+
 ```
@@
 values of our model shows that our model is not very strong at
-predicting attitude toward the justifiability of abortion, with a R\^2
+predicting attitude toward the justifiability of abortion, with a R-squared
 of 0.247 and RMSE of 2.53.
@@
 To see how well our model fits the data on the attitudes toward the
-justifiability of abortion, we conducted a 10-fold cross-validation,
+justifiability of abortion, we conducted a 10-fold cross-validation with seed 206,
 calculating various metrics -- the Akaike information criterion (AIC),
@@
     geom_line() +
-    labs(title = "R-squared")
+    labs(title = "R-squared", caption = "Fig.6: This plot of R-squared values demonstrates our chosen model 1 is the best of the three.")
 ```
@@
 final_table %>%
-  kable(digits = 3)
+  kable(digits = 3, caption = "With the lowest AIC and BIC and highest adjusted R-squared value, our chosen model is best of the three.")
 ```
@@
 
-Conditions check
+Below we will check each of the following conditions for our model:
+Linearity, Constant Variance, Normality, and Independence.
 
-Linearity \[residual plot vs fitted values\]
+```{r fit-aug}
+#| echo: false
+final_fit <- linear_reg() %>%
+  set_engine("lm") %>%
+  fit(aj ~ year + ideology + cai + godimportant +
+                        respectauthority + nationalpride,
+      data = abortion_data)
 
-Randomness
+final_fit_aug <- augment(final_fit$fit)
+```
 
-Independence \[residual plot vs id\]
+As shown in the following plot of residuals against fitted values, the linearity condition is not met as the residuals are not scattered around the horizontal red line.
+The constant variance condition is also not met.
+
+```{r linearity}
+#| echo: false
+c1 <- ggplot(data = final_fit_aug, aes(x = .fitted, y = .resid)) +
+  geom_point(alpha = 0.5) +
+  geom_hline(yintercept = 0, color = "red", linetype = "dashed") +
+  labs(title = "Plot of residuals v. fitted values", x = "Predicted values", y = "Residuals")
+
+c2 <- ggplot(final_fit_aug, aes(.resid)) +
+  geom_histogram(aes(y = after_stat(density)), binwidth = 1) +
+  stat_function(fun = dnorm, 
+    args = list(mean = mean(final_fit_aug$.resid), sd = sd(final_fit_aug$.resid)), 
+    lwd = 2,  color = "red"  )+
+ labs(title = "Distribution of residuals")
+
+c3 <- ggplot(final_fit_aug, aes(y = .resid, x = 1:nrow(final_fit_aug))) +
+  geom_point(alpha = 0.5) +
+  labs(x = "Order of data collection", y = "Residuals", title = "Plot of residuals v. order of collection")
+
+top_row1 <- plot_grid(c1,c2, ncol = 2, nrow = 1)
+
+caption4 <- ggdraw() + 
+  draw_label("Fig.7: The linearity condition and constant variance condition is clearly not met in this plot of residuals v. fitted values.",
+    x = 0,
+    hjust = 0, size = 8
+  ) +
+  theme(
+    plot.margin = margin(0, 0, 0, 7)
+  )
+caption5 <- ggdraw() + 
+  draw_label("Fig.8: The residuals appear to follow a normal distribution, which looks like a bell-shape, as illustrated in red.",
+    x = 0,
+    hjust = 0, size = 8
+  ) +
+  theme(
+    plot.margin = margin(0, 0, 0, 7)
+  )
+caption6 <- ggdraw() + 
+  draw_label("Fig.9: The residuals plotted by observation order are randomly scattered, showing no structure.",
+    x = 0,
+    hjust = 0, size = 8
+  ) +
+  theme(
+    plot.margin = margin(0, 0, 0, 7)
+  )
+
+plot_grid(top_row1, c3, caption4, caption5, caption6,
+          ncol = 1, rel_heights = c(1, 1,0.1,0.1,0.1))
+
+```
 

### 2022-06-20 Final/FinalReport.qmd
visualization sizing

@@
 abortion_data$respectauthority <- as.factor(abortion_data$respectauthority)
-abortion_data$respectauthority <- factor(abortion_data$respectauthority, levels=c('0', '-1', '1'))
+abortion_data$respectauthority <- factor(abortion_data$respectauthority, 
+                                         levels=c('0', '-1', '1'))
 abortion_data$nationalpride <- as.factor(abortion_data$nationalpride)
@@
 The dataset used in this analysis observes attitudes on the
-justifiability of abortion among respondents in the United States
-across six "waves" of the World Values Survey (1982-2011), which is
+justifiability of abortion among respondents in the United States across
+six "waves" of the World Values Survey (1982-2011), which is
 administered every few years and collects information about people's
@@
 
-```{r vis-response}
+```{r vis-response, fig.height=3.5, fig.width=6}
 #| warning: false
@@
        x = "Abortion Justifiability (1-Least, 10-Most)",
-       caption = "Fig.1: The visual shows that the distribution of attitude toward abortion is trimodal, rather than bell-shaped. This may worsen our model's ability to fit the data.")+
+       caption = "Fig.1: The visual shows that the distribution of attitude 
+       toward abortion is trimodal, rather than bell-shaped. This may worsen 
+       our model's ability to fit the data.")+
   theme(plot.caption = element_text(hjust = 0.5))
@@
 
-```{r eda-histograms}
+```{r eda-histograms, fig.height=3.5, fig.width=6}
 #| warning: false
@@
   draw_label(
-    "Fig.2: The visuals show that while some variables in the data set have bell-shaped distributions, others are left-skewed.",
+    "Fig.2: The visuals show that while some variables in the data set 
+    have bell-shaped distributions, others are left-skewed.",
     x = 0,
@@
 
-```{r eda-boxplots}
+```{r eda-boxplots, fig.height=3.5, fig.width=6}
 #| warning: false
@@
 
-b5 <- ggplot(abortion_data, aes(x = respectauthority, group = respectauthority, 
+b5 <- ggplot(abortion_data, aes(x = respectauthority, 
+                                group = respectauthority, 
                                 y = aj)) +
@@
 caption2 <- ggdraw() + 
-  draw_label("Fig.3: The visuals show that there may be a relationship between the response and each of these variables, making them likely predictors.",
+  draw_label("Fig.3: The visuals show that there may be a relationship 
+             between the response and each of these variables, 
+             making them likely predictors.",
     x = 0,
@@
 
-```{r correlations}
+```{r correlations, fig.height=3.5, fig.width=6}
 #| warning: false
@@
 ggpairs(abortion_data, 
-  columns = c("year", "ideology", "cai", "godimportant", "respectauthority", "nationalpride"),
+  columns = c("year", "ideology", "cai", "godimportant", 
+              "respectauthority", "nationalpride"),
   columnLabels = c("YS", "PI", "CAI", "IG", 
@@
     ) +
-  labs(title = "Correlation Matrix of Predictors of Interest", caption = "Fig.4: This matrix shows that the correlations between pairs of predictors are low, with the highest being 0.370.")
+  labs(title = "Correlation Matrix of Predictors of Interest", 
+       caption = "Fig.4: This matrix shows that the correlations between 
+       pairs of predictors are low, with the highest being 0.370.")
 
 ```
+
 \pagebreak
@@
 We conducted a 75% - 25% data split into training and testing sets,
-using the random seed $206$. Fitting the MLR, we obtain the following table of parameter point estimates.
+using the random seed $206$. Fitting the MLR, we obtain the following
+table of parameter point estimates.
 
@@
 tidy(abortion_fit) %>%
-  kable(digits = 3, caption = "This table displays the estimate, standard error, statistic and p.value for each parameter of the model.")
+  kable(digits = 3, caption = "This table displays the estimate, 
+        standard error, statistic and p.value for 
+        each parameter of the model.")
 ```
@@
 
-```{r ci}
+```{r ci, fig.height=4, fig.width=6}
 #| echo: false
@@
 
-p2 <- ggplot(YS, aes(x = term, y = estimate, ymin = conf.low, ymax = conf.high)) +
+p2 <- ggplot(YS, aes(x = term, y = estimate, ymin = conf.low, 
+                     ymax = conf.high)) +
   geom_pointrange() +
@@
 
-p3 <- ggplot(CAI, aes(x = term, y = estimate, ymin = conf.low, ymax = conf.high)) +
+p3 <- ggplot(CAI, aes(x = term, y = estimate, ymin = conf.low, 
+                      ymax = conf.high)) +
   geom_pointrange()+
@@
 
-p4 <- ggplot(PI_IG, aes(x = term, y = estimate, ymin = conf.low, ymax = conf.high)) +
+p4 <- ggplot(PI_IG, aes(x = term, y = estimate, ymin = conf.low, 
+                        ymax = conf.high)) +
   geom_pointrange()+
@@
 
-p5 <- ggplot(RA_NP, aes(x = term, y = estimate, ymin = conf.low, ymax = conf.high)) +
+p5 <- ggplot(RA_NP, aes(x = term, y = estimate, ymin = conf.low, 
+                        ymax = conf.high)) +
   geom_pointrange()+
   labs(x = NULL) +
-  scale_x_discrete(breaks=c("respectauthority_X.1", "respectauthority_X1", "nationalpride_X1"),
+  scale_x_discrete(breaks=c("respectauthority_X.1", "respectauthority_X1", 
+                            "nationalpride_X1"),
         labels=c("RA=No", "RA=Yes", "NP=Yes"))
@@
 caption3 <- ggdraw() + 
-  draw_label("Fig.5: 95% confidence intervals of estimates for model parameters, using points and error bars.",
+  draw_label("Fig.5: 95% confidence intervals of estimates for model 
+             parameters, using points and error bars.",
     x = 0,
@@
 values of our model shows that our model is not very strong at
-predicting attitude toward the justifiability of abortion, with a R-squared
-of 0.247 and RMSE of 2.53.
+predicting attitude toward the justifiability of abortion, with a
+R-squared of 0.247 and RMSE of 2.53.
 
@@
 To see how well our model fits the data on the attitudes toward the
-justifiability of abortion, we conducted a 10-fold cross-validation with seed 206,
-calculating various metrics -- the Akaike information criterion (AIC),
-Bayesian Information Criterion (BIC) and adjusted R-squared value -- to
-evaluate the performance of our model on 10 subsets of our data.
+justifiability of abortion, we conducted a 10-fold cross-validation with
+seed 206, calculating various metrics -- the Akaike information
+criterion (AIC), Bayesian Information Criterion (BIC) and adjusted
+R-squared value -- to evaluate the performance of our model on 10
+subsets of our data.
 
@@
 
-```{r cv}
+```{r cv, fig.height=3.5, fig.width=6}
 #| message: false
@@
     geom_line() +
-    labs(title = "R-squared", caption = "Fig.6: This plot of R-squared values demonstrates our chosen model 1 is the best of the three.")
+    labs(title = "R-squared", caption = "Fig.6: This plot of R-squared 
+         values demonstrates our chosen model 1 is the best of the three.")
 ```
@@
 final_table %>%
-  kable(digits = 3, caption = "With the lowest AIC and BIC and highest adjusted R-squared value, our chosen model is best of the three.")
+  kable(digits = 3, caption = "With the lowest AIC and BIC and highest 
+        adjusted R-squared value, our chosen model is best of the three.")
 ```
@@
 
-As shown in the following plot of residuals against fitted values, the linearity condition is not met as the residuals are not scattered around the horizontal red line.
-The constant variance condition is also not met.
+As shown in the following plot of residuals against fitted values, the
+linearity condition is not met as the residuals are not scattered around
+the horizontal red line. The constant variance condition is also not
+met.
 
-```{r linearity}
+```{r linearity, fig.height=3.5, fig.width=5}
 #| echo: false
@@
   geom_hline(yintercept = 0, color = "red", linetype = "dashed") +
-  labs(title = "Plot of residuals v. fitted values", x = "Predicted values", y = "Residuals")
+  labs(title = "Plot of residuals v. fitted values", x = "Predicted values", 
+       y = "Residuals")
 
@@
   stat_function(fun = dnorm, 
-    args = list(mean = mean(final_fit_aug$.resid), sd = sd(final_fit_aug$.resid)), 
+    args = list(mean = mean(final_fit_aug$.resid), 
+                sd = sd(final_fit_aug$.resid)), 
     lwd = 2,  color = "red"  )+
@@
   geom_point(alpha = 0.5) +
-  labs(x = "Order of data collection", y = "Residuals", title = "Plot of residuals v. order of collection")
+  labs(x = "Order of data collection", y = "Residuals", 
+       title = "Plot of residuals v. order of collection")
 
@@
 caption4 <- ggdraw() + 
-  draw_label("Fig.7: The linearity condition and constant variance condition is clearly not met in this plot of residuals v. fitted values.",
+  draw_label("Fig.7: The linearity condition and constant variance condition 
+             is clearly not met in this plot of residuals v. fitted values.",
     x = 0,
@@
 caption5 <- ggdraw() + 
-  draw_label("Fig.8: The residuals appear to follow a normal distribution, which looks like a bell-shape, as illustrated in red.",
+  draw_label("Fig.8: The residuals appear to follow a normal distribution, 
+             which looks like a bell-shape, as illustrated in red.",
     x = 0,
@@
 caption6 <- ggdraw() + 
-  draw_label("Fig.9: The residuals plotted by observation order are randomly scattered, showing no structure.",
+  draw_label("Fig.9: The residuals plotted by observation order are randomly 
+             scattered, showing no structure.",
     x = 0,

### 2022-06-20 Final/FinalReport.qmd
fixed some captions

@@
   draw_label(
-    "Fig.2: The visuals show that while some variables in the data set 
-    have bell-shaped distributions, others are left-skewed.",
+    "Fig.2: The visuals show that while some variables have bell-shaped distributions, others are left-skewed.",
     x = 0,
@@
 caption2 <- ggdraw() + 
-  draw_label("Fig.3: The visuals show that there may be a relationship 
-             between the response and each of these variables, 
+  draw_label("Fig.3: The visuals show that there may be a relationship between the response and each of these variables, 
 making them likely predictors.",
@@
   labs(title = "Correlation Matrix of Predictors of Interest", 
-       caption = "Fig.4: This matrix shows that the correlations between 
-       pairs of predictors are low, with the highest being 0.370.")
+       caption = "Fig.4: This matrix shows that the correlations between pairs of predictors are low, with the 
+highest being 0.370.")
 
@@
 caption3 <- ggdraw() + 
-  draw_label("Fig.5: 95% confidence intervals of estimates for model 
-             parameters, using points and error bars.",
+  draw_label("Fig.5: 95% confidence intervals of estimates for model parameters, using points
+and error bars.",
     x = 0,
@@
     geom_line() +
-    labs(title = "R-squared", caption = "Fig.6: This plot of R-squared 
-         values demonstrates our chosen model 1 is the best of the three.")
+    labs(title = "R-squared", caption = 
+"Fig.6: This plot of R-squared values demonstrates our chosen model 1 is the 
+best of the three.")
 ```
@@
 final_table %>%
-  kable(digits = 3, caption = "With the lowest AIC and BIC and highest 
-        adjusted R-squared value, our chosen model is best of the three.")
+  kable(digits = 3, caption = 
+"With the lowest AIC and BIC and highest adjusted R-squared value, our chosen 
+model is best of the three.")
 ```
@@
 caption4 <- ggdraw() + 
-  draw_label("Fig.7: The linearity condition and constant variance condition 
-             is clearly not met in this plot of residuals v. fitted values.",
+  draw_label(
+"Fig.7: The linearity condition and constant variance condition is clearly not 
+met in this plot of residuals v. fitted values.",
     x = 0,
@@
 caption5 <- ggdraw() + 
-  draw_label("Fig.8: The residuals appear to follow a normal distribution, 
-             which looks like a bell-shape, as illustrated in red.",
+  draw_label(
+"Fig.8: The residuals appear to follow a normal distribution, which looks like 
+a bell-shape, as illustrated in red.",
     x = 0,
@@
 caption6 <- ggdraw() + 
-  draw_label("Fig.9: The residuals plotted by observation order are randomly 
-             scattered, showing no structure.",
+  draw_label(
+"Fig.9: The residuals plotted by observation order are randomly scattered, 
+showing no structure.",
     x = 0,

### 2022-06-21 Final/FinalReport.qmd
added conclusion + discussion section

@@
 
-\[paste here after finalizing the google doc\]
+The model that we fitted to predict abortion attitudes generally confirmed our initial research question that liberal ideology and attitudes towards other issues predicts abortion attitudes.
+Inference on the model (through both a confidence interval and hypothesis test) showed that having liberal political ideology, low belief in the importance of God, and belief in greater child autonomy generally predict a belief that abortion is more justified.
+However, holding liberal attitudes on certain other issues, namely believing that respect for authority is bad and not having pride in being American, do not necessarily predict greater justifiability of abortion.
+
+Despite the model confirming our initial research question, we have reason to question the quality of the model, and whether the model accurately fits the data on abortion attitudes.
+The first reason for questioning our model’s quality is the low R-squared value, which indicates that the variance of our predictors may not greatly account for the variance in abortion attitudes.
+The second reason for questioning our model’s quality is that our model does not fit the linearity and constant variance conditions.
+Upon closer examination of the data, we can see that the distribution of our response variable was trimodal, with peaks at the extremes and in the middle.
+This likely led to the data having a large number of very different observations that had different predictor values, but the same `aj` score, and may be one explanation our model’s weaknesses.
+
+Given the nature of public opinion on political issues, we are likely limited by the precision and accuracy of each observation in our dataset.
+Previous political science research on public opinion indicates that Americans’ political beliefs are inconsistent and volatile.
+For example, most Americans’ political beliefs have been found to be shaped by partisan and group loyalties, rather than ideology or any specific set of issue positions.
+Indeed, even for committed partisans, one individual’s issue positions are likely to to be scattered across the ideological spectrum, and may even change over short periods of time (Achen and Bartels, 2017).
+This inconsistency between partisanship, ideology, and issue positions may account for the non-normal distribution of abortion attitudes in the dataset and the model’s poor fit to the data.
+In addition, although the data are supposed to capture a nationally representative sample of the US population, we do not know if they are weighted by other factors that may represent important distinctions in the US population.
+This possibility of non-weighting may have influenced our model’s quality as well.
+
+Future work may build on this analysis and correct some of its weaknesses.
+First, the data should be weighted by more relevant factors to ensure that the sample is as representative as possible of the US population.
+Relevant factors may include, but are not limited to, partisan affiliation, race, religion, socioeconomic status, and geographic area.
+In addition, it may be helpful to consider other predictors that may influence abortion attitudes, such as partisan affiliation and race.
+However, it is important to keep in mind that abortion is an incredibly divisive and complex issue, so it is not an easy task to precisely pinpoint any individual’s abortion attitude, no matter which predictors are used.
 
@@
 
+Achen, C.H., & Bartels, L.M. (2017). Democracy for realists: Why elections do not produce responsive government. Princeton University Press.
+
 Brenan, M. (2020, July 7). One in Four Americans Consider Abortion a Key
@@
 https://news.gallup.com/poll/313316/one-four-americans-consider-abortion-key-voting-issue.aspx
+
+Miller, Steven V. (2016, Jul 31).  “Attitudes on the Justifiability of Abortion in the United States.”  GitHub.  https://github.com/svmiller/wvs-usa-abortion-attitudes
 

### 2022-06-21 previous-work/draft/draft.qmd
FINAL REORG

--- 2022-06-15 draft/draft.qmd
@@
 #| message: false
-abortion_data_full <- read_csv(here::here("data/abortion-attitudes", "wvs-usa-abortion-attitudes-data.csv"))
+abortion_data_full<-read_csv(here::here("data/abortion-attitudes", 
+                                        "wvs-usa-abortion-attitudes-data.csv"))
 ```
@@
 
-Understanding public attitudes on divisive political issues is an important way for political leaders to mobilize voters and for lawmakers to draft laws that represent their constituents.
-While it is simple to poll constituents’ positions on an issue, it can be challenging to assess the complex factors that influence and predict those stances.
+Understanding public attitudes on divisive political issues is an
+important way for political leaders to mobilize voters and for lawmakers
+to draft laws that represent their constituents. While it is simple to
+poll constituents' positions on an issue, it can be challenging to
+assess the complex factors that influence and predict those stances.
 
-Abortion is one such divisive issue in the United States, with strong organization of pro-choice and pro-life groups in states across the country in support of legislation for their respective sides (Ziegler, 2020).
-However, following the Supreme Court’s 1973 decision in Roe v.
-Wade, a ruling that has historically protected women's rights to abortion without excessive government restriction in America, the issue has risen in political salience.
-Both the pro-choice and pro-life movements have gained national prominence, and the two major political parties have polarized around the issue, with the Democratic Party in favor of and the Republican Party against policies legalization and increased access to abortion (Weinberger 2022).
-Abortion has also increasingly become a key issue that voters consider, with an increasing share of Americans identifying as “single-issue voters” regarding abortion (Brenan 2020).
+Abortion is one such divisive issue in the United States, with strong
+organization of pro-choice and pro-life groups in states across the
+country in support of legislation for their respective sides (Ziegler,
+2020). However, following the Supreme Court's 1973 decision in Roe v.
+Wade, a ruling that has historically protected women's rights to
+abortion without excessive government restriction in America, the issue
+has risen in political salience. Both the pro-choice and pro-life
+movements have gained national prominence, and the two major political
+parties have polarized around the issue, with the Democratic Party in
+favor of and the Republican Party against policies legalization and
+increased access to abortion (Weinberger 2022). Abortion has also
+increasingly become a key issue that voters consider, with an increasing
+share of Americans identifying as "single-issue voters" regarding
+abortion (Brenan 2020).
 
-In May 2022, a leaked draft opinion revealed that the US Supreme Court is prepared to overturn Roe v.
-Wade.
-Overturning Roe would dramatically change the trajectory of abortion politics in the US.
-Unless the US Congress passed a national policy, states would be able to decide whether or not to legalize abortion and gain much greater leverage in regulating access to the procedure (Weinberger 2022).
+In May 2022, a leaked draft opinion revealed that the US Supreme Court
+is prepared to overturn Roe v. Wade. Overturning Roe would dramatically
+change the trajectory of abortion politics in the US. Unless the US
+Congress passed a national policy, states would be able to decide
+whether or not to legalize abortion and gain much greater leverage in
+regulating access to the procedure (Weinberger 2022).
 
-Given the potential overturning of Roe and the polarizing nature of the issue, it is important to understand how the American public feels about whether abortion should be legal or not, how accessible the procedure should be, and what factors influence these opinions.
-Understanding public opinion on the issue will ensure that political leaders are able to mobilize the correct constituencies, and that policy experts are able to pass policies on this issue that accurately reflect the preferences of the American people.
+Given the potential overturning of Roe and the polarizing nature of the
+issue, it is important to understand how the American public feels about
+whether abortion should be legal or not, how accessible the procedure
+should be, and what factors influence these opinions. Understanding
+public opinion on the issue will ensure that political leaders are able
+to mobilize the correct constituencies, and that policy experts are able
+to pass policies on this issue that accurately reflect the preferences
+of the American people.
 
@@
 
-We will conduct our analysis on data collected about Americans’ abortion attitudes, demographic information, and other ideologies, using an EDA-informed predictive model.
-The data are a nationally representative sample of the American people; hence we can infer that all the observations are independent and the variables may show a linear relationship.
-Given that this issue has become highly polarized by political party, we predict that liberal political attitudes and youth will correlate with belief that abortion is more justified.
+We will conduct our analysis on data collected about Americans' abortion
+attitudes, demographic information, and other ideologies, using an
+EDA-informed predictive model. The data are a nationally representative
+sample of the American people; hence we can infer that all the
+observations are independent and the variables may show a linear
+relationship. Given that this issue has become highly polarized by
+political party, we predict that liberal political attitudes and youth
+will correlate with belief that abortion is more justified.
 
@@
 
-In this project, we investigate the correlation between attitudes on the justifiability of abortion with demographic factors (such as age, gender, and education level) and personal attitudes towards other issues (such as political ideology, importance of religion, and respect for authority) in a representative sample of American citizens from 1982-2011.
+In this project, we investigate the correlation between attitudes on the
+justifiability of abortion with demographic factors (such as age,
+gender, and education level) and personal attitudes towards other issues
+(such as political ideology, importance of religion, and respect for
+authority) in a representative sample of American citizens from
+1982-2011.
 
-The response variable, `Justifiability of abortion`, is a numerical measure on a scale of 1 to 10 on the individual person’s attitude toward whether abortion is justifiable or not.
-Individuals responded 1 for “abortion is never justified” and 10 for “abortion is always justified.
+The response variable, `Justifiability of abortion`, is a numerical
+measure on a scale of 1 to 10 on the individual person's attitude toward
+whether abortion is justifiable or not. Individuals responded 1 for
+"abortion is never justified" and 10 for "abortion is always justified.
 
@@
 
-We create a visualization and summary statistics for the response variable.
+We create a visualization and summary statistics for the response
+variable.
 
@@
 #| warning: false
-ggplot(abortion_data, aes(x = aj)) +
+ggplot(abortion_data_full, aes(x = aj)) +
   geom_histogram(binwidth = 1, aes(y=..density..)) + 
   geom_density(color = "red") +
-  labs(title = "Histogram of abortion attitudes")
+  labs(title = "Histogram of Abortion Attitudes",
+       x = "Abortion Justifiability (1-Least, 10-Most)")
 
-summary(abortion_data$aj)
+summary(abortion_data_full$aj)
 ```
@@
 #Year of survey
-ggplot(abortion_data, aes(x = year)) +
+ggplot(abortion_data_full, aes(x = year)) +
   geom_histogram() + 
   labs(title = "Histogram of year")
-ggplot(abortion_data, aes(group = year, y = aj)) +
+ggplot(abortion_data_full, aes(group = year, y = aj)) +
   geom_boxplot() 
@@
 # ideology
-ggplot(abortion_data, aes(x = ideology)) +
+ggplot(abortion_data_full, aes(x = ideology)) +
   geom_histogram() + 
   labs(title = "Histogram of ideology")
-ggplot(abortion_data, aes(x = ideology,group = ideology, y = aj)) +
+ggplot(abortion_data_full, aes(x = ideology,group = ideology, y = aj)) +
   geom_boxplot()
@@
 # Child autonomy index
-ggplot(abortion_data, aes(x = cai)) +
+ggplot(abortion_data_full, aes(x = cai)) +
   geom_histogram() + 
   labs(title = "Histogram of Child autonomy index")
-ggplot(abortion_data, aes(x = cai, group = cai, y = aj)) +
+ggplot(abortion_data_full, aes(x = cai, group = cai, y = aj)) +
   geom_boxplot()
@@
 # Importance of God
-ggplot(abortion_data, aes(x = godimportant)) +
+ggplot(abortion_data_full, aes(x = godimportant)) +
   geom_histogram() + 
   labs(title = "Histogram of how respondent saw God's importance")
-ggplot(abortion_data, aes(x = godimportant, group = godimportant, y = aj)) +
+ggplot(abortion_data_full, aes(x = godimportant, group = godimportant, 
+                               y = aj)) +
   geom_boxplot()
@@
 # Respect for authority
-ggplot(abortion_data, aes(x = respectauthority)) +
+ggplot(abortion_data_full, aes(x = respectauthority)) +
   geom_histogram() + 
   labs(title = "Histogram of respect for authority")
-ggplot(abortion_data, aes(x = respectauthority, group = respectauthority, y = aj)) +
+ggplot(abortion_data_full, aes(x = respectauthority, group = respectauthority, 
+                          y = aj)) +
   geom_boxplot()
@@
 # National Pride
-ggplot(abortion_data, aes(x = nationalpride)) +
+ggplot(abortion_data_full, aes(x = nationalpride)) +
   geom_histogram() + 
   labs(title = "Histogram of National pride")
-ggplot(abortion_data, aes(x = nationalpride, group = nationalpride, y = aj)) +
+ggplot(abortion_data_full, aes(x = nationalpride, group = nationalpride, y = aj)) +
   geom_boxplot()
@@
 #| warning: false
-ggpairs(abortion_data, 
+ggpairs(abortion_data_full, 
   columns = c("year", "ideology", "cai", "godimportant", "respectauthority", 
               "nationalpride"),
-  columnLabels = c("year", "ideology", "child autonomy", "imptnce of God", "resp for autho", 
-              "natl pride")) +
+  columnLabels = c("year", "ideology", "child autonomy", "imptnce of God", 
+                   "resp for autho", "natl pride")) +
    theme(
@@
 Recipe1:
+
 -   Year
@@
 We also wanted to choose predictor variables that are not strongly
-correlated with each other, to avoid multicollinearity. From our EDA, we
-found that several predictors had correlations of 0.2 or higher with
+correlated with each other, to avoid multi-collinearity. From our EDA,
+we found that several predictors had correlations of 0.2 or higher with
 each other. Given these correlations, we decided to make a second recipe
@@
 
-## Fit Model
+We'll now split each of the three "recipes" into a training set (75%)
+and a testing set (25%)
 
-Note: make sure to create recipes, specify engine, make workflows, and
-fit models here
+```{r splits}
+set.seed(206)
+abortion_split <- initial_split(abortion_data)
+abortion_training <- training(abortion_split)
+abortion_testing <- testing(abortion_split)
+```
 
-## Inference
+## Fit Models
 
-\pagebreak
+Now that the data is split, we will specify the models and create
+recipes for each.
 
-##RESULTS
+```{r spec}
+abortion_spec <- linear_reg() %>%
+  set_engine("lm")
+```
 
-## Model Evaluation
+Now that our model is specified, we will create 3 recipes, each with the
+corresponding predictors listed above.
 
-## Model Comparison
+The first recipe will be predicting abortion attitudes from `year`,
+`ideology`, `cai`, `godimportant`, `respectauthority`, and
+`nationalpride`. We will center the variable `year` so that we can have
+a meaningful intercept, create dummy variables when needed, and
+eliminate any zero variance predictors.
 
-## Model Diagnosis
+```{r recipe1}
+abortion_rec1 <- recipe(aj ~ year + ideology + cai + godimportant +
+                        respectauthority, nationalpride,
+                        data = abortion_data) %>%
+  step_center(year) %>%
+  step_dummy(all_nominal_predictors()) %>%
+  step_zv(all_predictors())
 
-## Discussion of Influential Points
+abortion_rec1
+```
+
+Our second recipe will be predicting abortion attitudes from just `year`
+and `godimportant`. We will once again center the variable `year` so
+that we can have a meaningful intercept, create dummy variables when
+needed, and eliminate any zero variance predictors.
+
+```{r recipe2}
+abortion_rec2 <- recipe(aj ~ year + godimportant,
+                        data = abortion_data) %>%
+  step_center(year) %>%
+  step_dummy(all_nominal_predictors()) %>%
+  step_zv(all_predictors())
+
+abortion_rec2
+```
+
+Our third and final recipe will predict abortion attitudes from `year`,
+`ideology`, `cai`, and `nationalpride`. We will take the same steps from
+the last 2 recipes.
+
+```{r recipe3}
+abortion_rec3 <- recipe(aj ~ year + ideology + cai + nationalpride,
+                        data = abortion_data) %>%
+  step_center(year) %>%
+  step_dummy(all_nominal_predictors()) %>%
+  step_zv(all_predictors())
+
+abortion_rec3
+```
+
+With the recipes made, we are ready to create our workflows.
+
+```{r workflows}
+abortion_wflow1 <- workflow() %>%
+  add_model(abortion_spec) %>%
+  add_recipe(abortion_rec1)
+
+abortion_wflow2 <- workflow() %>%
+  add_model(abortion_spec) %>%
+  add_recipe(abortion_rec2)
+
+abortion_wflow3 <- workflow() %>%
+  add_model(abortion_spec) %>%
+  add_recipe(abortion_rec3)
+```
+
+And finally, we will fit each of the models above
+
+```{r fit-models}
+abortion_fit1 <- abortion_wflow1 %>%
+  fit(data = abortion_training)
+tidy(abortion_fit1) %>%
+  kable(digits = 3)
+
+abortion_fit2 <- abortion_wflow2 %>%
+  fit(data = abortion_training)
+tidy(abortion_fit2) %>%
+  kable(digits = 3)
+
+abortion_fit3 <- abortion_wflow3 %>%
+  fit(data = abortion_training)
+tidy(abortion_fit3) %>%
+  kable(digits = 3)
+```
+
+## Model Comparison and Selection
+
+Now that we have all of our models fit and displayed, we need to
+interpret the results and determine the best model.
+
+Let's start by seeing how well our models predict the abortion attitudes
+
+```{r predict}
+abortion_training_pred1 <- predict(abortion_fit1, abortion_training) %>%
+  bind_cols(abortion_training %>% select(aj, year, ideology, cai, 
+                                         godimportant, respectauthority, 
+                                         nationalpride))
+abortion_training_pred1
+
+abortion_training_pred2 <- predict(abortion_fit2, abortion_training) %>%
+  bind_cols(abortion_training %>% select(aj, year, godimportant))
+abortion_training_pred2
+
+abortion_training_pred3 <- predict(abortion_fit3, abortion_training) %>%
+  bind_cols(abortion_training %>% select(aj, year, ideology, cai, 
+                                         nationalpride))
+abortion_training_pred3
+```
+
+A quick glance at our data might seem that the models are not that
+strong at predicting the abortion attitude (aj) of a given observation.
+We will look into this further now with cross-validation, then a test in
+AIC and BIC statistics.
+
+```{r cv}
+set.seed(206)
+folds <- vfold_cv(abortion_training, v = 10)
+abortion_fit_rs1 <- abortion_wflow1 %>%
+  fit_resamples(folds)
+
+abortion_fit_rs2 <- abortion_wflow2 %>%
+  fit_resamples(folds)
+
+abortion_fit_rs3 <- abortion_wflow3 %>%
+  fit_resamples(folds)
+```
+
+```{r cv-table}
+cv_metrics1 <- collect_metrics(abortion_fit_rs1, summarize = FALSE) 
+cv_metrics2 <- collect_metrics(abortion_fit_rs2, summarize = FALSE) 
+cv_metrics3 <- collect_metrics(abortion_fit_rs3, summarize = FALSE)
+
+cv_metrics1 %>%
+  mutate(.estimate = round(.estimate, 3)) %>%
+  pivot_wider(id_cols = id, names_from = .metric, values_from = .estimate) %>%
+  kable(col.names = c("Fold", "RMSE", "R-squared"))
+
+cv_metrics2 %>%
+  mutate(.estimate = round(.estimate, 3)) %>%
+  pivot_wider(id_cols = id, names_from = .metric, values_from = .estimate) %>%
+  kable(col.names = c("Fold", "RMSE", "R-squared"))
+
+cv_metrics3 %>%
+  mutate(.estimate = round(.estimate, 3)) %>%
+  pivot_wider(id_cols = id, names_from = .metric, values_from = .estimate) %>%
+  kable(col.names = c("Fold", "RMSE", "R-squared"))
+```
+
+As with the predictions from before, the RMSE and R-squared values for
+each of these models do not show promising results. Of the three models,
+the first is seemingly the strongest model as it has the closest
+R-squared values to 1. We will continue our analysis still with AIC and
+BIC analysis.
+
+```{r aic-bic}
+glance(abortion_fit1) %>% 
+  select(AIC, BIC)
+
+glance(abortion_fit2) %>% 
+  select(AIC, BIC)
+
+glance(abortion_fit3) %>% 
+  select(AIC, BIC)
+```
+
+From the above AIC and BIC values, we have further evidence that model 1
+is the strongest model of the three. Model 1 has the lowest AIC and BIC
+values, which are usually penalized for having more predictors, but in
+this case the model is strong enough to overcome that penalty.
+
+## Inference: Confidence Interval and Hypothesis Test
+
+Before we discuss results, we will take a quick look at the resulting
+confidence intervals from our fitted models and the resulting
+coefficients.
+
+```{r ci}
+tidy(abortion_fit1, conf.int = TRUE) %>%
+  kable(digits = 3)
+tidy(abortion_fit2, conf.int = TRUE) %>%
+  kable(digits = 3)
+tidy(abortion_fit3, conf.int = TRUE) %>%
+  kable(digits = 3)
+```
+
+## Results
+
+From the analysis of the models through multiple different calculations,
+we know that of our three models, model 1 is the strongest, predicting
+abortion attitude from `year`, `ideology`, `cai`, `godimportant`,
+`respectauthority`, and `nationalpride`. Given that model 1 is our
+strongest model, it must also be noted that all 3 of our models had
+difficulty predicting an accurate attitude towards abortion of a given
+observation. This is most likely due to the data having a large number
+of very different observations given the same `aj` score. In model 1,
+however, our p-values for the coefficients given are all at an
+acceptable range except for that of `nationalpride`. This means that
+every other predictor was able to be helpful in predicting the response
+variable.
+
+## Discussion + Conclusion
+
+Ultimately, our investigation in the correlation between attitudes on
+the justifiability of abortion and demographic factors + personal
+attitudes towards other issues presented some problems. Through our
+model creation and comparison, we are able to determine that the best
+combination of predictors for abortion attitudes are `year`, `ideology`,
+`cai`, `godimportant`, `respectauthority`, and `nationalpride`. While
+every one of these predictors are shown to be helpful in predicting the
+response variable, the model itself is generally weak although it is the
+best of our 3.
+
+Unfortunately, given our data set, we are limited by the truthfulness
+and accuracy of each observation. While we can assume each observation
+is answered honestly and thoughtfully, there was the option to leave
+some answers empty, and some predictors are quite hard to place nicely
+into a 1-10 scale (take ideology for example). Steps that could be taken
+to counter this problem are to either categorize some of these 1-10
+scales into even smaller categories (like 3/4) to possibly show better
+trends, or even more categories (like 1-50 scale) to increase accuracy
+of responses and also possibly reveal more trends.
+
+With abortion being such a divisive issue in our society, it is
+incredibly difficult to pinpoint just a few predictors to try and
+accurately tell how any one person might feel about abortion. Knowing
+that, however, we do not expect that abortion attitudes are impossible
+to more accurately predict. We have shown through our modeling and
+analysis that some of the predictors presented are helpful in predicting
+our response variable. Given this, it is not unreasonable to assume that
+there might be even more helpful predictors to use in place, or in
+addition to the ones we have selected. For future work, we might be able
+to produce stronger models given additional demographic/personality
+predictors to work with.
 
@@
 Brenan, M. (2020, July 7). One in Four Americans Consider Abortion a Key
-Voting Issue. Gallup. https://news.gallup.com/poll/313316/one-four-americans-consider-abortion-key-voting-issue.aspx
+Voting Issue. Gallup.
+https://news.gallup.com/poll/313316/one-four-americans-consider-abortion-key-voting-issue.aspx
 
 Weinberger, J. (2022, May 6). How we got here: Roe v. Wade from 1973 to
-today. Vox. https://www.vox.com/23055389/roe-v-wade-timeline-abortion-overturn-political-polarization
+today. Vox.
+https://www.vox.com/23055389/roe-v-wade-timeline-abortion-overturn-political-polarization
 
 Ziegler, M. (2020, October 22). Abortion politics polarized before Roe.
-When it's gone, the fighting won't stop. The Washington Post. https://www.washingtonpost.com/outlook/2020/10/22/roe-polarize-abortion-politics/
+When it's gone, the fighting won't stop. The Washington Post.
+https://www.washingtonpost.com/outlook/2020/10/22/roe-polarize-abortion-politics/

### 2022-06-21 written-report.qmd
FINAL REORG

--- 2022-06-21 Final/FinalReport.qmd
@@
 plot_grid(title_factor, top_row, title_numeric, bottom_row, caption3, 
-          ncol = 1, rel_heights = c(0.1, 1,0.1,1,0.1))
+          ncol = 1, rel_heights = c(0.1, 1,0.1,1,0.2))
 
@@
 plot_grid(top_row1, c3, caption4, caption5, caption6,
-          ncol = 1, rel_heights = c(1, 1,0.1,0.1,0.1))
+          ncol = 1, rel_heights = c(1, 1,0.2,0.2,0.2))
 
@@
 
-The model that we fitted to predict abortion attitudes generally confirmed our initial research question that liberal ideology and attitudes towards other issues predicts abortion attitudes.
-Inference on the model (through both a confidence interval and hypothesis test) showed that having liberal political ideology, low belief in the importance of God, and belief in greater child autonomy generally predict a belief that abortion is more justified.
-However, holding liberal attitudes on certain other issues, namely believing that respect for authority is bad and not having pride in being American, do not necessarily predict greater justifiability of abortion.
+The model that we fitted to predict abortion attitudes generally
+confirmed our initial research question that liberal ideology and
+attitudes towards other issues predicts abortion attitudes. Inference on
+the model (through both a confidence interval and hypothesis test)
+showed that having liberal political ideology, low belief in the
+importance of God, and belief in greater child autonomy generally
+predict a belief that abortion is more justified. However, holding
+liberal attitudes on certain other issues, namely believing that respect
+for authority is bad and not having pride in being American, do not
+necessarily predict greater justifiability of abortion.
 
-Despite the model confirming our initial research question, we have reason to question the quality of the model, and whether the model accurately fits the data on abortion attitudes.
-The first reason for questioning our model’s quality is the low R-squared value, which indicates that the variance of our predictors may not greatly account for the variance in abortion attitudes.
-The second reason for questioning our model’s quality is that our model does not fit the linearity and constant variance conditions.
-Upon closer examination of the data, we can see that the distribution of our response variable was trimodal, with peaks at the extremes and in the middle.
-This likely led to the data having a large number of very different observations that had different predictor values, but the same `aj` score, and may be one explanation our model’s weaknesses.
+Despite the model confirming our initial research question, we have
+reason to question the quality of the model, and whether the model
+accurately fits the data on abortion attitudes. The first reason for
+questioning our model's quality is the low R-squared value, which
+indicates that the variance of our predictors may not greatly account
+for the variance in abortion attitudes. The second reason for
+questioning our model's quality is that our model does not fit the
+linearity and constant variance conditions. Upon closer examination of
+the data, we can see that the distribution of our response variable was
+trimodal, with peaks at the extremes and in the middle. This likely led
+to the data having a large number of very different observations that
+had different predictor values, but the same `aj` score, and may be one
+explanation our model's weaknesses.
 
-Given the nature of public opinion on political issues, we are likely limited by the precision and accuracy of each observation in our dataset.
-Previous political science research on public opinion indicates that Americans’ political beliefs are inconsistent and volatile.
-For example, most Americans’ political beliefs have been found to be shaped by partisan and group loyalties, rather than ideology or any specific set of issue positions.
-Indeed, even for committed partisans, one individual’s issue positions are likely to to be scattered across the ideological spectrum, and may even change over short periods of time (Achen and Bartels, 2017).
-This inconsistency between partisanship, ideology, and issue positions may account for the non-normal distribution of abortion attitudes in the dataset and the model’s poor fit to the data.
-In addition, although the data are supposed to capture a nationally representative sample of the US population, we do not know if they are weighted by other factors that may represent important distinctions in the US population.
-This possibility of non-weighting may have influenced our model’s quality as well.
+Given the nature of public opinion on political issues, we are likely
+limited by the precision and accuracy of each observation in our
+dataset. Previous political science research on public opinion indicates
+that Americans' political beliefs are inconsistent and volatile. For
+example, most Americans' political beliefs have been found to be shaped
+by partisan and group loyalties, rather than ideology or any specific
+set of issue positions. Indeed, even for committed partisans, one
+individual's issue positions are likely to to be scattered across the
+ideological spectrum, and may even change over short periods of time
+(Achen and Bartels, 2017). This inconsistency between partisanship,
+ideology, and issue positions may account for the non-normal
+distribution of abortion attitudes in the dataset and the model's poor
+fit to the data. In addition, although the data are supposed to capture
+a nationally representative sample of the US population, we do not know
+if they are weighted by other factors that may represent important
+distinctions in the US population. This possibility of non-weighting may
+have influenced our model's quality as well.
 
-Future work may build on this analysis and correct some of its weaknesses.
-First, the data should be weighted by more relevant factors to ensure that the sample is as representative as possible of the US population.
-Relevant factors may include, but are not limited to, partisan affiliation, race, religion, socioeconomic status, and geographic area.
-In addition, it may be helpful to consider other predictors that may influence abortion attitudes, such as partisan affiliation and race.
-However, it is important to keep in mind that abortion is an incredibly divisive and complex issue, so it is not an easy task to precisely pinpoint any individual’s abortion attitude, no matter which predictors are used.
+Future work may build on this analysis and correct some of its
+weaknesses. First, the data should be weighted by more relevant factors
+to ensure that the sample is as representative as possible of the US
+population. Relevant factors may include, but are not limited to,
+partisan affiliation, race, religion, socioeconomic status, and
+geographic area. In addition, it may be helpful to consider other
+predictors that may influence abortion attitudes, such as partisan
+affiliation and race. However, it is important to keep in mind that
+abortion is an incredibly divisive and complex issue, so it is not an
+easy task to precisely pinpoint any individual's abortion attitude, no
+matter which predictors are used.
 
@@
 
-Achen, C.H., & Bartels, L.M. (2017). Democracy for realists: Why elections do not produce responsive government. Princeton University Press.
+Achen, C.H., & Bartels, L.M. (2017). Democracy for realists: Why
+elections do not produce responsive government. Princeton University
+Press.
 
@@
 
-Miller, Steven V. (2016, Jul 31).  “Attitudes on the Justifiability of Abortion in the United States.”  GitHub.  https://github.com/svmiller/wvs-usa-abortion-attitudes
+Miller, Steven V. (2016, Jul 31). "Attitudes on the Justifiability of
+Abortion in the United States." GitHub.
+https://github.com/svmiller/wvs-usa-abortion-attitudes
 
